### 2015-05-09 reports/paper/experiment.tex
paper update

\section{Experiment}
To test localization accuracy, an one meter by one meter grid was set up and the arrays are placed at the top left and top right corners of the grid. Fig~\ref{fig:setup_point} shows a picture of the setup.

\begin{figure}[]
  \centering
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{setup_1.JPG}
    \caption{1 meter by 1 meter grid}
  \end{subfigure}
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{setup_2.JPG}
    \caption{array placement}
  \end{subfigure}
  \caption{Setup for localization accuracy testing}
  \label{fig:setup_point}
\end{figure}

A total of $32$ positions are chosen uniformly in this region where microphone data is recorded. To test how accuracy varies with window size, the algorithm is fed with recorded microphone data with different segment length. Fig~\ref{fig:accuracy_vs_window} shows how accuracy changes with window length for three GCC algorithms. The error lowers as window size increases and plataeus after window size exceeds around $10$ milli-second. The lowest error achieved is $2.53$ centimeters. It is achieved when window size is set to 12 millisecond and GCC\_PHAT is used to estimate difference of delay

Although accuracy improves with window length, the calculation time also increases with window length. The part of calculation that depends on window length is using cross correlation to estimate difference of arrival time. cross correlation can be calculated with FFT and the runtime is of order $O(N\log N)$. We measured how the computation time varies with window length and Figure~\ref{fig:speed_vs_window} shows the result. The runtime increases approximately linearly in the window size region of interest.

We also calculated the localization error for each tested point in the region. Figure~\ref{fig:error_distribution} shows a heatmap of the error distribution inside the grid. The error is below $3$ cm for most areas inside the region. There is one error spike in the mid-left region and we contribute this to audio source placecment error becuase the error is fairly low and consistent around that spike region.

\begin{figure}[]
\includegraphics[width=0.5\textwidth]{average_error_window_size.eps}
\caption{accuracy versus window size}
\label{fig:accuracy_vs_window}
\end{figure}

\begin{figure}[]
\includegraphics[width=0.5\textwidth]{calculation_time.eps}
\caption{speed versus window size}
\label{fig:speed_vs_window}
\end{figure}

\begin{figure}[]
\includegraphics[width=0.5\textwidth]{error_distribution.eps}
\caption{error distribution in the grid}
\label{fig:error_distribution}
\end{figure}

\begin{figure}[]
  \centering
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{setup_ring.JPG}
  \end{subfigure}
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{setup_rotating_disk.JPG}
  \end{subfigure}
  \caption{Setup for circle movement localization}
  \label{fig:setup_circle}
\end{figure}

To test how well the arrays track movement, we mounted a rotating disk $40$ centimeter in diameter onto the grid at $(x=0, y=-0.3)$. Fig~\ref{fig:setup_circle} shows a picture of the sestup. A sound source is placed on the edge of rotating disk and the arrays localizes the sound source as it rotates in a circle. In this experiment, we want to test how accuracy changes with:
\begin{itemize}
\item different window sizes
\item different audio sources
\item different movement tracking filters
\item different movement speeds
\end{itemize}

\begin{figure*}[]
\centering
  \includegraphics[width=\textwidth]{trace_window_size_movement.eps}
\caption{Localization quality versus window size}\label{fig:wn}
\label{fig:trace_win_circle}
\end{figure*}

\begin{figure}[]
\includegraphics[width=0.5\textwidth]{error_window_size_movement.eps}
\caption{Localization error versus window size}
\label{fig:err_win_circle}
\end{figure}

Fig~\ref{fig:err_win_circle} gives an intuitive representation of how accuracy changes with window size. When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate arrival time differences. The localization is very noisy. As window size increases, the localization converges to the shape of ground truth circle. Fig~\ref{fig:err_win_circle} shows how the error changes with window size. The general trend is similar to that in point localization case. The error decreases as window size increases and plateaus after window size exceeds around $10$ milliseconds.

To test how different sound sources impact localization quality, three experiments on the same movement track were conducted with three different sound sources:

\begin{description}

\item[White Noise] A recording of white noise.

\item[Music A] A music that has normal audio amplitude throughout experiment period was chosen. "Honest Eyes" by Black Tide was chosen

\item[Music B] A music with intermittent low amplitude sections was chosen. "Canon" was used.

\end{description}

To test how audio movement source affects localization quality, each of three experiments were conducted at two different speeds:
\begin{description}
\item[Normal] An angular speed of $0.5$ rad/s was maintained, which translates to linear speed of $10$ cm/s.
\item[Fast] An angular speed of $1.0$ rad/s was maintained, which translates to linear speed of $20$ cm/s.
\end{description}

For each experiment conducted, two different movement filters are tests:
\begin{description}
\item[Averaging filter] localization for past $0.5$ seconds are averaged and outputed as current estimate.
\item[Kalman filter] A 2nd order Kalman filtering is used.
\end{description}

\begin{figure*}[]
\centering
\begin{subfigure}[]{1.0\textwidth}
  \includegraphics[width=1.0\textwidth]{trace_output_circle_wnn.eps}
  \caption{white noise}
  \label{fig:circle_wn}
\end{subfigure}
\begin{subfigure}[]{1.0\textwidth}
  \includegraphics[width=\textwidth]{trace_output_circle_man.eps}
  \caption{music A}
  \label{fig:circle_musica}
\end{subfigure}
\begin{subfigure}[]{1.0\textwidth}
  \includegraphics[width=\textwidth]{trace_output_circle_mbn.eps}
  \caption{music B}
  \label{fig:circle_musicb}
\end{subfigure}
\caption{Localization of circle movement with different sound sources. Sound source is moving at $10$ cm per second}
\label{fig:circle_normal}
\end{figure*}

Fig~\ref{fig:circle_normal} shows results for experiments with three different audio sources at normal speed.

with three different audio source were used for the same circular movement. In the first trial, white noise is used to set a baseline for audio sources, since our understanding is that white noise performs the best under GCC\_PHAT. In the second trial, a randomly picked music is used. In the third trial, a music that contains low amplitude segments is picked to reflect real life situations. Fig~\ref{fig:circle} shows the result for all three trials. From the result, white noise does perform the best and the localized shape matches well with the shape of the rotating disk. There exists a bit of jiggling, but can be improved with filtering. Fig~\ref{fig:circle} also shows the result with averaging filtering and Kalman filtering.

For the music A, the array still localizaes correctly and the output represetns a circular shape, but the localizaed points are more spread out with more jiggling. This is because music does not have components in all frequency bands and GCC\_PHAT estimate is less accuracy with such audio sources. Similar to the white noise case, we can improve the raw tracking estimates from the array with filtering.

For music B, the music has a low amplitude section and it is also shown in the traced result. The "blank" region in the output corresponds to the low amplitude segment in the audio. The general tracking quality of this music is not as good as that with Music A due to these low amplitude audio segments.

\begin{figure*}[]
\centering
\begin{subfigure}[]{1.0\textwidth}
  \includegraphics[width=\textwidth]{trace_output_circle_wnf.eps}
  \caption{white noise}
  \label{fig:circle_wn}
\end{subfigure}
\begin{subfigure}[]{1.0\textwidth}
  \includegraphics[width=\textwidth]{trace_output_circle_maf.eps}
  \caption{music A}
  \label{fig:circle_musica}
\end{subfigure}
\begin{subfigure}[]{1.0\textwidth}
  \includegraphics[width=\textwidth]{trace_output_circle_mbf.eps}
  \caption{music B}
  \label{fig:circle_musicb}
\end{subfigure}
\caption{Localization of circle movement with different sound sources. Sound source is moving at $20$ cm per second}
\label{fig:circle_fast}
\end{figure*}

### 2015-05-12 reports/paper/experiment.tex
update paper

@@
 A sound source is placed on the edge of rotating disk and the arrays localizes the sound source as it rotates in a circle.
-In this experiment, we want to test how accuracy changes with:
+In this experiment, we tested how accuracy changes with:
 \begin{itemize}
@@
 When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate arrival time differences.
-The localization is very noisy.
+The localization is noisy.
 As window size increases, the localization converges to the shape of ground truth circle.
@@
 
-To test how different sound sources impact localization quality, three experiments on the same movement track were conducted with three different sound sources:
+To test how different sound sources impact localization quality, we conducted three experiments on the same movement track with three different sound sources:
 
-\begin{description}
+\begin{description}[\IEEEsetlabelwidth{long a label}\IEEEusemathlabelsep]
 
@@
 
-To test how audio movement source affects localization quality, each of three experiments were conducted at two different speeds:
+To test how sound source movement speed affects localization quality, each of three experiments were conducted at two different speeds:
 \begin{description}
@@
 For each experiment conducted, two different movement filters are tests:
-\begin{description}
+\begin{description}[\IEEEsetlabelwidth{Very long label}\IEEEusemathlabelsep]
 \item[Averaging filter] localization for past $0.5$ seconds are averaged and outputed as current estimate.
@@
   \caption{white noise}
-  \label{fig:circle_wn}
+  \label{fig:circle_wnn}
 \end{subfigure}
@@
   \caption{music A}
-  \label{fig:circle_musica}
+  \label{fig:circle_musican}
 \end{subfigure}
@@
   \caption{music B}
-  \label{fig:circle_musicb}
+  \label{fig:circle_musicbn}
 \end{subfigure}
@@
 \end{figure*}
-
-Fig~\ref{fig:circle_normal} shows results for experiments with three different audio sources at normal speed.
-
-with three different audio source were used for the same circular movement.
-In the first trial, white noise is used to set a baseline for audio sources, since our understanding is that white noise performs the best under GCC\_PHAT.
-In the second trial, a randomly picked music is used.
-In the third trial, a music that contains low amplitude segments is picked to reflect real life situations.
-Fig~\ref{fig:circle} shows the result for all three trials.
-From the result, white noise does perform the best and the localized shape matches well with the shape of the rotating disk.
-There exists a bit of jiggling, but can be improved with filtering.
-Fig~\ref{fig:circle} also shows the result with averaging filtering and Kalman filtering.
-
-For the music A, the array still localizaes correctly and the output represetns a circular shape, but the localizaed points are more spread out with more jiggling.
-This is because music does not have components in all frequency bands and GCC\_PHAT estimate is less accuracy with such audio sources.
-Similar to the white noise case, we can improve the raw tracking estimates from the array with filtering.
-
-For music B, the music has a low amplitude section and it is also shown in the traced result.
-The "blank" region in the output corresponds to the low amplitude segment in the audio.
-The general tracking quality of this music is not as good as that with Music A due to these low amplitude audio segments.
 
@@
   \caption{white noise}
-  \label{fig:circle_wn}
+  \label{fig:circle_wnf}
 \end{subfigure}
@@
   \caption{music A}
-  \label{fig:circle_musica}
+  \label{fig:circle_musicaf}
 \end{subfigure}
@@
   \caption{music B}
-  \label{fig:circle_musicb}
+  \label{fig:circle_musicbf}
 \end{subfigure}
@@
 \end{figure*}
+
+Fig~\ref{fig:circle_normal} shows results for experiments with at normal speed, and Fig~\ref{fig:circle_fast} shows results at fast speed.
+By comparing localization error for each audio source between normal movement speed and fast movement speed, we find the localization error does depend on how fast the sound sorce is moving.
+For example fig~\ref{fig:circle_musican} and fig~\ref{fig:circle_musicaf} shows that localization error is $1.289$ cm at normal movement speed and $1.291$ cm at fast movement speed.
+
+From fig~\ref{fig:circle_normal}, localization error is $0.9$ cm for white noise, $1.29$ cm for Music A, and $2.9$ cm for Music B.
+Localization accuracy is the best for white noise source, and worst for Music B.
+This is consistent with our expectation because low amplitude regions in Music B would make arrays lose track of where the source is.
+This can also be seen from the "blank" regions in fig~\ref{fig:circle_musicbn}
+
+Fig ~\ref{fig:circle_normal} also shows that raw detection has the most amount of jiggling.
+Kalman filter reduces the amount of jiggling from raw detection.
+Averaging filter has the least amount of jiggling.
+However, averaging filter averages detection outputs from past 0.5 seconds, which makes the filtered output lags the real movement.

### 2015-05-12 reports/paper/sound_localization.tex
update paper

\section{Sound Localization}
Constant difference of distance to two fixed points trace out a hyperbola in the 2D plane. Since each pair of microphones gives a hyperbola curve in the plane, localization becomes finding intersection of hyperbolas when more than two microphones are used. Accurate localization relies on accuracte estimate of delay differences between microphones.

Generalized Cross Correlation provides a framework to estimate delay differences $t_0$ between two signals $x_1(t)$ and $x_2(t)$:
\[
t_0 = \arg\max_{\tau} \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
\]
, where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transform of $x_1(t)$ and $x_2(t)$. $W(\omega)$ provides a way to prefilter the signals passed to cross correlation estimator. We experiment with three ways of prefiltering the signal:
\begin{description}[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]
\item[GCC] $W(\omega) = 1$. No prefiltering is done. This is normal cross correlation.
\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divideded by its magnitude. Only phase information contributes to delay estimation
\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. part of magnitude information is included in delay estimation.
\end{description}

### 2015-05-12 reports/paper/system.tex
paper update

\section{System}
The system uses two arrays, each with three microphones. Three microphones along with a micro-controller are mounted at three vertices of an equilateral triangle. Fig~\ref{fig:setup_array} shows a picture of the array. Micro-controller used in this project is \emph{teensy 3.1}. The micro-controller collects microphone data on all three channels for $12$ millisecond and then send the recorded data to a computer for localization.

\begin{figure}[]
  \centering
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{array_close.JPG}
  \end{subfigure}
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{array.JPG}
  \end{subfigure}
  \caption{array}
  \label{fig:setup_array}
\end{figure}

From a timing point of view, the micro controller spends $12$ millisecond on sampling microphone data before sending the data to the computer for processing. Sending data through USB port takes another $18$ millisecond, and processing on computer takes around $50$ millisecond. Therefore, the totoal time lag between sound source and localization is around $80$ millisecond.

### 2015-05-12 reports/paper/sound_localization.tex
paper update

@@
 \section{Sound Localization}
-Constant difference of distance to two fixed points trace out a hyperbola in the 2D plane.
-Since each pair of microphones gives a hyperbola curve in the plane, localization becomes finding intersection of hyperbolas when more than two microphones are used.
-Accurate localization relies on accuracte estimate of delay differences between microphones.
+points with the same difference of distance to two fixed points form a hyperbola in the 2D plane.
+Since each pair of microphones gives a hyperbola in the plane, localization becomes finding intersection of hyperbolas when more than two microphones are used.
+Localization relies on accuracte estimate of delay differences between microphones.
 
-Generalized Cross Correlation provides a framework to estimate delay differences $t_0$ between two signals $x_1(t)$ and $x_2(t)$:
-\[
+Generalized Cross Correlation(GCC) provides a framework to estimate delay differences $t_0$ between two signals $x_1(t)$ and $x_2(t)$:
+\begin{equation} \label{eq:gcc}
 t_0 = \arg\max_{\tau} \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
-\]
+\end{equation}
 , where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transform of $x_1(t)$ and $x_2(t)$.
@@
 \item[GCC] $W(\omega) = 1$. No prefiltering is done. This is normal cross correlation.
-\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divideded by its magnitude. Only phase information contributes to delay estimation
+\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divideded by its magnitude. Only phase information contributes to delay estimation.
 \item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. part of magnitude information is included in delay estimation.

### 2015-05-12 reports/paper/system.tex
paper update

@@
 The system uses two arrays, each with three microphones.
-Three microphones along with a micro-controller are mounted at three vertices of an equilateral triangle.
+Three microphones along with a micro-controller are mounted at three vertices of an $20$cm equilateral triangle.
 Fig~\ref{fig:setup_array} shows a picture of the array.
 Micro-controller used in this project is \emph{teensy 3.1}.
-The micro-controller collects microphone data on all three channels for $12$ millisecond and then send the recorded data to a computer for localization.
+The micro-controller has $64$k RAM memory and the ADC is capable of sampling at $500$kHz.
+In this project, the micro-controller collects microphone data on all three channels for $12$ millisecond and then send the recorded data to a computer for localization.
 
@@
 
+To speed up computation for real time localization, instead of search for $t_0$ that maximizes equation~\ref{eq:gcc}.
+A grid-search in 2D grid is performed instead.
+With this approach, for each point in the grid, the arrival time difference to each microphone pair can be precomputed.
+Then localization resolves to calculating GCC for each microphone pair and performing a lookup for each point in the grid.
+To further improve localization accuracy, instead of using a point estimate for difference of arrival time, a liklihoood map is built.
+Each entry in GCC output is used as the liklihood for that delay.
+With three microphones $m_1,m_2,m_3$, for each location $(x,y)$ in the gird, the difference of arrival time to each microphones pair can be precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$.
+Let $R_{m_1,m_2}(\tau)$,$R_{m_1,m_2}(\tau)$, and $R_{m_1,m_2}(\tau)$ denote the GCC output for each microphone pair, then a liklihood map in the grid is calculated as:
+\begin{eqnarray*}
+  L(x,y) &=& R_{m_1,m_2}(D_{m_1,m_2}(x,y)) + R_{m_1,m_3}(D_{m_1,m_3}(x,y)) \\
+ & & +R_{m_2,m_3}(D_{m_2,m_3}(x,y)) 
+\end{eqnarray*}
+
+Liklihood map from each array can be combined into final liklihood map:
+\begin{equation}\label{eq:combine_l}
+L(x,y) = L_1(x,y) L_2(x,y)
+\end{equation}
+, where $L_1(x,y)$ and $L_2(x,y)$ represents the liklihood map from array $1$ and array $2$
+
+\begin{figure}[]
+  \centering
+  \begin{subfigure}[]{.15\textwidth}
+    \includegraphics[width=\textwidth]{left.png}
+    \caption{localization with only array 1}
+  \end{subfigure}
+  \begin{subfigure}[]{.15\textwidth}
+    \includegraphics[width=\textwidth]{right.png}
+    \caption{localization with only array 2}
+  \end{subfigure}
+  \begin{subfigure}[]{.15\textwidth}
+    \includegraphics[width=\textwidth]{combined.png}
+    \caption{localization with both arrays}
+  \end{subfigure}
+  \caption{Liklihood maps for localizating point at $(0.0,-0.3)$ m}
+  \label{fig:liklihood}
+\end{figure}
+
+To see the effect of using multiple arrays, fig~\ref{fig:liklihood} shows the individual liklihood map from each array and also the combined liklihood according to equation~\ref{eq:combine_l}.
+Individual array gives accurate angle estimate, but has high uncertainty in distance estimate.
+By using two arrays, the angle estimate can be effectively combined to estimate distance.
+
 From a timing point of view, the micro controller spends $12$ millisecond on sampling microphone data before sending the data to the computer for processing.

### 2015-05-12 reports/paper/experiment.tex
paper update

@@
-\section{Experiment}
+\section{Experiment Results}
 To test localization accuracy, an one meter by one meter grid was set up and the arrays are placed at the top left and top right corners of the grid.
@@
 The lowest error achieved is $2.53$ centimeters.
-It is achieved when window size is set to 12 millisecond and GCC\_PHAT is used to estimate difference of delay
+It is achieved when window size is set to 12 millisecond and GCC\_PHAT is used for TDOA estimation.
 
 Although accuracy improves with window length, the calculation time also increases with window length.
-The part of calculation that depends on window length is using cross correlation to estimate difference of arrival time.
+The part of calculation that depends on window length is using cross correlation to estimate TDOA.
 cross correlation can be calculated with FFT and the runtime is of order $O(N\log N)$.
@@
 Fig~\ref{fig:err_win_circle} gives an intuitive representation of how accuracy changes with window size.
-When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate arrival time differences.
+When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate TDOA.
 The localization is noisy.
@@
 
-\item[Music A] A music that has normal audio amplitude throughout experiment period was chosen. "Honest Eyes" by Black Tide was chosen
+\item[Music A] A music that has normal audio amplitude throughout experiment period was chosen. "Honest Eyes" by Black Tide was the music used.
 
-\item[Music B] A music with intermittent low amplitude sections was chosen. "Canon" was used.
+\item[Music B] A music with intermittent low amplitude sections was chosen. "Canon" was the music used.
 
@@
 Fig~\ref{fig:circle_normal} shows results for experiments with at normal speed, and Fig~\ref{fig:circle_fast} shows results at fast speed.
-By comparing localization error for each audio source between normal movement speed and fast movement speed, we find the localization error does depend on how fast the sound sorce is moving.
+By comparing localization error for each audio source between normal movement speed and fast movement speed, we find the localization error does not depend on how fast the sound sorce is moving.
 For example fig~\ref{fig:circle_musican} and fig~\ref{fig:circle_musicaf} shows that localization error is $1.289$ cm at normal movement speed and $1.291$ cm at fast movement speed.
@@
 This is consistent with our expectation because low amplitude regions in Music B would make arrays lose track of where the source is.
-This can also be seen from the "blank" regions in fig~\ref{fig:circle_musicbn}
+This can also be seen from the "blank" regions in fig~\ref{fig:circle_musicbn}.
 
@@
 Averaging filter has the least amount of jiggling.
-However, averaging filter averages detection outputs from past 0.5 seconds, which makes the filtered output lags the real movement.
+However, averaging filter averages detection outputs from past 0.5 seconds, which makes the filtered output lag the real movement.

### 2015-05-12 reports/paper/system.tex
paper update

--- 2015-05-12 reports/paper/system.tex
@@
 \section{System}
-The system uses two arrays, each with three microphones.
-Three microphones along with a micro-controller are mounted at three vertices of an equilateral triangle.
-Fig~\ref{fig:setup_array} shows a picture of the array.
+The system uses two arrays, each with three microphones mounted on vetices of a $20$cm equilateral triangle.
+A micro-controller is also attached to one of the vertices.
+Fig~\ref{fig:setup_array} shows a picture of the array setup.
 Micro-controller used in this project is \emph{teensy 3.1}.
-The micro-controller collects microphone data on all three channels for $12$ millisecond and then send the recorded data to a computer for localization.
+It has $64$k RAM memory and the ADC is capable of sampling at $500$kHz.
+In this project, the micro-controller collects microphone data on all three channels for $12$ millisecond and then send the recorded data to a computer for localization.
 
@@
   \centering
-  \begin{subfigure}[]{.2\textwidth}
+  \begin{subfigure}[]{.15\textwidth}
     \includegraphics[width=\textwidth]{array_close.JPG}
+    \caption{micro-controller}
   \end{subfigure}
-  \begin{subfigure}[]{.2\textwidth}
+  \begin{subfigure}[]{.15\textwidth}
     \includegraphics[width=\textwidth]{array.JPG}
+    \caption{array}
   \end{subfigure}
-  \caption{array}
+  \begin{subfigure}[]{.15\textwidth}
+    \includegraphics[width=\textwidth]{setup_2.JPG}
+    \caption{two arrays}
+  \end{subfigure}
+  \caption{Localization arrays}
   \label{fig:setup_array}
@@
 
-From a timing point of view, the micro controller spends $12$ millisecond on sampling microphone data before sending the data to the computer for processing.
+To speed up computation for real time localization, instead of searching for $t_0$ that maximizes equation~\ref{eq:gcc}, a grid search in 2D grid is performed.
+For each point in the grid, the TDOA to each microphone pair can be precomputed.
+Then localization resolves to calculating GCC for each microphone pair and performing a lookup for each point in the grid.
+To further improve localization accuracy, instead of using point estimate for TDOA, a liklihoood map is built.
+Each entry in GCC output is used as the liklihood for that delay.
+With three microphones $m_1,m_2,m_3$, there are three microphones pairs: $m_1m_2,m_1m_3,m_2m_3$.
+TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$.
+Then the liklihood map can be built as:
+\begin{eqnarray*}
+  L(x,y) &=& R_{m_1,m_2}(D_{m_1,m_2}(x,y)) + R_{m_1,m_3}(D_{m_1,m_3}(x,y)) \\
+ & & +R_{m_2,m_3}(D_{m_2,m_3}(x,y)) 
+\end{eqnarray*}
+,where $R_{m_1,m_2}(\tau)$,$R_{m_1,m_2}(\tau)$, and $R_{m_1,m_2}(\tau)$ denote GCC output from microphone pairs $m_1m_2,m_1m_3,$ and $m_2m_3$.
+
+Liklihood maps from two arrays can be combined into final liklihood map:
+\begin{equation}\label{eq:combine_l}
+L(x,y) = L_1(x,y) L_2(x,y)
+\end{equation}
+, where $L_1(x,y)$ and $L_2(x,y)$ represents the liklihood map from array $1$ and array $2$.
+
+\begin{figure}[]
+  \centering
+  \begin{subfigure}[]{.15\textwidth}
+    \includegraphics[width=\textwidth]{left.png}
+    \caption{localization with only array 1}
+    \label{fig:liklihood1}
+  \end{subfigure}
+  \begin{subfigure}[]{.15\textwidth}
+    \includegraphics[width=\textwidth]{right.png}
+    \caption{localization with only array 2}
+    \label{fig:liklihood2}
+  \end{subfigure}
+  \begin{subfigure}[]{.15\textwidth}
+    \includegraphics[width=\textwidth]{combined.png}
+    \caption{localization with both arrays}
+    \label{fig:liklihood3}
+  \end{subfigure}
+  \caption{Liklihood maps for localizating point at $(0.0,-0.3)$ m}
+  \label{fig:liklihood}
+\end{figure}
+
+To see the effect of using multiple arrays, fig~\ref{fig:liklihood} shows the individual liklihood map from each array and also the combined liklihood map according to equation~\ref{eq:combine_l}.
+Individual array gives accurate angle estimate, but has high uncertainty in distance estimate.
+By combining estimates from two arrays, the angle estimate can be effectively combined to estimate distance.
+
+From a timing point of view, the micro-controller spends $12$ millisecond on sampling microphone data before sending the data to the computer for processing.
 Sending data through USB port takes another $18$ millisecond, and processing on computer takes around $50$ millisecond.

### 2015-05-13 reports/paper/simulation.tex
paper update added simulation section

\section{Simulation}
In a 2D region, points with the same TDOA to two fixed locations form a hyperbola. However, in practical systems, we can only measure TDOA up to a precision.Therefore we look at all points with difference of distance close to some target value within measurement error $\epsilon$. This $\epsilon$ prepresents accuracy on measurement of difference of distances, and in practice it is related to sampling rate and TDOA methods.

\begin{figure}[]
  \centering
  \begin{subfigure}[]{.23\textwidth}
    \includegraphics[width=\textwidth]{sim/sim_2_1}
    \caption{micro-controller}
  \end{subfigure}
  \begin{subfigure}[]{.23\textwidth}
    \includegraphics[width=\textwidth]{sim/sim_2_2}
    \caption{array}
  \end{subfigure}
  \begin{subfigure}[]{.23\textwidth}
    \includegraphics[width=\textwidth]{sim/sim_2_3}
    \caption{two arrays}
  \end{subfigure}
  \caption{Localization arrays}
  \label{fig:sim_2_5}
\end{figure}

To see how precision affects localization accuracy, we simulated two microphones placed at: $M_1:(x=-10\mbox{cm},y=0\mbox{cm})$ and $M_2:(x=10\mbox{cm},y=0\mbox{cm})$. A test sound source is emitted at point $P$ $50$ centimeters away from $(0,0)$. Fig~\ref{fig:sim_2_5} shows the region where all points $\hat P$ satisfy:
\[
 |(\hat P M_1 - \hat P M_2) - (P M_1 - P M_2)| < 1 \mbox{cm}
\]
Intuitively, points in the region have difference of distance very similar to each other. From fig~\ref{fig:sim_2_5}, the region still has the shape of a hyperbola, but with an uncertainty region around the curvce. The uncertainty region is not uniform around the curve, the farther away the point is, the larger the uncertainty region becomes. It indicates that the samve delta distance movement will generate smaller difference of distance when you are farther away from the array. The size of the uncertianty region is also angle dependent: points closer to the line of microphones have larger region compared to points close to the line perpenticular to microphones.

\begin{figure*}[]
  \centering
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_20cm_0_degrees}
    \caption{0 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_20cm_45_degrees}
    \caption{45 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_20cm_90_degrees}
    \caption{90 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_20cm_135_degrees}
    \caption{135 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_20cm_180_degrees}
    \caption{180 degrees}
  \end{subfigure}
  \caption{$20$cm equilateral triangle array. Source is $20$cm away from the array}
  \label{fig:sim_3_2}
\end{figure*}

\begin{figure*}[]
  \centering
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_80cm_0_degrees}
    \caption{0 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_80cm_45_degrees}
    \caption{45 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_80cm_90_degrees}
    \caption{90 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_80cm_135_degrees}
    \caption{135 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_80cm_180_degrees}
    \caption{180 degrees}
  \end{subfigure}
  \caption{$20$cm equilateral triangle array. Source is $80$cm away from the array}
  \label{fig:sim_3_8}
\end{figure*}

With more than two multiple microphones, each pair of microphones generates a hyperbolic region and localization becomes finding the intersection of regions. The smaller the intersection region, the better the localization accuracy. To see how accuracy changes with array placement and sound source location, three microphones are placed at three vertices of an $20$cm equilateral triangle. An audio source is placed at $20$ cm away from the center of the array. Fig~\ref{fig:sim_3_2} shows the intersection of regions for $5$ different placement of the sound source. It can be seen that accuracy is worse when sound source is close to the line of any two microphones. This observation is consistent with two microphone case, since points close to line of microphones have a larger uncertainty region.

To see how sound source distance affects localization accuracy, the same simulation is carried out with sound source moved from $20$cm to $80$cm away from the center of the array. Results are presetned in fig~\ref{fig:sim_3_8}. Comparing with fig~\ref{fig:sim_3_2}, accuracy decreases with distance to the array. This is also consistent with our observation in $2$ microphone case where source farther away would result in larger uncertainty region.

\begin{figure*}[]
  \centering
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_center}
    \caption{$3$ microphones placed at vertices of an $20$cm equilateral triangle}
    \label{fig:sim_hm_3}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_p1}
    \caption{another microphone added at origin}
    \label{fig:sim_hm_3_p1}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_2x}
    \caption{$3$ microphones placed at vertices of an $40$cm equilateral triangle}
    \label{fig:sim_hm_3_2x}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_line}
    \caption{$3$ microphones placed in a line. Total length is $20$cm}
    \label{fig:sim_hm_3_line}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_4corner}
    \caption{$4$ microphones placed at $4$ corners of the region}
    \label{fig:sim_hm_4}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_center}
    \caption{two $3$ microphone arrays placed $1$ meter apart}
    \label{fig:sim_hm_2_array}
  \end{subfigure}
  \caption{Accuracy for different array configurations}
  \label{fig:sim_hm}
\end{figure*}

Intersection area is a measure of the localization accuracy. To evaluate an array's accuracy in a region, we can place sound source at predetermined grid points in the region and look at the intersection area for each point in the grid. Results for a few different configurations are presented in fig~\ref{fig:sim_hm}.

Fig~\ref{fig:sim_hm_3} shows the accuracy when microphones are placed at three vertices of an $20$ cm equilateral triangle. The region inside the array has good accuracy. However, for regions along line of any two microphones, the accuracy drops significantly.

To evaluate how adding one microphone(without increasing array size) improves accuracy,  another microphone is added to the array at $(0,0)$. Result is presented in fig~\ref{fig:sim_hm_3_p1}. Addition of the new microphone only slightly improved the accuracy around the array region. Regions near lines of microphones still have significantly larger uncertainty region.

To evaluate array size's impact on accuracy, the size of original array from fig~\ref{fig:sim_hm_3} is increased by a factor of $2$. The result is presented in fig~\ref{fig:sim_hm_3_2x}. The overall uncertainty area decreased across the region.

In fig~\ref{fig:sim_hm_3_line}, three microphones $10$cm apart from each other are placed in a line on x-axis. The result indicates that this configuration has high uncertainty in nearly all regions except for regions near the microphones. In fig~\ref{fig:sim_hm_4}, four microphones are placed at four corners of the region. Accuracy is consistently good across the region. However, this configuration requires microphone to be placed far apart.

To get benefit of far away microphone placement without incurring overhead of placing individual microphones, two array solutions are explored. Fig~\ref{fig:sim_hm_2_array} shows the accuracy map when two $3$ microphone array are placed $1$ meter apart.  It shows that the array has good accuracy inside a one meter by one meter region

### 2015-05-13 reports/paper/system.tex
paper update added simulation section

@@
 It has $64$k RAM memory and the ADC is capable of sampling at $500$kHz.
-In this project, the micro-controller collects microphone data on all three channels for $12$ millisecond and then send the recorded data to a computer for localization.
+In this project, the micro-controller collects microphone data on all three channels for $12$ millisecond and then send the recorded data to a computer through USB port for localization.
 
@@
 To speed up computation for real time localization, instead of searching for $t_0$ that maximizes equation~\ref{eq:gcc}, a grid search in 2D grid is performed.
-For each point in the grid, the TDOA to each microphone pair can be precomputed.
+For each point in the grid, the theoretical TDOA to each microphone pair can be precomputed.
 Then localization resolves to calculating GCC for each microphone pair and performing a lookup for each point in the grid.
@@
 With three microphones $m_1,m_2,m_3$, there are three microphones pairs: $m_1m_2,m_1m_3,m_2m_3$.
-TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$.
+Theoretical TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$.
 Then the liklihood map can be built as:

### 2015-05-13 reports/paper/simulation.tex
paper update. simulation results

@@
 In a 2D region, points with the same TDOA to two fixed locations form a hyperbola.
-However, in practical systems, we can only measure TDOA up to a precision.Therefore we look at all points with difference of distance close to some target value within measurement error $\epsilon$.
+However, in practical systems, we can only measure TDOA up to a precision.
+Therefore we look at all points with difference of distance close to some target value within measurement error $\epsilon$.
 This $\epsilon$ prepresents accuracy on measurement of difference of distances, and in practice it is related to sampling rate and TDOA methods.
@@
     \includegraphics[width=\textwidth]{sim/sim_2_1}
-    \caption{micro-controller}
+    \caption{source at $(r=50$ cm,$\theta = 0$ degrees$)$}
   \end{subfigure}
@@
     \includegraphics[width=\textwidth]{sim/sim_2_2}
-    \caption{array}
+    \caption{source at $(r=50$ cm,$\theta = 45$ degrees$)$}
   \end{subfigure}
@@
     \includegraphics[width=\textwidth]{sim/sim_2_3}
-    \caption{two arrays}
+    \caption{source at $(r=50$ cm,$\theta = 90$ degrees$)$}
   \end{subfigure}
-  \caption{Localization arrays}
+  \caption{Uncertainty region}
   \label{fig:sim_2_5}
@@
 
-To see how precision affects localization accuracy, we simulated two microphones placed at: $M_1:(x=-10\mbox{cm},y=0\mbox{cm})$ and $M_2:(x=10\mbox{cm},y=0\mbox{cm})$.
+To see how precision affects localization accuracy, we simulated two microphones placed at: $M_1:(x=-10\mbox{ cm},y=0\mbox{ cm})$ and $M_2:(x=10\mbox{ cm},y=0\mbox{ cm})$.
 A test sound source is emitted at point $P$ $50$ centimeters away from $(0,0)$.
@@
 \[
- |(\hat P M_1 - \hat P M_2) - (P M_1 - P M_2)| < 1 \mbox{cm}
+ (\hat P M_1 - \hat P M_2) - (P M_1 - P M_2) < 1 \mbox{cm}
 \]
@@
 The uncertainty region is not uniform around the curve, the farther away the point is, the larger the uncertainty region becomes.
-It indicates that the samve delta distance movement will generate smaller difference of distance when you are farther away from the array.
+It indicates that the same delta distance movement will generate smaller difference of distance when the source is farther away from the array.
 The size of the uncertianty region is also angle dependent: points closer to the line of microphones have larger region compared to points close to the line perpenticular to microphones.
@@
   \end{subfigure}
-  \caption{$20$cm equilateral triangle array. Source is $80$cm away from the array}
+  \caption{$20$ cm equilateral triangle array. Source is $80$ cm away from the array}
   \label{fig:sim_3_8}
@@
 
-With more than two multiple microphones, each pair of microphones generates a hyperbolic region and localization becomes finding the intersection of regions.
+With more than two multiple microphones, each pair of microphones generates a hyperbolic region and localization becomes finding the intersection of hyperbolic regions.
 The smaller the intersection region, the better the localization accuracy.
-To see how accuracy changes with array placement and sound source location, three microphones are placed at three vertices of an $20$cm equilateral triangle.
+To see how accuracy changes with array placement and sound source location, three microphones are placed at three vertices of an $20$ cm equilateral triangle.
 An audio source is placed at $20$ cm away from the center of the array.
@@
 
-To see how sound source distance affects localization accuracy, the same simulation is carried out with sound source moved from $20$cm to $80$cm away from the center of the array.
+To see how sound source distance affects localization accuracy, the same simulation is carried out with sound source moved from $20$ cm to $80$ cm away from the center of the array.
 Results are presetned in fig~\ref{fig:sim_3_8}.
@@
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_center}
-    \caption{$3$ microphones placed at vertices of an $20$cm equilateral triangle}
+    \caption{$3$ microphones placed at vertices of an $20$cm equilateral triangle. Average error is $18.6$ cm}
     \label{fig:sim_hm_3}
@@
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_p1}
-    \caption{another microphone added at origin}
+    \caption{another microphone added at origin. Average error is $17.1$ cm}
     \label{fig:sim_hm_3_p1}
@@
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_2x}
-    \caption{$3$ microphones placed at vertices of an $40$cm equilateral triangle}
+    \caption{$3$ microphones placed at vertices of an $40$cm equilateral triangle. Average error is $10.04$ cm}
     \label{fig:sim_hm_3_2x}
@@
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_line}
-    \caption{$3$ microphones placed in a line. Total length is $20$cm}
+    \caption{$3$ microphones placed in a line. Total length is $20$ cm. Average error is $55.05$ cm}
     \label{fig:sim_hm_3_line}
@@
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_4corner}
-    \caption{$4$ microphones placed at $4$ corners of the region}
+    \caption{$4$ microphones placed at $4$ corners of the region. Average error is $0.05$ cm}
     \label{fig:sim_hm_4}
@@
   \begin{subfigure}[]{.3\textwidth}
-    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_center}
-    \caption{two $3$ microphone arrays placed $1$ meter apart}
+    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_2a}
+    \caption{two $3$ microphone arrays placed $1$ meter apart. Average error is $2.60$ cm}
     \label{fig:sim_hm_2_array}
@@
 Intersection area is a measure of the localization accuracy.
-To evaluate an array's accuracy in a region, we can place sound source at predetermined grid points in the region and look at the intersection area for each point in the grid.
-Results for a few different configurations are presented in fig~\ref{fig:sim_hm}.
+To evaluate an array's accuracy in a region, we can place sound source at predetermined grid points in the region and look at the intersection area for each tested point in the grid.
+The center location of intersection region can be used as localization estimate to calculate localization error.
+Results for a few different microphone array configurations are presented in fig~\ref{fig:sim_hm}.
 
-Fig~\ref{fig:sim_hm_3} shows the accuracy when microphones are placed at three vertices of an $20$ cm equilateral triangle.
+Fig~\ref{fig:sim_hm_3} shows the accuracy when microphones are placed at three vertices of a $20$ cm equilateral triangle.
 The region inside the array has good accuracy.
 However, for regions along line of any two microphones, the accuracy drops significantly.
+Average error across the region is $18.6$ cm.
 
@@
 Addition of the new microphone only slightly improved the accuracy around the array region.
+Average error dropped from $18.6$ cm to $17.1$ cm.
 Regions near lines of microphones still have significantly larger uncertainty region.
@@
 The overall uncertainty area decreased across the region.
+Average error improved to $10.04$ cm.
 
-In fig~\ref{fig:sim_hm_3_line}, three microphones $10$cm apart from each other are placed in a line on x-axis.
-The result indicates that this configuration has high uncertainty in nearly all regions except for regions near the microphones.
-In fig~\ref{fig:sim_hm_4}, four microphones are placed at four corners of the region.
-Accuracy is consistently good across the region.
-However, this configuration requires microphone to be placed far apart.
+In fig~\ref{fig:sim_hm_3_line}, three microphones are placed $10$ cm apart from each other on x-axis.
+Error heatmap showed high uncertainty on the x axis, and the overall accuracy is not as good as that with three microphones placed in a triangle.
+The average error is $55.05$ cm.
 
-To get benefit of far away microphone placement without incurring overhead of placing individual microphones, two array solutions are explored.
-Fig~\ref{fig:sim_hm_2_array} shows the accuracy map when two $3$ microphone array are placed $1$ meter apart.
-It shows that the array has good accuracy inside a one meter by one meter region
+To further increase the distance between microphones, we placed four microphones at four corners of the region.
+Fig~\ref{fig:sim_hm_4} showed the result.
+With this configuration, accuracy is consistently good across the region.
+The average error is $0.05$ cm.
+However, placing microphones far apart at corners of the region requires accurate placement of all four individual microphones.
+The system is less portable compared to small arrays with microphones near each other.
+
+To avoid the need to accurate place four microphones at far distance(as required by fig~\ref{fig:sim_hm_4}), we explored configuration with two arrays.
+Two $3$ microphone array are placed $1$ meter apart and the result is presented in fig~\ref{fig:sim_hm_2_array}.
+The result indicates that this configuration has good accuracy when source is close to the arrays.
+Accuracy decreases as sound source moves outside the one meter by one meter region.
+The average error is $2.60$ cm.

### 2015-05-18 reports/paper/prior.tex
rename

--- 2015-05-12 reports/paper/sound_localization.tex
@@
 \section{Sound Localization}
-points with the same difference of distance to two fixed points form a hyperbola in the 2D plane.
-Since each pair of microphones gives a hyperbola in the plane, localization becomes finding intersection of hyperbolas when more than two microphones are used.
-Localization relies on accuracte estimate of delay differences between microphones.
+Points with the same time difference of arrival(TDOA) to two fixed points on a plane form a hyperbola.
+Since TDOA from each pair of microphones gives a hyperbola in the plane, localization becomes finding intersection of hyperbolas when more than two microphones are used.
+Localization relies on accurate estimate of delay differences between microphones.
 
 Generalized Cross Correlation(GCC) provides a framework to estimate delay differences $t_0$ between two signals $x_1(t)$ and $x_2(t)$:
-\begin{equation} \label{eq:gcc}
-t_0 = \arg\max_{\tau} \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
-\end{equation}
+\begin{eqnarray} \label{eq:gcc}
+t_0 &=& \arg\max_{\tau} R_{x_1x_2}(\tau) \\
+R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
+\end{eqnarray}
 , where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transform of $x_1(t)$ and $x_2(t)$.
-$W(\omega)$ provides a way to prefilter the signals passed to cross correlation estimator.
-We experiment with three ways of prefiltering the signal:
+$W(\omega)$ provides a way to prefilter signals passed to cross correlation estimator.
+We experimented with three ways of prefiltering the signal:
 \begin{description}[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]

### 2015-05-19 reports/paper/introduction.tex
add
working on prior section

--- 2015-05-12 reports/paper/sound_localization.tex
@@
-\section{Sound Localization}
-Constant difference of distance to two fixed points trace out a hyperbola in the 2D plane.
-Since each pair of microphones gives a hyperbola curve in the plane, localization becomes finding intersection of hyperbolas when more than two microphones are used.
-Accurate localization relies on accuracte estimate of delay differences between microphones.
-
-Generalized Cross Correlation provides a framework to estimate delay differences $t_0$ between two signals $x_1(t)$ and $x_2(t)$:
-\[
-t_0 = \arg\max_{\tau} \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
-\]
-, where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transform of $x_1(t)$ and $x_2(t)$.
-$W(\omega)$ provides a way to prefilter the signals passed to cross correlation estimator.
-We experiment with three ways of prefiltering the signal:
-\begin{description}[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]
-\item[GCC] $W(\omega) = 1$. No prefiltering is done. This is normal cross correlation.
-\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divideded by its magnitude. Only phase information contributes to delay estimation
-\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. part of magnitude information is included in delay estimation.
-\end{description}
+\section{Introduction}
+ Inexpensive local localization. Options:
+\begin{enumerate}
+\item GPS not accurate.
+\item Ultrasound. expensive
+\item use sound. Everyone has access to. One application is to draw with music
+\end{enumerate}

### 2015-05-19 reports/paper/prior.tex
working on prior section

@@
-\section{Sound Localization}
-Points with the same time difference of arrival(TDOA) to two fixed points on a plane form a hyperbola.
-Since TDOA from each pair of microphones gives a hyperbola in the plane, localization becomes finding intersection of hyperbolas when more than two microphones are used.
+\section{Background}
+Acoustic localization has been researched extensively in the literature. Localization techniques can be broadly put into two categories:
+\begin{enumerate}
+\item Location Template Matching (LTM) based approaches.
+\item Time Difference of Arrival (TDOA) based approaches.
+\end{enumerate}
+
+\subsection{LTM}
+In LTM based approaches, acoustic templates of locations are first stored in the system during a ``training'' phase.
+Then, incoming acoustic waveform is compared with stored templates.
+Localization result is the location that has the best matching template.
+Different ways of extracting templates from raw acoustic source and different similarity measures have been investigated in the past.
+
+\cite{extended:tusi} and \cite{ltm:pham} investigated using max value from corss-correlation as similarity measure to localize user tap on interactive surface.
+\cite{ltm:lpc} used L2 distance in Linear Predictive Coding coeeficient space as similarity measure to localize taps on surfaces.
+\cite{ltm:tusi2} further explored accuracy improvement by using multiple templates for each location and speed improvement by merging multiple templates into one representative templates.
+
+The requirement of having a template for each location to be detected makes this approach too restrictive for our project, since we want the drawing to be continuous in the 2D region.
+Moreover, the need to recalibrate all locations during setup is too curbursome for end users in a portable drawing application.
+Therefore, our main focus is on TDOA approaches.
+
+\subsection{TDOA}
+Points with the same TDOA to two fixed points on a plane form a hyperbola.
+With more than two microphones, intersection of hyperbola curves generated by each pair of microphoens give the source location.
 Localization relies on accurate estimate of delay differences between microphones.
+
+In \cite{tdoa:ppp}, authors used eight microphnes mounted on corners of a ping pong table to localize the point where the ping pong ball hits the table.
+They used a threshold to determine the arrival time of acoustic signal.
+This approach works well in noise free environment but performance degrades with background noise.
+It also suffers from dispersive deflections that arrives before the main wavefront.
+To make it more robust, authors in \cite{tdoa:mit3} extracted descriptive parameters for each significant peak(e.g., peak height, width, mean arrival time).
+The algorithm then used extracted parameters to predict arrival time with a second order polynomial, the parameters of which were fitted during calibration at fixed locations.
+
+In~\cite{tdoa:mit2, tdoa:3}, authors explored using cross-correlation to measure the TDOA between signals.
 

### 2015-05-19 reports/paper/prior.tex
paper prior update

@@
 It also suffers from dispersive deflections that arrives before the main wavefront.
-To make it more robust, authors in \cite{tdoa:mit3} extracted descriptive parameters for each significant peak(e.g., peak height, width, mean arrival time).
+To make it more robust, authors in \cite{tdoa:mit3} and \cite{tdoa:mit4} extracted descriptive parameters for each significant peak(e.g., peak height, width, mean arrival time).
 The algorithm then used extracted parameters to predict arrival time with a second order polynomial, the parameters of which were fitted during calibration at fixed locations.
 
-In~\cite{tdoa:mit2, tdoa:3}, authors explored using cross-correlation to measure the TDOA between signals.
+Cross-correlation has been used extensively in measure signal delays\cite{tdoa:mit2, tdoa:3}.
+Cross-correlation with different prefiterings (known as \emph{generalized cross correlation (GCC)}) have also been investigated to improve delay estimation~\cite{tdoa:gcc1,tdoa:gcc2,tdoa:gcc3}.
 
-Generalized Cross Correlation(GCC) provides a framework to estimate delay differences $t_0$ between two signals $x_1(t)$ and $x_2(t)$:
+GCC provides a framework to estimate delay differences $t_0$ between two signals $x_1(t)$ and $x_2(t)$:
 \begin{eqnarray} \label{eq:gcc}

### 2015-05-19 reports/paper/experiment.tex
paper update added simulation section
paper update

@@
-\section{Experiment Results}
-To test localization accuracy, an one meter by one meter grid was set up and the arrays are placed at the top left and top right corners of the grid.
-Fig~\ref{fig:setup_point} shows a picture of the setup.
+\section{Experiment}
+\subsection{Setup}
+\subsubsection{Point localization}
 
@@
 
+To test localization accuracy, an one meter by one meter grid was set up and the arrays are placed at the top left and top right corners of the grid.
+Fig~\ref{fig:setup_point} shows a picture of the setup.
 A total of $32$ positions are chosen uniformly in this region where microphone data is recorded.
+
+\subsubsection{Movement tracking}
+To test how well the arrays track movement, we mounted a rotating disk $40$ centimeter in diameter onto the grid at $(x=0, y=-0.3)$.
+Fig~\ref{fig:setup_circle} shows a picture of the sestup.
+A sound source is placed on the edge of rotating disk and the arrays localizes the sound source as it rotates in a circle.
+In this experiment, we tested how accuracy changes with:
+\begin{itemize}
+\item different window sizes
+\item different audio sources
+\item different movement tracking filters
+\item different movement speeds
+\end{itemize}
+
+To test how different sound sources impact localization quality, we conducted three experiments on the same movement track with three different sound sources:
+
+\begin{description}[\IEEEsetlabelwidth{long a label}\IEEEusemathlabelsep]
+
+\item[White Noise] A recording of white noise.
+
+\item[Music A] A music that has normal audio amplitude throughout experiment period was chosen. "Honest Eyes" by Black Tide was the music used.
+
+\item[Music B] A music with intermittent low amplitude sections was chosen. "Canon" was the music used.
+
+\end{description}
+
+To test how sound source movement speed affects localization quality, each of three experiments were conducted at two different speeds:
+\begin{description}
+\item[Normal] An angular speed of $0.5$ rad/s was maintained, which translates to linear speed of $10$ cm/s.
+\item[Fast] An angular speed of $1.0$ rad/s was maintained, which translates to linear speed of $20$ cm/s.
+\end{description}
+
+For each experiment conducted, two different movement filters are tests:
+\begin{description}[\IEEEsetlabelwidth{Very long label}\IEEEusemathlabelsep]
+\item[Averaging filter] localization for past $0.5$ seconds are averaged and outputed as current estimate.
+\item[Kalman filter] A 2nd order Kalman filtering is used.
+\end{description}
+
+\subsection{Results}
+\subsubsection{Point localization}
+
 To test how accuracy varies with window size, the algorithm is fed with recorded microphone data with different segment length.
@@
 
-To test how well the arrays track movement, we mounted a rotating disk $40$ centimeter in diameter onto the grid at $(x=0, y=-0.3)$.
-Fig~\ref{fig:setup_circle} shows a picture of the sestup.
-A sound source is placed on the edge of rotating disk and the arrays localizes the sound source as it rotates in a circle.
-In this experiment, we tested how accuracy changes with:
-\begin{itemize}
-\item different window sizes
-\item different audio sources
-\item different movement tracking filters
-\item different movement speeds
-\end{itemize}
-
 \begin{figure*}[]
@@
 
+\subsubsection{Movement tracking}
+
 Fig~\ref{fig:err_win_circle} gives an intuitive representation of how accuracy changes with window size.
@@
 The error decreases as window size increases and plateaus after window size exceeds around $10$ milliseconds.
-
-To test how different sound sources impact localization quality, we conducted three experiments on the same movement track with three different sound sources:
-
-\begin{description}[\IEEEsetlabelwidth{long a label}\IEEEusemathlabelsep]
-
-\item[White Noise] A recording of white noise.
-
-\item[Music A] A music that has normal audio amplitude throughout experiment period was chosen. "Honest Eyes" by Black Tide was the music used.
-
-\item[Music B] A music with intermittent low amplitude sections was chosen. "Canon" was the music used.
-
-\end{description}
-
-To test how sound source movement speed affects localization quality, each of three experiments were conducted at two different speeds:
-\begin{description}
-\item[Normal] An angular speed of $0.5$ rad/s was maintained, which translates to linear speed of $10$ cm/s.
-\item[Fast] An angular speed of $1.0$ rad/s was maintained, which translates to linear speed of $20$ cm/s.
-\end{description}
-
-For each experiment conducted, two different movement filters are tests:
-\begin{description}[\IEEEsetlabelwidth{Very long label}\IEEEusemathlabelsep]
-\item[Averaging filter] localization for past $0.5$ seconds are averaged and outputed as current estimate.
-\item[Kalman filter] A 2nd order Kalman filtering is used.
-\end{description}
 
@@
 Localization accuracy is the best for white noise source, and worst for Music B.
-This is consistent with our expectation because low amplitude regions in Music B would make arrays lose track of where the source is.
+This is consistent with our expectation because low amplitude regions in Music B would cause arrays to lose track of where the source is.
 This can also be seen from the "blank" regions in fig~\ref{fig:circle_musicbn}.

### 2015-05-19 reports/paper/prior.tex
paper update

@@
 
-Cross-correlation has been used extensively in measure signal delays\cite{tdoa:mit2, tdoa:3}.
+Cross-correlation has been used extensively in measuring signal delays\cite{tdoa:mit2, tdoa:3}.
 Cross-correlation with different prefiterings (known as \emph{generalized cross correlation (GCC)}) have also been investigated to improve delay estimation~\cite{tdoa:gcc1,tdoa:gcc2,tdoa:gcc3}.
@@
 $W(\omega)$ provides a way to prefilter signals passed to cross correlation estimator.
-We experimented with three ways of prefiltering the signal:
+We focused with three ways of prefiltering the signal:
 \begin{description}[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]

### 2015-05-19 reports/paper/simulation.tex
rename
paper update

@@
-\section{Simulation}
-In a 2D region, points with the same TDOA to two fixed locations form a hyperbola.
-However, in practical systems, we can only measure TDOA up to a precision.
+\section{Array Architecture}
+As was mentioned in the previous section, points with the same TDOA to two fixed locations form a hyperbola in a 2D plane.
+In practical systems, we can only measure TDOA up to a precision.
 Therefore we look at all points with difference of distance close to some target value within measurement error $\epsilon$.
 This $\epsilon$ prepresents accuracy on measurement of difference of distances, and in practice it is related to sampling rate and TDOA methods.
+In this section we evaluate delay estimation's impact on localization accuracy.
 
@@
 To see how precision affects localization accuracy, we simulated two microphones placed at: $M_1:(x=-10\mbox{ cm},y=0\mbox{ cm})$ and $M_2:(x=10\mbox{ cm},y=0\mbox{ cm})$.
-A test sound source is emitted at point $P$ $50$ centimeters away from $(0,0)$.
+A test sound source is emitted at point $P$ which is $50$ centimeters away from $(0,0)$.
 Fig~\ref{fig:sim_2_5} shows the region where all points $\hat P$ satisfy:
 \[
- (\hat P M_1 - \hat P M_2) - (P M_1 - P M_2) < 1 \mbox{cm}
+ (\hat P M_1 - \hat P M_2) - (P M_1 - P M_2) < 1 \mbox{ cm}
 \]
@@
 The average error is $2.60$ cm.
+
+With simulation results, we decided to build the two array system as described in fig~\ref{fig:sim_hm_2_array}.
+The setup is reasonably portable (compared to fig~\ref{fig:sim_hm_4}), while at the same time having significantly better accuracy compared to one array systems.

### 2015-05-19 reports/paper/system.tex
rename
paper update

@@
-\section{System}
-The system uses two arrays, each with three microphones mounted on vetices of a $20$cm equilateral triangle.
+\section{System Setup}
+The end system has two arrays, each with three microphones mounted on vetices of a $20$ cm equilateral triangle.
 A micro-controller is also attached to one of the vertices.
@@
   \centering
-  \begin{subfigure}[]{.15\textwidth}
+  \begin{subfigure}[]{.23\textwidth}
     \includegraphics[width=\textwidth]{array_close.JPG}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.15\textwidth}
+  \begin{subfigure}[]{.23\textwidth}
     \includegraphics[width=\textwidth]{array.JPG}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.15\textwidth}
+  \begin{subfigure}[]{.23\textwidth}
     \includegraphics[width=\textwidth]{setup_2.JPG}
@@
 \end{eqnarray*}
-,where $R_{m_1,m_2}(\tau)$,$R_{m_1,m_2}(\tau)$, and $R_{m_1,m_2}(\tau)$ denote GCC output from microphone pairs $m_1m_2,m_1m_3,$ and $m_2m_3$.
+,where $R_{m_1,m_2}(\tau)$,$R_{m_1,m_3}(\tau)$, and $R_{m_2,m_3}(\tau)$ denote GCC output from microphone pairs $m_1m_2,m_1m_3,$ and $m_2m_3$.
 
@@
   \centering
-  \begin{subfigure}[]{.15\textwidth}
+  \begin{subfigure}[]{.23\textwidth}
     \includegraphics[width=\textwidth]{left.png}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.15\textwidth}
+  \begin{subfigure}[]{.23\textwidth}
     \includegraphics[width=\textwidth]{right.png}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.15\textwidth}
+  \begin{subfigure}[]{.23\textwidth}
     \includegraphics[width=\textwidth]{combined.png}

### 2015-05-19 reports/paper/introduction.tex
paper update

@@
 \section{Introduction}
- Inexpensive local localization. Options:
-\begin{enumerate}
-\item GPS not accurate.
-\item Ultrasound. expensive
-\item use sound. Everyone has access to. One application is to draw with music
-\end{enumerate}
+
+%Localization has been a widely researched topic in both academia and industry due to its wide range of applications.
+Companies and individuals have bennefited from technologies such as GPS, ultrasound, and blah.
+Since xxx  GPS technology has been invented and built into cars for navigation which provided great convenience for drivers.
+Ultrasound when used by sonar, help greatly helped underwater mapping.
+However, when it comes to localizationon a smaller scale such as the inside of a room, GPS would not be of much help because it's accuracy range is $1.4$ m~\cite{intro:gps}.
+
+Accuracte indoor locazation allows creation of noval applications with surrounding awareness that uses position and movment information as input.
+One application is to allow users to draw with music, without physically touching the computer.
+Another example is to build AI games with physical pieces such as toy car racing where the computer controls some of the pieces.
+In this work, we aim to build such a source localization system that is portable, inexpensive, yet accuracte for localization in a small area.
+
+Global Positioning System (GPS) is the prevailing technology used for outdoor localization.
+Commercial grade GPS has an average error of a few meters, depending on the size and quality of the receiver~\cite{intro:gps}.
+While accuracy in this range is good for many applications including driving navigation and vehicle tracking, it does not provide enough precision for local movement tracking.
+Ultrasound based indoor localization has achieved sub-centimeter accuracy~\cite{intro:ultra}.
+However, ultrasound systems require use of expensive transducers.
+
+Bluetooth and Wi-Fi based technologies gained popularity in indoor positioning recently, mainly due to the widespread deployment of bluetooth tags and Wi-Fi stations in public spaces.
+In these systems, signal strength received from different base stations are used for the estimation of the device location.
+However, their reported accuracy are in the range of $1$ to $5$ meters~\cite{intro:blue, intro:loc}, which is not enough for local movement tracking.
+
+In this project, we have built a local area localization system using microphone array that localizes normal audio source.
+The system is built with inexpensive electret microphones.
+People can interact with the system using any device that has audio output such as a mobile phone.
+
+In this paper, we discuss prior relevant research and the theory in sound localization in Section II.
+In Section III, we carried out a simulation to evaluate different array achitectures and their impact on accuracy.
+In Section IV, we present the chosen architecture along with hardware details.
+Finally, we provide the experiment details and results in Section V.

### 2015-05-19 reports/paper/prior.tex
paper update

@@
 \section{Background}
-Acoustic localization has been researched extensively in the literature. Localization techniques can be broadly put into two categories:
-\begin{enumerate}
-\item Location Template Matching (LTM) based approaches.
-\item Time Difference of Arrival (TDOA) based approaches.
-\end{enumerate}
+Acoustic localization has been researched extensively in the literature.
+Localization techniques can be broadly categorized into Location Template Matching (LTM) based approaches and Time Difference of Arrival (TDOA) based approaches.
 

### 2015-05-20 reports/paper/introduction.tex
paper update

@@
 Another example is to build AI games with physical pieces such as toy car racing where the computer controls some of the pieces.
-In this work, we aim to build such a source localization system that is portable, inexpensive, yet accuracte for localization in a small area.
+In this work, we aim to build such a source localization system that is portable, inexpensive, yet resonably accuracte for localization in a small area.
 
@@
 
-Bluetooth and Wi-Fi based technologies gained popularity in indoor positioning recently, mainly due to the widespread deployment of bluetooth tags and Wi-Fi stations in public spaces.
+Bluetooth and Wi-Fi based technologies have gained popularity in indoor positioning recently, mainly due to the widespread deployment of bluetooth tags and Wi-Fi stations in public spaces.
 In these systems, signal strength received from different base stations are used for the estimation of the device location.
@@
 In this paper, we discuss prior relevant research and the theory in sound localization in Section II.
-In Section III, we carried out a simulation to evaluate different array achitectures and their impact on accuracy.
+In Section III, we evaluated different array achitectures and their impact on accuracy.
+We demonstrated that a two array system outperforms a single array system of similar size.
 In Section IV, we present the chosen architecture along with hardware details.
-Finally, we provide the experiment details and results in Section V.
+Finally, experiment details and results are presented in Section V.

### 2015-05-20 reports/paper/experiment.tex
paper update

@@
 \subsubsection{Movement tracking}
-To test how well the arrays track movement, we mounted a rotating disk $40$ centimeter in diameter onto the grid at $(x=0, y=-0.3)$.
+To test how well the arrays track movement, we mounted a rotating disk $40$ centimeter in diameter onto the grid at $(x=0$ cm$, y=-30$ cm$)$.
 Fig~\ref{fig:setup_circle} shows a picture of the sestup.
@@
 
-\item[Music A] A music that has normal audio amplitude throughout experiment period was chosen. "Honest Eyes" by Black Tide was the music used.
+\item[Music A] A music that has normal audio amplitude throughout experiment period was chosen. ``Honest Eyes'' by Black Tide was the music used.
 
-\item[Music B] A music with intermittent low amplitude sections was chosen. "Canon" was the music used.
+\item[Music B] A music with intermittent low amplitude sections was chosen. ``Canon'' was the music used.
 
@@
 Fig~\ref{fig:accuracy_vs_window} shows how accuracy changes with window length for three GCC algorithms.
-The error lowers as window size increases and plataeus after window size exceeds around $10$ milli-second.
+The error decreases as window size increases and plataeus after window size exceeds around $10$ millisecond.
 The lowest error achieved is $2.53$ centimeters.
@@
 The error is below $3$ cm for most areas inside the region.
-There is one error spike in the mid-left region and we contribute this to audio source placecment error becuase the error is fairly low and consistent around that spike region.
+There is one error spike in the mid-left region and we contribute this to audio source misplacecment becuase the error is fairly low and consistent around that spike region.
 
@@
 
-Fig~\ref{fig:circle_normal} shows results for experiments with at normal speed, and Fig~\ref{fig:circle_fast} shows results at fast speed.
+Fig~\ref{fig:circle_normal} shows results for experiments at normal speed, and Fig~\ref{fig:circle_fast} shows results at fast speed.
 By comparing localization error for each audio source between normal movement speed and fast movement speed, we find the localization error does not depend on how fast the sound sorce is moving.
@@
 This is consistent with our expectation because low amplitude regions in Music B would cause arrays to lose track of where the source is.
-This can also be seen from the "blank" regions in fig~\ref{fig:circle_musicbn}.
+This can also be seen from the ``blank'' regions in fig~\ref{fig:circle_musicbn}.
 

### 2015-05-20 reports/paper/prior.tex
paper update

@@
 The requirement of having a template for each location to be detected makes this approach too restrictive for our project, since we want the drawing to be continuous in the 2D region.
-Moreover, the need to recalibrate all locations during setup is too curbursome for end users in a portable drawing application.
+Moreover, the need to recalibrate all locations during setup is too curbursome for end users in a portable localization system.
 Therefore, our main focus is on TDOA approaches.
@@
 
-Cross-correlation has been used extensively in measuring signal delays\cite{tdoa:mit2, tdoa:3}.
+Cross-correlation has also been used in measuring signal delays\cite{tdoa:mit2, tdoa:micloc, tdoa:3}.
 Cross-correlation with different prefiterings (known as \emph{generalized cross correlation (GCC)}) have also been investigated to improve delay estimation~\cite{tdoa:gcc1,tdoa:gcc2,tdoa:gcc3}.
@@
 \begin{eqnarray} \label{eq:gcc}
-t_0 &=& \arg\max_{\tau} R_{x_1x_2}(\tau) \\
+t_0 &=& \arg\max_{\tau} R_{x_1x_2}(\tau) \\\label{eq:gcc2}
 R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
@@
 $W(\omega)$ provides a way to prefilter signals passed to cross correlation estimator.
-We focused with three ways of prefiltering the signal:
+We focused on three ways of prefiltering the signal:
 \begin{description}[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]
-\item[GCC] $W(\omega) = 1$. No prefiltering is done. This is normal cross correlation.
+\item[GCC] $W(\omega) = 1$. No prefiltering is done. This is unfiltered normal cross correlation.
 \item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divideded by its magnitude. Only phase information contributes to delay estimation.

### 2015-05-20 reports/paper/simulation.tex
paper update

@@
 
-To see how sound source distance affects localization accuracy, the same simulation is carried out with sound source moved from $20$ cm to $80$ cm away from the center of the array.
+To see how sound source distance affects localization accuracy, the same simulation is carried out with the sound source moved from $20$ cm to $80$ cm away from the center of the array.
 Results are presetned in fig~\ref{fig:sim_3_8}.
@@
 In fig~\ref{fig:sim_hm_3_line}, three microphones are placed $10$ cm apart from each other on x-axis.
-Error heatmap showed high uncertainty on the x axis, and the overall accuracy is not as good as that with three microphones placed in a triangle.
+Error heatmap shows high uncertainty on the x axis, and the overall accuracy is not as good as that with three microphones placed in a triangle.
 The average error is $55.05$ cm.
@@
 To further increase the distance between microphones, we placed four microphones at four corners of the region.
-Fig~\ref{fig:sim_hm_4} showed the result.
+Fig~\ref{fig:sim_hm_4} shows the result.
 With this configuration, accuracy is consistently good across the region.
@@
 The system is less portable compared to small arrays with microphones near each other.
+Placing microphones far apart from each other also causes problems in TDOA estimation, because sampling of microphones in the same array requires synchronized clock.
 
-To avoid the need to accurate place four microphones at far distance(as required by fig~\ref{fig:sim_hm_4}), we explored configuration with two arrays.
+To avoid the need to accurately place four microphones at far distances(as required by fig~\ref{fig:sim_hm_4}), we explored configuration with two arrays.
 Two $3$ microphone array are placed $1$ meter apart and the result is presented in fig~\ref{fig:sim_hm_2_array}.

### 2015-05-20 reports/paper/system.tex
paper update

@@
 
-To speed up computation for real time localization, instead of searching for $t_0$ that maximizes equation~\ref{eq:gcc}, a grid search in 2D grid is performed.
+To handle uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation function~\ref{eq:gcc2} as a mapping from delay to liklihood.
+For each microphone array, we build a liklihood heatmap of each point in the grid being the sound sorce.
 For each point in the grid, the theoretical TDOA to each microphone pair can be precomputed.
-Then localization resolves to calculating GCC for each microphone pair and performing a lookup for each point in the grid.
-To further improve localization accuracy, instead of using point estimate for TDOA, a liklihoood map is built.
-Each entry in GCC output is used as the liklihood for that delay.
+Then the heatmap can be generated by going through points in the grid and perform a lookup using equation~\ref{eq:gcc2}.
 With three microphones $m_1,m_2,m_3$, there are three microphones pairs: $m_1m_2,m_1m_3,m_2m_3$.
 Theoretical TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$.
-Then the liklihood map can be built as:
+Then the liklihood map $L(x,y)$ can be built as:
 \begin{eqnarray*}
@@
 
-To see the effect of using multiple arrays, fig~\ref{fig:liklihood} shows the individual liklihood map from each array and also the combined liklihood map according to equation~\ref{eq:combine_l}.
-Individual array gives accurate angle estimate, but has high uncertainty in distance estimate.
-By combining estimates from two arrays, the angle estimate can be effectively combined to estimate distance.
+To see the effect of accuracy improvement using multiple arrays, fig~\ref{fig:liklihood} shows a real life localization where the source is placed at $(0$ cm$,-30$ cm$)$ .
+It shows the individual liklihood map from each array and also the combined liklihood map according to equation~\ref{eq:combine_l}.
+Individual array gives accurate angle estimate, but has high uncertainty in source distance.
+It demonstrated that by combining estimates from two arrays the system is able to perform more accurate localization.
 

### 2015-05-21 reports/paper/simulation.tex
paper update

@@
 The size of the uncertianty region is also angle dependent: points closer to the line of microphones have larger region compared to points close to the line perpenticular to microphones.
+
+This can also be seen analytically. Assuming two microphones are placed at $M_1:(-c,0)$ and $M2:(c,0)$. All points $P:(x,y)$ with difference of distance $ |PM_1 - PM_2| = 2a$ satisfies:
+\begin{eqnarray}\label{eqn:hyperbola}
+\frac{x^2}{a^2} - \frac{y^2}{c^2-a^2} = 1
+\end{eqnarray}
+To see how difference of distance changes with respect to distance, we can expand the equation and take partial derivative with respect to $x$ we can get:
+\begin{eqnarray}\label{eqn:derivative}
+\frac{\partial a}{\partial x} = \frac{x(c^2-a^2)}{a(x^2+y^2+c^2)-2a^3}
+\end{eqnarray}
+Since all points in equation~\ref{eqn:derivative} must lie on the hyperbola, we can substitute~\ref{eqn:hyperbola} into~\ref{eqn:derivative}:
+\begin{eqnarray}\label{eqn:derivativeF}
+\frac{\partial a}{\partial x} = \frac{c^2-a^2}{\frac{c^2}{a}x - \frac{a^3}{x}}
+\end{eqnarray}
+
+The denominator of equation~\ref{eqn:derivativeF} increases monotonically as $|x|$ increases, which indicates $\frac{\partial a}{\partial x}$ decreases as we move farther away.
+The same distance move $\delta x$ would generate smaller change in difference of distance $a$ when source is farther away from the microphones.
 

### 2015-05-21 reports/paper/simulation.tex
paper update

@@
 \[
- (\hat P M_1 - \hat P M_2) - (P M_1 - P M_2) < 1 \mbox{ cm}
+|(\hat P M_1 - \hat P M_2) - (P M_1 - P M_2)|< 1 \mbox{ cm}
 \]
@@
 \end{eqnarray}
-To see how difference of distance changes with respect to distance, we can expand the equation and take partial derivative with respect to $x$ we can get:
+To see how difference of distance changes with respect to distance, we can expand the equation and find partial differential $\frac{\partial a}{\partial x}$:
 \begin{eqnarray}\label{eqn:derivative}
@@
 
-The denominator of equation~\ref{eqn:derivativeF} increases monotonically as $|x|$ increases, which indicates $\frac{\partial a}{\partial x}$ decreases as we move farther away.
+The denominator of equation~\ref{eqn:derivativeF} increases monotonically as $|x|$ increases, which indicates $\frac{\partial a}{\partial x}$ decreases as we move farther away along the hyperbola.
 The same distance move $\delta x$ would generate smaller change in difference of distance $a$ when source is farther away from the microphones.

### 2015-05-21 reports/paper/experiment.tex
paper update

@@
 \begin{figure}[]
-\includegraphics[width=0.5\textwidth]{error_distribution.eps}
+\centering
+\includegraphics[width=0.3\textwidth]{error_distribution.eps}
 \caption{error distribution in the grid}
@@
 \centering
-\begin{subfigure}[]{1.0\textwidth}
-  \includegraphics[width=1.0\textwidth]{trace_output_circle_wnn.eps}
-  \caption{white noise}
+\includegraphics[width=.8\textwidth]{trace_output_circle_wnn.eps}
+\caption{white noise ($10$ cm per second)}
 \label{fig:circle_wnn}
-\end{subfigure}
-\begin{subfigure}[]{1.0\textwidth}
-  \includegraphics[width=\textwidth]{trace_output_circle_man.eps}
-  \caption{music A}
-  \label{fig:circle_musican}
-\end{subfigure}
-\begin{subfigure}[]{1.0\textwidth}
-  \includegraphics[width=\textwidth]{trace_output_circle_mbn.eps}
-  \caption{music B}
-  \label{fig:circle_musicbn}
-\end{subfigure}
-\caption{Localization of circle movement with different sound sources. Sound source is moving at $10$ cm per second}
-\label{fig:circle_normal}
 \end{figure*}
@@
 \centering
-\begin{subfigure}[]{1.0\textwidth}
-  \includegraphics[width=\textwidth]{trace_output_circle_wnf.eps}
-  \caption{white noise}
-  \label{fig:circle_wnf}
-\end{subfigure}
-\begin{subfigure}[]{1.0\textwidth}
-  \includegraphics[width=\textwidth]{trace_output_circle_maf.eps}
-  \caption{music A}
-  \label{fig:circle_musicaf}
-\end{subfigure}
-\begin{subfigure}[]{1.0\textwidth}
-  \includegraphics[width=\textwidth]{trace_output_circle_mbf.eps}
-  \caption{music B}
-  \label{fig:circle_musicbf}
-\end{subfigure}
-\caption{Localization of circle movement with different sound sources. Sound source is moving at $20$ cm per second}
-\label{fig:circle_fast}
+\includegraphics[width=.8\textwidth]{trace_output_circle_man.eps}
+\caption{music A ($10$ cm per second)}
+\label{fig:circle_musican}
 \end{figure*}
 
-Fig~\ref{fig:circle_normal} shows results for experiments at normal speed, and Fig~\ref{fig:circle_fast} shows results at fast speed.
+\begin{figure*}[]
+\centering
+\includegraphics[width=.8\textwidth]{trace_output_circle_mbn.eps}
+\caption{music B ($10$ cm per second)}
+\label{fig:circle_musicbn}
+\end{figure*}
+
+\begin{figure*}[]
+\centering
+\includegraphics[width=.8\textwidth]{trace_output_circle_wnf.eps}
+\caption{white noise ($20$ cm per second)}
+\label{fig:circle_wnf}
+\end{figure*}
+
+\begin{figure*}[]
+\centering
+  \includegraphics[width=.8\textwidth]{trace_output_circle_maf.eps}
+  \caption{music A ($20$ cm per second)}
+  \label{fig:circle_musicaf}
+\end{figure*}
+
+\begin{figure*}[]
+\centering
+  \includegraphics[width=.8\textwidth]{trace_output_circle_mbf.eps}
+  \caption{music B ($20$ cm per second)}
+  \label{fig:circle_musicbf}
+\end{figure*}
+
+Fig~\ref{fig:circle_wnn} to \ref{fig:circle_musicbn} shows results for experiments at normal speed, and Fig~\ref{fig:circle_wnf} to~\ref{fig:circle_musicbf} shows results at fast speed.
 By comparing localization error for each audio source between normal movement speed and fast movement speed, we find the localization error does not depend on how fast the sound sorce is moving.
@@
 
-From fig~\ref{fig:circle_normal}, localization error is $0.9$ cm for white noise, $1.29$ cm for Music A, and $2.9$ cm for Music B.
+For normal speed movement tracking, localization error is $0.9$ cm for white noise, $1.29$ cm for Music A, and $2.9$ cm for Music B.
 Localization accuracy is the best for white noise source, and worst for Music B.
@@
 
-Fig ~\ref{fig:circle_normal} also shows that raw detection has the most amount of jiggling.
+It also shows that raw detection has the most amount of jiggling.
 Kalman filter reduces the amount of jiggling from raw detection.

### 2015-05-22 reports/paper/experiment.tex
update
paper update. Significiant help from Max Yuan

@@
 
-To test how accuracy varies with window size, the algorithm is fed with recorded microphone data with different segment length.
+To test how accuracy varies with window size, the algorithm is fed with recorded microphone data of different window size.
 Fig~\ref{fig:accuracy_vs_window} shows how accuracy changes with window length for three GCC algorithms.
-The error decreases as window size increases and plataeus after window size exceeds around $10$ millisecond.
+The error decreases as window size increases and plateaus after window size exceeds around $10$ millisecond.
 The lowest error achieved is $2.53$ centimeters.
@@
 \begin{figure}[]
-\includegraphics[width=0.5\textwidth]{calculation_time.eps}
+\centering
+\includegraphics[width=0.4\textwidth]{calculation_time.eps}
 \caption{speed versus window size}
@@
 \begin{figure}[]
-\includegraphics[width=0.5\textwidth]{error_window_size_movement.eps}
+\centering
+\includegraphics[width=0.4\textwidth]{error_window_size_movement.eps}
 \caption{Localization error versus window size}
@@
 
-Fig~\ref{fig:err_win_circle} gives an intuitive representation of how accuracy changes with window size.
+Fig~\ref{fig:err_wn} gives an intuitive representation of how accuracy changes with window size.
 When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate TDOA.

### 2015-05-22 reports/paper/introduction.tex
paper update. Significiant help from Max Yuan

@@
 Accuracte indoor locazation allows creation of noval applications with surrounding awareness that uses position and movment information as input.
-One application is to allow users to draw with music, without physically touching the computer.
-Another example is to build AI games with physical pieces such as toy car racing where the computer controls some of the pieces.
+One application is to allow users to draw with music without physically touching the computer.
+Another example is to build AI games with physical pieces such as toy car racing where the computer controls some them.
 In this work, we aim to build such a source localization system that is portable, inexpensive, yet resonably accuracte for localization in a small area.
@@
 While accuracy in this range is good for many applications including driving navigation and vehicle tracking, it does not provide enough precision for local movement tracking.
-Ultrasound based indoor localization has achieved sub-centimeter accuracy~\cite{intro:ultra}.
-However, ultrasound systems require use of expensive transducers.
+Ultrasound based indoor localization approaches on the other hand, has achieved sub-centimeter accuracy~\cite{intro:ultra}.
+However, ultrasound systems require the use of expensive transducers.
 
@@
 
-In this project, we have built a local area localization system using microphone array that localizes normal audio source.
-The system is built with inexpensive electret microphones.
-People can interact with the system using any device that has audio output such as a mobile phone.
+In this project, we have built a localization system with reasonable high precision for small area using microphone arrays that localize typical audio source.
+Our system is built with inexpensive electret microphones mounted on portable frames.
+Users can interact with our system using any device that has audio output such as a mobile phone.
 
-In this paper, we discuss prior relevant research and the theory in sound localization in Section II.
-In Section III, we evaluated different array achitectures and their impact on accuracy.
-We demonstrated that a two array system outperforms a single array system of similar size.
-In Section IV, we present the chosen architecture along with hardware details.
+In this paper, we first discussed in Section II some prior relevant research and approaches in sound localization.
+In Section III, we evaluated different array achitectures and their impact on localization accuracy.
+We demonstrated that a two array system outperforms a single array system of similar physical dimension.
+In Section IV, we presented the chosen architecture along with hardware details.
 Finally, experiment details and results are presented in Section V.

### 2015-05-22 reports/paper/prior.tex
paper update. Significiant help from Max Yuan

@@
 Acoustic localization has been researched extensively in the literature.
-Localization techniques can be broadly categorized into Location Template Matching (LTM) based approaches and Time Difference of Arrival (TDOA) based approaches.
+Localization techniques can be broadly categorized into Location Template Matching (LTM) and Time Difference of Arrival (TDOA) based approaches.
 
 \subsection{LTM}
-In LTM based approaches, acoustic templates of locations are first stored in the system during a ``training'' phase.
-Then, incoming acoustic waveform is compared with stored templates.
-Localization result is the location that has the best matching template.
+In LTM based approaches, acoustic templates acquired from different locations are first stored in the system during a ``training'' phase.
+Localization can be performed by comparing the incoming waveform with the stored templates, and the location with the best matching template is chosen as the output.
 Different ways of extracting templates from raw acoustic source and different similarity measures have been investigated in the past.
 
-\cite{extended:tusi} and \cite{ltm:pham} investigated using max value from corss-correlation as similarity measure to localize user tap on interactive surface.
-\cite{ltm:lpc} used L2 distance in Linear Predictive Coding coeeficient space as similarity measure to localize taps on surfaces.
-\cite{ltm:tusi2} further explored accuracy improvement by using multiple templates for each location and speed improvement by merging multiple templates into one representative templates.
+\cite{extended:tusi} and \cite{ltm:pham} investigated using max value from cross-correlation as a similarity measure to localize user tap on interactive surfaces.
+\cite{ltm:lpc} used L2 distance in the Linear Predictive Coding coefficient space as a similarity measure to localize taps on surfaces.
+\cite{ltm:tusi2} further explored accuracy improvement by using multiple templates for each location and speed improvement by merging multiple templates into one representative template.
 
-The requirement of having a template for each location to be detected makes this approach too restrictive for our project, since we want the drawing to be continuous in the 2D region.
-Moreover, the need to recalibrate all locations during setup is too curbursome for end users in a portable localization system.
-Therefore, our main focus is on TDOA approaches.
+The requirement of having a template for each location to be detected makes this approach too restrictive for our project, since we want the localization to be continuous in a 2D region.
+Moreover, the need to recalibrate all locations during setup is too cumbersome for the end users in a portable system.
+Therefore, our main focus will be on TDOA based approaches.
 
 \subsection{TDOA}
-Points with the same TDOA to two fixed points on a plane form a hyperbola.
-With more than two microphones, intersection of hyperbola curves generated by each pair of microphoens give the source location.
-Localization relies on accurate estimate of delay differences between microphones.
+TDOA approaches exploit the difference of arrival time between the acoustic source and two fixed microphones on the plane.
+It can be easily shown that the acoustic sources with the same TDOA to two fixed microphones on the plane form a hyperbola.
+When you have more than two microphones, each pair would give a different hyperbola.
+The intersection of all the hyperbolas marks the source location.
+TDOA approaches rely on accurate estimates of arrival time differences between microphones.
 
-In \cite{tdoa:ppp}, authors used eight microphnes mounted on corners of a ping pong table to localize the point where the ping pong ball hits the table.
+In \cite{tdoa:ppp}, authors used eight microphones mounted on the corners of a ping pong table to localize points where the ball hits the table.
 They used a threshold to determine the arrival time of acoustic signal.
-This approach works well in noise free environment but performance degrades with background noise.
-It also suffers from dispersive deflections that arrives before the main wavefront.
+This approach works well in noise free environment but the performance degrades with background noise.
+Their approach also suffers from dispersive deflections that arrive before the main wavefront of the acoustic signal.
 To make it more robust, authors in \cite{tdoa:mit3} and \cite{tdoa:mit4} extracted descriptive parameters for each significant peak(e.g., peak height, width, mean arrival time).
@@
 
-Cross-correlation has also been used in measuring signal delays\cite{tdoa:mit2, tdoa:micloc, tdoa:3}.
-Cross-correlation with different prefiterings (known as \emph{generalized cross correlation (GCC)}) have also been investigated to improve delay estimation~\cite{tdoa:gcc1,tdoa:gcc2,tdoa:gcc3}.
+Cross-correlation has also been used to measure signal arrival time differences\cite{tdoa:mit2, tdoa:micloc, tdoa:3}.
+Cross-correlation with prefilterings is known as \emph{generalized cross correlation (GCC)}.
+Different prefilterings have been investigated to improve arrival time difference estimation~\cite{tdoa:gcc1,tdoa:gcc2,tdoa:gcc3}.
 
-GCC provides a framework to estimate delay differences $t_0$ between two signals $x_1(t)$ and $x_2(t)$:
+Under the GCC framework, the arrival time difference $t_0$ between two signals $x_1(t)$ and $x_2(t)$ can be estimated as:
 \begin{eqnarray} \label{eq:gcc}
@@
 , where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transform of $x_1(t)$ and $x_2(t)$.
-$W(\omega)$ provides a way to prefilter signals passed to cross correlation estimator.
+$W(\omega)$ provides a way to prefilter signals passed to the cross correlation estimator.
 We focused on three ways of prefiltering the signal:

### 2015-05-22 reports/paper/simulation.tex
paper update
paper update. Significiant help from Max Yuan

@@
 \section{Array Architecture}
-As was mentioned in the previous section, points with the same TDOA to two fixed locations form a hyperbola in a 2D plane.
+As was mentioned in the previous section, points with the same TDOA to two fixed locations form a hyperbola on a 2D plane.
 In practical systems, we can only measure TDOA up to a precision.
 Therefore we look at all points with difference of distance close to some target value within measurement error $\epsilon$.
-This $\epsilon$ prepresents accuracy on measurement of difference of distances, and in practice it is related to sampling rate and TDOA methods.
+This $\epsilon$ represents accuracy on the measurement of difference of distances, and in practice it is related to sampling rate and TDOA methods.
 In this section we evaluate delay estimation's impact on localization accuracy.
@@
 Intuitively, points in the region have difference of distance very similar to each other.
-From fig~\ref{fig:sim_2_5}, the region still has the shape of a hyperbola, but with an uncertainty region around the curvce.
+From fig~\ref{fig:sim_2_5}, the region still has the shape of a hyperbola, but with an uncertainty region around the curve.
 The uncertainty region is not uniform around the curve, the farther away the point is, the larger the uncertainty region becomes.
-It indicates that the same delta distance movement will generate smaller difference of distance when the source is farther away from the array.
+This indicates for the the same delta distance movement it will generate smaller difference of distance change when the source is farther away from the array.
 The size of the uncertianty region is also angle dependent: points closer to the line of microphones have larger region compared to points close to the line perpenticular to microphones.
 
-This can also be seen analytically. Assuming two microphones are placed at $M_1:(-c,0)$ and $M2:(c,0)$. All points $P:(x,y)$ with difference of distance $ |PM_1 - PM_2| = 2a$ satisfies:
+This can also be seen analytically. Assuming two microphones are placed on the x-axis at $M_1:(-c,0)$ and $M2:(c,0)$. All points $P:(x,y)$ with difference of distance $ |PM_1 - PM_2| = 2a$ satisfies:
 \begin{eqnarray}\label{eqn:hyperbola}
@@
 The denominator of equation~\ref{eqn:derivativeF} increases monotonically as $|x|$ increases, which indicates $\frac{\partial a}{\partial x}$ decreases as we move farther away along the hyperbola.
-The same distance move $\delta x$ would generate smaller change in difference of distance $a$ when source is farther away from the microphones.
+The same distance move $\delta x$ would generate smaller change in difference of distance $a$ when the source is farther away from the microphones.
 
@@
 
-With more than two multiple microphones, each pair of microphones generates a hyperbolic region and localization becomes finding the intersection of hyperbolic regions.
+With more than two microphones, each pair of microphones generates a hyperbolic region and localization becomes finding the intersection of hyperbolic regions.
 The smaller the intersection region, the better the localization accuracy.
@@
 Fig~\ref{fig:sim_3_2} shows the intersection of regions for $5$ different placement of the sound source.
-It can be seen that accuracy is worse when sound source is close to the line of any two microphones.
-This observation is consistent with two microphone case, since points close to line of microphones have a larger uncertainty region.
+It can be seen that accuracy is worse when sound source is close to the line connecting any two microphones.
+This observation is consistent with two microphone case, since points close to lines connecting microphones have a larger uncertainty region.
 
 To see how sound source distance affects localization accuracy, the same simulation is carried out with the sound source moved from $20$ cm to $80$ cm away from the center of the array.
-Results are presetned in fig~\ref{fig:sim_3_8}.
-Comparing with fig~\ref{fig:sim_3_2}, accuracy decreases with distance to the array.
+Results are presented in fig~\ref{fig:sim_3_8}.
+Comparing with fig~\ref{fig:sim_3_2}, accuracy decreases as the distance to the array increases.
 This is also consistent with our observation in $2$ microphone case where source farther away would result in larger uncertainty region.
@@
 The result indicates that this configuration has good accuracy when source is close to the arrays.
-Accuracy decreases as sound source moves outside the one meter by one meter region.
+Accuracy decreases as sound source moves outside of the one meter by one meter region.
 The average error is $2.60$ cm.

### 2015-05-22 reports/paper/system.tex
paper update. Significiant help from Max Yuan

@@
 \section{System Setup}
-The end system has two arrays, each with three microphones mounted on vetices of a $20$ cm equilateral triangle.
+The end system has two arrays, each with three microphones mounted on vertices of a $20$ cm equilateral triangle.
 A micro-controller is also attached to one of the vertices.
@@
 To handle uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation function~\ref{eq:gcc2} as a mapping from delay to liklihood.
-For each microphone array, we build a liklihood heatmap of each point in the grid being the sound sorce.
-For each point in the grid, the theoretical TDOA to each microphone pair can be precomputed.
-Then the heatmap can be generated by going through points in the grid and perform a lookup using equation~\ref{eq:gcc2}.
+For each microphone array, we build a heatmap for the region.
+The intensity at each point represents the likelihood of it being the source.
+For each point on the grid, the theoretical TDOA to each microphone pair can be precomputed.
+Then the heatmap can be generated by going through points on the grid and perform a lookup using equation~\ref{eq:gcc2}.
 With three microphones $m_1,m_2,m_3$, there are three microphones pairs: $m_1m_2,m_1m_3,m_2m_3$.

### 2015-05-22 reports/paper/simulation.tex
yuan's improvement

@@
 As was mentioned in the previous section, points with the same TDOA to two fixed locations form a hyperbola on a 2D plane.
-In practical systems, we can only measure TDOA up to a precision.
+However, in practical systems we can only measure TDOA up to a precision.
 Therefore we look at all points with difference of distance close to some target value within measurement error $\epsilon$.
-This $\epsilon$ represents accuracy on the measurement of difference of distances, and in practice it is related to sampling rate and TDOA methods.
-In this section we evaluate delay estimation's impact on localization accuracy.
+This $\epsilon$ represents accuracy on the measurement of difference of distances, and in practice it is related to sampling rate and estimation of difference of arrival time.
+In this section we evaluate the impact of difference of distance estimation on localization accuracy.
 
@@
 To see how precision affects localization accuracy, we simulated two microphones placed at: $M_1:(x=-10\mbox{ cm},y=0\mbox{ cm})$ and $M_2:(x=10\mbox{ cm},y=0\mbox{ cm})$.
-A test sound source is emitted at point $P$ which is $50$ centimeters away from $(0,0)$.
-Fig~\ref{fig:sim_2_5} shows the region where all points $\hat P$ satisfy:
+A test sound source is emitted at point $P$ which is $50$ centimeters away from the origin $(0,0)$.
+Fig~\ref{fig:sim_2_5} shows the region $R$ where all points $\hat P$ satisfy:
 \[
-|(\hat P M_1 - \hat P M_2) - (P M_1 - P M_2)|< 1 \mbox{ cm}
+R:\{\hat P: |(\hat P M_1 - \hat P M_2) - a|< 1 \mbox{ cm}\}
 \]
-Intuitively, points in the region have difference of distance very similar to each other.
-From fig~\ref{fig:sim_2_5}, the region still has the shape of a hyperbola, but with an uncertainty region around the curve.
-The uncertainty region is not uniform around the curve, the farther away the point is, the larger the uncertainty region becomes.
+, where $a$ is the difference of distance between $P M_1$ and $P M_2$:
+\[
+P M_1 - P M_2 = a
+\]
+Intuitively, points in $R$ have difference of distance very similar to each other.
+Looking at fig~\ref{fig:sim_2_5}, we can still see that $R$ has the shape of a hyperbola, but with an uncertainty region around it.
+The thickness of the uncertainty region is not uniform around the hyperbola, the farther away the point is, the larger the uncertainty region becomes.
 This indicates for the the same delta distance movement it will generate smaller difference of distance change when the source is farther away from the array.
-The size of the uncertianty region is also angle dependent: points closer to the line of microphones have larger region compared to points close to the line perpenticular to microphones.
+The size of the uncertianty region is also angle dependent: points closer to the line connecting microphones have larger region compared to points close to the line bisecting microphones.
 
@@
 The smaller the intersection region, the better the localization accuracy.
-To see how accuracy changes with array placement and sound source location, three microphones are placed at three vertices of an $20$ cm equilateral triangle.
+To see how accuracy changes with array placement and sound source location, three microphones are placed at the three vertices of an $20$ cm equilateral triangle.
 An audio source is placed at $20$ cm away from the center of the array.
 Fig~\ref{fig:sim_3_2} shows the intersection of regions for $5$ different placement of the sound source.
-It can be seen that accuracy is worse when sound source is close to the line connecting any two microphones.
+It can be seen that accuracy is decreases when sound source becomes close to the line connecting any two microphones.
 This observation is consistent with two microphone case, since points close to lines connecting microphones have a larger uncertainty region.
@@
 The region inside the array has good accuracy.
-However, for regions along line of any two microphones, the accuracy drops significantly.
+However, for regions along the line of any two microphones, the accuracy drops significantly.
 Average error across the region is $18.6$ cm.
@@
 
-With simulation results, we decided to build the two array system as described in fig~\ref{fig:sim_hm_2_array}.
+With the simulation results, we decided to build the two array system as described in fig~\ref{fig:sim_hm_2_array}.
 The setup is reasonably portable (compared to fig~\ref{fig:sim_hm_4}), while at the same time having significantly better accuracy compared to one array systems.

### 2015-05-22 reports/paper/experiment.tex
paper update

@@
 Fig~\ref{fig:setup_point} shows a picture of the setup.
-A total of $32$ positions are chosen uniformly in this region where microphone data is recorded.
+A total of $32$ testing locations are chosen uniformly in this.
 
@@
 Fig~\ref{fig:setup_circle} shows a picture of the sestup.
-A sound source is placed on the edge of rotating disk and the arrays localizes the sound source as it rotates in a circle.
+A sound source is placed on the edge of the rotating disk and the arrays localizes the sound source as it rotates in a circle.
 In this experiment, we tested how accuracy changes with:
@@
 
-To test how different sound sources impact localization quality, we conducted three experiments on the same movement track with three different sound sources:
+To test how different sound sources impact localization quality, we conducted the experiments with three different sound sources:
 
@@
 
-\item[Music A] A music that has normal audio amplitude throughout experiment period was chosen. ``Honest Eyes'' by Black Tide was the music used.
+\item[Music A] A randomly picked music that has normal audio amplitude throughout experiment period. ``Honest Eyes'' by Black Tide was the music used.
 
-\item[Music B] A music with intermittent low amplitude sections was chosen. ``Canon'' was the music used.
+\item[Music B] A randomly picked music with intermittent low amplitude sections. ``Canon'' was the music used.
 
@@
 
-To test how sound source movement speed affects localization quality, each of three experiments were conducted at two different speeds:
+To test how the movement speed of sound source affects localization quality, each experiment was conducted at two different speeds:
 \begin{description}
-\item[Normal] An angular speed of $0.5$ rad/s was maintained, which translates to linear speed of $10$ cm/s.
-\item[Fast] An angular speed of $1.0$ rad/s was maintained, which translates to linear speed of $20$ cm/s.
+\item[Normal] An angular speed of $0.5$ rad/s was maintained, which translates to a linear speed of $10$ cm/s.
+\item[Fast] An angular speed of $1.0$ rad/s was maintained, which translates to a linear speed of $20$ cm/s.
 \end{description}
 
-For each experiment conducted, two different movement filters are tests:
+For each experiment conducted, two different movement filters were tests:
 \begin{description}[\IEEEsetlabelwidth{Very long label}\IEEEusemathlabelsep]
@@
 \subsubsection{Point localization}
-
-To test how accuracy varies with window size, the algorithm is fed with recorded microphone data of different window size.
-Fig~\ref{fig:accuracy_vs_window} shows how accuracy changes with window length for three GCC algorithms.
-The error decreases as window size increases and plateaus after window size exceeds around $10$ millisecond.
+To test how accuracy varies with window size, the algorithm is fed with microphone data of different window size.
+Fig~\ref{fig:accuracy_vs_window} shows how accuracy changes with the window size for GCC, GCC\_PHAT and GCC\_PHAT\_SQRT.
+The error decreases as window size increases and plateaus after the window size exceeds around $10$ millisecond.
 The lowest error achieved is $2.53$ centimeters.
@@
 
-Although accuracy improves with window length, the calculation time also increases with window length.
-The part of calculation that depends on window length is using cross correlation to estimate TDOA.
-cross correlation can be calculated with FFT and the runtime is of order $O(N\log N)$.
-We measured how the computation time varies with window length and Figure~\ref{fig:speed_vs_window} shows the result.
+Although accuracy improves with the window size, calculation time also increases with it.
+The part of calculation that depends on the window size is using cross correlation for TDOA estimation.
+Cross correlation can be calculated with Fast Fourier Transform (FFT) and the runtime is of order $O(N\log N)$.
+We measured how the computation time varies with window size and Figure~\ref{fig:speed_vs_window} shows the result.
 The runtime increases approximately linearly in the window size region of interest.
@@
 We also calculated the localization error for each tested point in the region.
-Figure~\ref{fig:error_distribution} shows a heatmap of the error distribution inside the grid.
+Figure~\ref{fig:error_distribution} shows the error distribution inside the grid.
 The error is below $3$ cm for most areas inside the region.
-There is one error spike in the mid-left region and we contribute this to audio source misplacecment becuase the error is fairly low and consistent around that spike region.
+There is one error spike in the mid-left region and we contribute this to audio source misplacecment becuase the error is fairly low and consistent around that spike.
 
@@
 \includegraphics[width=0.3\textwidth]{error_distribution.eps}
-\caption{error distribution in the grid}
+\caption{error distribution in the grid. Arrays are placed at $(-0.5$ m$, 0$ m$)$ and $(0.5$ m$, 0$ m$)$}
 \label{fig:error_distribution}
@@
 
-Fig~\ref{fig:err_wn} gives an intuitive representation of how accuracy changes with window size.
+Fig~\ref{fig:wn} gives an intuitive representation of how accuracy changes with window size.
 When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate TDOA.
@@
 The general trend is similar to that in point localization case.
-The error decreases as window size increases and plateaus after window size exceeds around $10$ milliseconds.
+Error decreases as the window size increases and plateaus after it exceeds around $10$ milliseconds.
 

### 2015-05-22 reports/paper/simulation.tex
paper update

@@
 Therefore we look at all points with difference of distance close to some target value within measurement error $\epsilon$.
-This $\epsilon$ represents accuracy on the measurement of difference of distances, and in practice it is related to sampling rate and estimation of difference of arrival time.
+This $\epsilon$ represents accuracy on the measurement of difference of distances, and in practice it is related to sampling rate and estimation techniques.
 In this section we evaluate the impact of difference of distance estimation on localization accuracy.
@@
 A test sound source is emitted at point $P$ which is $50$ centimeters away from the origin $(0,0)$.
-Fig~\ref{fig:sim_2_5} shows the region $R$ where all points $\hat P$ satisfy:
+Let $2a$ denote the difference of distance between $P$ and two microphones:
 \[
-R:\{\hat P: |(\hat P M_1 - \hat P M_2) - a|< 1 \mbox{ cm}\}
+2a = P M_1 - P M_2 
 \]
-, where $a$ is the difference of distance between $P M_1$ and $P M_2$:
+, then fig~\ref{fig:sim_2_5} shows the region $R$ where all points have difference of distance to two microphones close to $a$ within $1$ cm:
 \[
-P M_1 - P M_2 = a
+R=\{\hat P: |(\hat P M_1 - \hat P M_2) - 2a|< 1 \mbox{ cm}\}
 \]
-Intuitively, points in $R$ have difference of distance very similar to each other.
+Intuitively, points in $R$ have difference of distance to two microphones very similar to each other.
 Looking at fig~\ref{fig:sim_2_5}, we can still see that $R$ has the shape of a hyperbola, but with an uncertainty region around it.
@@
 Fig~\ref{fig:sim_3_2} shows the intersection of regions for $5$ different placement of the sound source.
-It can be seen that accuracy is decreases when sound source becomes close to the line connecting any two microphones.
-This observation is consistent with two microphone case, since points close to lines connecting microphones have a larger uncertainty region.
+It can be seen that accuracy decreases when sound source becomes close to the line connecting any two microphones.
+This observation is consistent with the two microphone case, since points close to lines connecting microphones have a larger uncertainty region.
 
@@
 Comparing with fig~\ref{fig:sim_3_2}, accuracy decreases as the distance to the array increases.
-This is also consistent with our observation in $2$ microphone case where source farther away would result in larger uncertainty region.
+This is also consistent with our observation in $2$ microphone case where sources farther away would result in larger uncertainty region.
 
@@
 
-Intersection area is a measure of the localization accuracy.
+Each microphone pair generates a hyperbolic region, and the intersection area of these regions is a measure of the localization accuracy.
 To evaluate an array's accuracy in a region, we can place sound source at predetermined grid points in the region and look at the intersection area for each tested point in the grid.
-The center location of intersection region can be used as localization estimate to calculate localization error.
+The center location of intersection region can be used as the localization estimate to calculate localization error.
 Results for a few different microphone array configurations are presented in fig~\ref{fig:sim_hm}.
 
-Fig~\ref{fig:sim_hm_3} shows the accuracy when microphones are placed at three vertices of a $20$ cm equilateral triangle.
+Fig~\ref{fig:sim_hm_3} shows the accuracy when microphones are placed at the three vertices of a $20$ cm equilateral triangle.
 The region inside the array has good accuracy.
-However, for regions along the line of any two microphones, the accuracy drops significantly.
+However, for regions along the line connecting any two microphones, the accuracy drops significantly.
 Average error across the region is $18.6$ cm.
 
-To evaluate how adding one microphone(without increasing array size) improves accuracy,  another microphone is added to the array at $(0,0)$.
+To evaluate how adding one microphone(without increasing the array size) improves accuracy, another microphone is added to the array at $(0,0)$.
 Result is presented in fig~\ref{fig:sim_hm_3_p1}.
@@
 Average error dropped from $18.6$ cm to $17.1$ cm.
-Regions near lines of microphones still have significantly larger uncertainty region.
+Regions near lines connecting microphones still have significantly larger uncertainty region.
 
-To evaluate array size's impact on accuracy, the size of original array from fig~\ref{fig:sim_hm_3} is increased by a factor of $2$.
+To evaluate the array size's impact on accuracy, the size of the original array (as in fig~\ref{fig:sim_hm_3}) is increased by a factor of $2$.
 The result is presented in fig~\ref{fig:sim_hm_3_2x}.
@@
 
-In fig~\ref{fig:sim_hm_3_line}, three microphones are placed $10$ cm apart from each other on x-axis.
+In fig~\ref{fig:sim_hm_3_line}, three microphones are placed $10$ cm apart from each other on the x-axis.
 Error heatmap shows high uncertainty on the x axis, and the overall accuracy is not as good as that with three microphones placed in a triangle.
@@
 
-To avoid the need to accurately place four microphones at far distances(as required by fig~\ref{fig:sim_hm_4}), we explored configuration with two arrays.
+To avoid the need to accurately place microphones at far distances(as required by fig~\ref{fig:sim_hm_4}), we explored configuration with two arrays.
 Two $3$ microphone array are placed $1$ meter apart and the result is presented in fig~\ref{fig:sim_hm_2_array}.

### 2015-05-22 reports/paper/system.tex
paper update

@@
 \section{System Setup}
-The end system has two arrays, each with three microphones mounted on vertices of a $20$ cm equilateral triangle.
-A micro-controller is also attached to one of the vertices.
+The end system has two arrays, each with three microphones mounted on the vertices of a $20$ cm equilateral triangle.
+A micro-controller is attached to one of the vertices.
 Fig~\ref{fig:setup_array} shows a picture of the array setup.
-Micro-controller used in this project is \emph{teensy 3.1}.
-It has $64$k RAM memory and the ADC is capable of sampling at $500$kHz.
-In this project, the micro-controller collects microphone data on all three channels for $12$ millisecond and then send the recorded data to a computer through USB port for localization.
+The micro-controller used in this project is \emph{teensy 3.1}, which has $64$k RAM memory and the ADC is capable of sampling at $500$kHz~\cite{tdoa:micloc, sys:teensy}.
+In this project, the micro-controller collects microphone data on all three channels for $12$ millisecond and then sends the recorded data to a computer through the USB port for localization.
 
@@
 
-To handle uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation function~\ref{eq:gcc2} as a mapping from delay to liklihood.
+To handle the uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation output~\ref{eq:gcc2} as a measure of the likelihood of different arrival time differences.
 For each microphone array, we build a heatmap for the region.
@@
 For each point on the grid, the theoretical TDOA to each microphone pair can be precomputed.
-Then the heatmap can be generated by going through points on the grid and perform a lookup using equation~\ref{eq:gcc2}.
-With three microphones $m_1,m_2,m_3$, there are three microphones pairs: $m_1m_2,m_1m_3,m_2m_3$.
-Theoretical TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$.
-Then the liklihood map $L(x,y)$ can be built as:
+Then the heatmap can be generated by going through points on the grid and performing a lookup using equation~\ref{eq:gcc2}.
+With three microphones $m_1,m_2,$ and $m_3$, there are three microphone pairs: $m_1m_2,m_1m_3,$ and $m_2m_3$.
+The theoretical TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$.
+Then the liklihood map $L(x,y)$ can be built by superposing the liklihood from each microphone pair:
 \begin{eqnarray*}
@@
 
-Liklihood maps from two arrays can be combined into final liklihood map:
+Liklihood maps from two arrays can be combined into the final liklihood map:
 \begin{equation}\label{eq:combine_l}
@@
 \end{equation}
-, where $L_1(x,y)$ and $L_2(x,y)$ represents the liklihood map from array $1$ and array $2$.
+, where $L_1(x,y)$ and $L_2(x,y)$ represent the liklihood map from array $1$ and array $2$.
 
@@
 
-To see the effect of accuracy improvement using multiple arrays, fig~\ref{fig:liklihood} shows a real life localization where the source is placed at $(0$ cm$,-30$ cm$)$ .
+To see the effect of accuracy improvement using multiple arrays, fig~\ref{fig:liklihood} shows a real life localization where the source is placed at $(0$ cm$,-30$ cm$)$.
 It shows the individual liklihood map from each array and also the combined liklihood map according to equation~\ref{eq:combine_l}.
-Individual array gives accurate angle estimate, but has high uncertainty in source distance.
-It demonstrated that by combining estimates from two arrays the system is able to perform more accurate localization.
+Individual arrays give accurate angle estimate, but have high uncertainty in distance estimate.
+The combined liklihood map demonstrated that by merging estimates from two arrays the system is able to perform more accurate localization.
 
-From a timing point of view, the micro-controller spends $12$ millisecond on sampling microphone data before sending the data to the computer for processing.
-Sending data through USB port takes another $18$ millisecond, and processing on computer takes around $50$ millisecond.
+From a timing point of view, the micro-controller spends $12$ millisecond on sampling the microphone data before sending it to a computer for processing.
+Sending the data through the USB port takes another $15$ millisecond, and processing on the computer takes around $50$ millisecond.
 Therefore, the totoal time lag between sound source and localization is around $80$ millisecond.

### 2015-05-22 reports/paper/introduction.tex
spell check

@@
 
-Accuracte indoor locazation allows creation of noval applications with surrounding awareness that uses position and movment information as input.
+Accurate indoor localization allows creation of novel applications with surrounding awareness that uses position and movement information as input.
 One application is to allow users to draw with music without physically touching the computer.
 Another example is to build AI games with physical pieces such as toy car racing where the computer controls some them.
-In this work, we aim to build such a source localization system that is portable, inexpensive, yet resonably accuracte for localization in a small area.
+In this work, we aim to build such a source localization system that is portable, inexpensive, yet reasonably accurate for localization in a small area.
 
@@
 In this paper, we first discussed in Section II some prior relevant research and approaches in sound localization.
-In Section III, we evaluated different array achitectures and their impact on localization accuracy.
+In Section III, we evaluated different array architectures and their impact on localization accuracy.
 We demonstrated that a two array system outperforms a single array system of similar physical dimension.

### 2015-05-22 reports/paper/prior.tex
spell check

@@
 Cross-correlation has also been used to measure signal arrival time differences\cite{tdoa:mit2, tdoa:micloc, tdoa:3}.
-Cross-correlation with prefilterings is known as \emph{generalized cross correlation (GCC)}.
-Different prefilterings have been investigated to improve arrival time difference estimation~\cite{tdoa:gcc1,tdoa:gcc2,tdoa:gcc3}.
+Cross-correlation with prefiltering is known as \emph{generalized cross correlation (GCC)}.
+Different prefiltering approaches have been investigated to improve arrival time difference estimation~\cite{tdoa:gcc1,tdoa:gcc2,tdoa:gcc3}.
 
@@
 \item[GCC] $W(\omega) = 1$. No prefiltering is done. This is unfiltered normal cross correlation.
-\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divideded by its magnitude. Only phase information contributes to delay estimation.
-\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. part of magnitude information is included in delay estimation.
+\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divided by its magnitude. Only phase information contributes to delay estimation.
+\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. Part of magnitude information is included in delay estimation.
 \end{description}

### 2015-05-22 reports/paper/simulation.tex
spell check

@@
 The thickness of the uncertainty region is not uniform around the hyperbola, the farther away the point is, the larger the uncertainty region becomes.
-This indicates for the the same delta distance movement it will generate smaller difference of distance change when the source is farther away from the array.
-The size of the uncertianty region is also angle dependent: points closer to the line connecting microphones have larger region compared to points close to the line bisecting microphones.
+This indicates for the same delta distance movement it will generate smaller difference of distance change when the source is farther away from the array.
+The size of the uncertainty region is also angle dependent: points closer to the line connecting microphones have larger region compared to points close to the line bisecting microphones.
 

### 2015-05-22 reports/paper/system.tex
spell check

@@
 The theoretical TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$.
-Then the liklihood map $L(x,y)$ can be built by superposing the liklihood from each microphone pair:
+Then the likelihood map $L(x,y)$ can be built by superposing the likelihood from each microphone pair:
 \begin{eqnarray*}
@@
 
-Liklihood maps from two arrays can be combined into the final liklihood map:
+Likelihood maps from two arrays can be combined into the final likelihood map:
 \begin{equation}\label{eq:combine_l}
@@
 \end{equation}
-, where $L_1(x,y)$ and $L_2(x,y)$ represent the liklihood map from array $1$ and array $2$.
+, where $L_1(x,y)$ and $L_2(x,y)$ represent the likelihood map from array $1$ and array $2$.
 
@@
   \end{subfigure}
-  \caption{Liklihood maps for localizating point at $(0.0,-0.3)$ m}
+  \caption{Likelihood maps for localizing point at $(0.0,-0.3)$ m}
   \label{fig:liklihood}
@@
 To see the effect of accuracy improvement using multiple arrays, fig~\ref{fig:liklihood} shows a real life localization where the source is placed at $(0$ cm$,-30$ cm$)$.
-It shows the individual liklihood map from each array and also the combined liklihood map according to equation~\ref{eq:combine_l}.
+It shows the individual likelihood map from each array and also the combined likelihood map according to equation~\ref{eq:combine_l}.
 Individual arrays give accurate angle estimate, but have high uncertainty in distance estimate.
-The combined liklihood map demonstrated that by merging estimates from two arrays the system is able to perform more accurate localization.
+The combined likelihood map demonstrated that by merging estimates from two arrays the system is able to perform more accurate localization.
 
@@
 Sending the data through the USB port takes another $15$ millisecond, and processing on the computer takes around $50$ millisecond.
-Therefore, the totoal time lag between sound source and localization is around $80$ millisecond.
+Therefore, the total time lag between sound source and localization is around $80$ millisecond.

### 2015-05-22 reports/paper/experiment.tex
update

@@
 \section{Experiment}
+We conducted two sets of experiments to evaluate the system's localization accuracy: one on point localization and the other on movement tracking.
+
 \subsection{Setup}
@@
 
-To test localization accuracy, an one meter by one meter grid was set up and the arrays are placed at the top left and top right corners of the grid.
+An one meter by one meter grid was set up where the arrays were placed at the top left and top right corners of the grid.
 Fig~\ref{fig:setup_point} shows a picture of the setup.
-A total of $32$ testing locations are chosen uniformly in this.
+A total of $32$ testing locations were chosen uniformly in this grid.
+We reported the error as the average distance between our placement of the audio source and the location estimated from the arrays.
 
 \subsubsection{Movement tracking}
-To test how well the arrays track movement, we mounted a rotating disk $40$ centimeter in diameter onto the grid at $(x=0$ cm$, y=-30$ cm$)$.
-Fig~\ref{fig:setup_circle} shows a picture of the sestup.
-A sound source is placed on the edge of the rotating disk and the arrays localizes the sound source as it rotates in a circle.
-In this experiment, we tested how accuracy changes with:
+To test how well the system tracks movement, we mounted a rotating disk $40$ centimeters in diameter onto the grid at $(x=0$ cm$, y=-30$ cm$)$.
+Fig~\ref{fig:setup_circle} shows a picture of the setup.
+A sound source is placed on the edge of the rotating disk and the arrays track the sound source as it rotates in a circle.
+In this experiment, we evaluated how accuracy changes with:
 \begin{itemize}
-\item different window sizes
-\item different audio sources
-\item different movement tracking filters
-\item different movement speeds
+\item window sizes
+\item audio sources
+\item movement tracking filters
+\item movement speeds
 \end{itemize}
@@
 To test how different sound sources impact localization quality, we conducted the experiments with three different sound sources:
-
 \begin{description}[\IEEEsetlabelwidth{long a label}\IEEEusemathlabelsep]
@@
 
-\item[Music A] A randomly picked music that has normal audio amplitude throughout experiment period. ``Honest Eyes'' by Black Tide was the music used.
+\item[Music A] A randomly picked music that has normal audio amplitude throughout the experiment period. \emph{Honest Eyes} by \emph{Black Tide} was the music used.
 
-\item[Music B] A randomly picked music with intermittent low amplitude sections. ``Canon'' was the music used.
+\item[Music B] A randomly picked music with intermittent low amplitude sections. \emph{Canon} was the music used.
 
@@
 
-For each experiment conducted, two different movement filters were tests:
+For each experiment conducted, two different movement filters were evaluated:
 \begin{description}[\IEEEsetlabelwidth{Very long label}\IEEEusemathlabelsep]
-\item[Averaging filter] localization for past $0.5$ seconds are averaged and outputed as current estimate.
-\item[Kalman filter] A 2nd order Kalman filtering is used.
+\item[Averaging filter] localization for past $0.5$ seconds were averaged and outputed as current estimate.
+\item[Kalman filter] A 2nd order Kalman filtering was used.
 \end{description}
+
+In the movement tracking experiments described above, we have the ground truth location of the circle, but not the exact location of the audio source at each moment during the movement.
+Therefore, the error is reported as the distance between the localized point to its closest point on the ground truth circle.
 
@@
 \subsubsection{Point localization}
-To test how accuracy varies with window size, the algorithm is fed with microphone data of different window size.
-Fig~\ref{fig:accuracy_vs_window} shows how accuracy changes with the window size for GCC, GCC\_PHAT and GCC\_PHAT\_SQRT.
+To test how accuracy varies with window size, the algorithm is fed with microphone data of different lengths, and the result is shown in fig~\ref{fig:accuracy_vs_window}.
 The error decreases as window size increases and plateaus after the window size exceeds around $10$ millisecond.
-The lowest error achieved is $2.53$ centimeters.
-It is achieved when window size is set to 12 millisecond and GCC\_PHAT is used for TDOA estimation.
+The lowest error achieved is $2.53$ centimeters, which occurred when the window size is set to 12 millisecond and GCC\_PHAT is used for TDOA estimation.
 
-Although accuracy improves with the window size, calculation time also increases with it.
+Although accuracy improves with the window size, computation time also increases with it.
 The part of calculation that depends on the window size is using cross correlation for TDOA estimation.
 Cross correlation can be calculated with Fast Fourier Transform (FFT) and the runtime is of order $O(N\log N)$.
-We measured how the computation time varies with window size and Figure~\ref{fig:speed_vs_window} shows the result.
+We measured how the computation time varied with window size and Figure~\ref{fig:speed_vs_window} shows the result.
 The runtime increases approximately linearly in the window size region of interest.
 
-We also calculated the localization error for each tested point in the region.
+We also calculated the localization error for each tested point in the grid.
 Figure~\ref{fig:error_distribution} shows the error distribution inside the grid.
-The error is below $3$ cm for most areas inside the region.
+The error is below $3$ cm for most areas inside the grid.
 There is one error spike in the mid-left region and we contribute this to audio source misplacecment becuase the error is fairly low and consistent around that spike.
@@
 When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate TDOA.
-The localization is noisy.
-As window size increases, the localization converges to the shape of ground truth circle.
+Which makes localization noisy.
+As window size increases, localization converges to the shape of the ground truth circle.
 Fig~\ref{fig:err_win_circle} shows how the error changes with window size.
@@
 Fig~\ref{fig:circle_wnn} to \ref{fig:circle_musicbn} shows results for experiments at normal speed, and Fig~\ref{fig:circle_wnf} to~\ref{fig:circle_musicbf} shows results at fast speed.
-By comparing localization error for each audio source between normal movement speed and fast movement speed, we find the localization error does not depend on how fast the sound sorce is moving.
+By comparing localization error for each audio source between normal movement speed and fast movement speed, we find that localization error does not depend on how fast the sound sorce is moving.
 For example fig~\ref{fig:circle_musican} and fig~\ref{fig:circle_musicaf} shows that localization error is $1.289$ cm at normal movement speed and $1.291$ cm at fast movement speed.

### 2015-05-22 reports/paper/experiment.tex
paper update. emailed

@@
   \end{subfigure}
-  \caption{Setup for localization accuracy testing}
+  \caption{Setup for point localization evaluation}
   \label{fig:setup_point}
@@
 \begin{description}[\IEEEsetlabelwidth{Very long label}\IEEEusemathlabelsep]
-\item[Averaging filter] localization for past $0.5$ seconds were averaged and outputed as current estimate.
-\item[Kalman filter] A 2nd order Kalman filtering was used.
+\item[Averaging filter] localization for past $0.5$ seconds were averaged and outputted as current estimate.
+\item[Kalman filter] A 2nd order Kalman filter was used.
 \end{description}
@@
 The error is below $3$ cm for most areas inside the grid.
-There is one error spike in the mid-left region and we contribute this to audio source misplacecment becuase the error is fairly low and consistent around that spike.
+There is one error spike in the mid-left region and we contribute this to audio source misplacement because the error is fairly low and consistent around that spike.
 
@@
 \includegraphics[width=0.5\textwidth]{average_error_window_size.eps}
-\caption{accuracy versus window size}
+\caption{Localization error versus window size}
 \label{fig:accuracy_vs_window}
@@
 \includegraphics[width=0.4\textwidth]{calculation_time.eps}
-\caption{speed versus window size}
+\caption{Computation time versus window size}
 \label{fig:speed_vs_window}
@@
   \end{subfigure}
-  \caption{Setup for circle movement localization}
+  \caption{Setup for movement tracking evaluation}
   \label{fig:setup_circle}
@@
 Fig~\ref{fig:wn} gives an intuitive representation of how accuracy changes with window size.
-When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate TDOA.
-Which makes localization noisy.
+When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate TDOA, which results in noisy localization.
 As window size increases, localization converges to the shape of the ground truth circle.
@@
 Fig~\ref{fig:circle_wnn} to \ref{fig:circle_musicbn} shows results for experiments at normal speed, and Fig~\ref{fig:circle_wnf} to~\ref{fig:circle_musicbf} shows results at fast speed.
-By comparing localization error for each audio source between normal movement speed and fast movement speed, we find that localization error does not depend on how fast the sound sorce is moving.
+By comparing localization error for each audio source between normal movement speed and fast movement speed, we find that localization error does not depend on how fast the sound source is moving.
 For example fig~\ref{fig:circle_musican} and fig~\ref{fig:circle_musicaf} shows that localization error is $1.289$ cm at normal movement speed and $1.291$ cm at fast movement speed.

### 2015-05-22 reports/paper/introduction.tex
paper update. emailed

@@
 \section{Introduction}
-
-%Localization has been a widely researched topic in both academia and industry due to its wide range of applications.
-Companies and individuals have bennefited from technologies such as GPS, ultrasound, and blah.
-Since xxx  GPS technology has been invented and built into cars for navigation which provided great convenience for drivers.
-Ultrasound when used by sonar, help greatly helped underwater mapping.
-However, when it comes to localizationon a smaller scale such as the inside of a room, GPS would not be of much help because it's accuracy range is $1.4$ m~\cite{intro:gps}.
 
@@
 One application is to allow users to draw with music without physically touching the computer.
-Another example is to build AI games with physical pieces such as toy car racing where the computer controls some them.
+Another example is to build AI games with physical pieces such as toy car racing where the computer controls some of the toy cars.
 In this work, we aim to build such a source localization system that is portable, inexpensive, yet reasonably accurate for localization in a small area.
@@
 While accuracy in this range is good for many applications including driving navigation and vehicle tracking, it does not provide enough precision for local movement tracking.
-Ultrasound based indoor localization approaches on the other hand, has achieved sub-centimeter accuracy~\cite{intro:ultra}.
+Ultrasound based indoor localization approaches on the other hand, have achieved sub-centimeter accuracy~\cite{intro:ultra}.
 However, ultrasound systems require the use of expensive transducers.
@@
 
-In this project, we have built a localization system with reasonable high precision for small area using microphone arrays that localize typical audio source.
+In this project, we have built a localization system with reasonably high precision for small area using microphone arrays that localize typical audio sources.
 Our system is built with inexpensive electret microphones mounted on portable frames.
@@
 In Section IV, we presented the chosen architecture along with hardware details.
-Finally, experiment details and results are presented in Section V.
+Finally, experiment details and results were presented in Section V.

### 2015-05-22 reports/paper/prior.tex
paper update. emailed

@@
 
-\cite{extended:tusi} and \cite{ltm:pham} investigated using max value from cross-correlation as a similarity measure to localize user tap on interactive surfaces.
+\cite{extended:tusi} and \cite{ltm:pham} investigated using max value from cross-correlation as a similarity measure to localize user taps on interactive surfaces.
 \cite{ltm:lpc} used L2 distance in the Linear Predictive Coding coefficient space as a similarity measure to localize taps on surfaces.
@@
 \end{eqnarray}
-, where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transform of $x_1(t)$ and $x_2(t)$.
+, where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transforms of $x_1(t)$ and $x_2(t)$.
 $W(\omega)$ provides a way to prefilter signals passed to the cross correlation estimator.

### 2015-05-22 reports/paper/simulation.tex
paper update. emailed

@@
     \includegraphics[width=\textwidth]{sim/sim_2_1}
-    \caption{source at $(r=50$ cm,$\theta = 0$ degrees$)$}
+    \caption{source at $(r=50$ cm, $\theta = 0$ degrees$)$}
   \end{subfigure}
@@
     \includegraphics[width=\textwidth]{sim/sim_2_2}
-    \caption{source at $(r=50$ cm,$\theta = 45$ degrees$)$}
+    \caption{source at $(r=50$ cm, $\theta = 45$ degrees$)$}
   \end{subfigure}
@@
     \includegraphics[width=\textwidth]{sim/sim_2_3}
-    \caption{source at $(r=50$ cm,$\theta = 90$ degrees$)$}
+    \caption{source at $(r=50$ cm, $\theta = 90$ degrees$)$}
   \end{subfigure}
-  \caption{Uncertainty region}
+  \caption{Uncertainty region. Yellow dots represent microphones' locations and the white dot represents the location of the source.}
   \label{fig:sim_2_5}
@@
 \]
-, then fig~\ref{fig:sim_2_5} shows the region $R$ where all points have difference of distance to two microphones close to $a$ within $1$ cm:
+, then fig~\ref{fig:sim_2_5} shows the region $R$ where all points have difference of distance to two microphones close to $2a$ within $1$ cm:
 \[
@@
 \end{eqnarray}
-To see how difference of distance changes with respect to distance, we can expand the equation and find partial differential $\frac{\partial a}{\partial x}$:
+To see how the difference of distance changes with respect to source location, we can expand the equation and find the partial differential $\frac{\partial a}{\partial x}$:
 \begin{eqnarray}\label{eqn:derivative}
@@
   \end{subfigure}
-  \caption{$20$cm equilateral triangle array. Source is $20$cm away from the array}
+  \caption{Uncertainty region. Microphones are at the vertices of a $20$cm equilateral triangle. The source is $20$cm away from the array.}
   \label{fig:sim_3_2}
@@
   \end{subfigure}
-  \caption{$20$ cm equilateral triangle array. Source is $80$ cm away from the array}
+  \caption{Uncertainty region. Microphones are at the vertices of a $20$cm equilateral triangle. The source is $80$cm away from the array.}
   \label{fig:sim_3_8}
@@
 The smaller the intersection region, the better the localization accuracy.
-To see how accuracy changes with array placement and sound source location, three microphones are placed at the three vertices of an $20$ cm equilateral triangle.
+To see how accuracy changes with array placement and sound source location, three microphones are placed at the three vertices of a $20$ cm equilateral triangle.
 An audio source is placed at $20$ cm away from the center of the array.
@@
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_center}
-    \caption{$3$ microphones placed at vertices of an $20$cm equilateral triangle. Average error is $18.6$ cm}
+    \caption{$3$ microphones are placed at vertices of a $20$cm equilateral triangle. The average error is $18.6$ cm}
     \label{fig:sim_hm_3}
@@
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_p1}
-    \caption{another microphone added at origin. Average error is $17.1$ cm}
+    \caption{Another microphone is added at origin. The average error is $17.1$ cm}
     \label{fig:sim_hm_3_p1}
@@
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_2x}
-    \caption{$3$ microphones placed at vertices of an $40$cm equilateral triangle. Average error is $10.04$ cm}
+    \caption{$3$ microphones are placed at vertices of a $40$cm equilateral triangle. The average error is $10.04$ cm}
     \label{fig:sim_hm_3_2x}
@@
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_line}
-    \caption{$3$ microphones placed in a line. Total length is $20$ cm. Average error is $55.05$ cm}
+    \caption{$3$ microphones are placed in a line, $10$ cm apart from each other. The average error is $55.05$ cm}
     \label{fig:sim_hm_3_line}
@@
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_4corner}
-    \caption{$4$ microphones placed at $4$ corners of the region. Average error is $0.05$ cm}
+    \caption{$4$ microphones are placed at $4$ corners of the grid. The average error is $0.05$ cm}
     \label{fig:sim_hm_4}
@@
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_2a}
-    \caption{two $3$ microphone arrays placed $1$ meter apart. Average error is $2.60$ cm}
+    \caption{Two $3$ microphone arrays are placed $1$ meter apart. The average error is $2.60$ cm}
     \label{fig:sim_hm_2_array}
   \end{subfigure}
-  \caption{Accuracy for different array configurations}
+  \caption{Error heatmap for different array configurations}
   \label{fig:sim_hm}
@@
 
-Each microphone pair generates a hyperbolic region, and the intersection area of these regions is a measure of the localization accuracy.
+Each microphone pair generates a hyperbolic region, and the source location is in the intersection of these regions.
+The area of the intersection region is a measure of the localization accuracy: the smaller the area, the more certain we are about the source location.
 To evaluate an array's accuracy in a region, we can place sound source at predetermined grid points in the region and look at the intersection area for each tested point in the grid.
-The center location of intersection region can be used as the localization estimate to calculate localization error.
+The center location of the intersection region can be used as the localization estimate to calculate localization error.
 Results for a few different microphone array configurations are presented in fig~\ref{fig:sim_hm}.
@@
 However, for regions along the line connecting any two microphones, the accuracy drops significantly.
-Average error across the region is $18.6$ cm.
+The average error across the region is $18.6$ cm.
 
@@
 Addition of the new microphone only slightly improved the accuracy around the array region.
-Average error dropped from $18.6$ cm to $17.1$ cm.
+The average error dropped from $18.6$ cm to $17.1$ cm.
 Regions near lines connecting microphones still have significantly larger uncertainty region.
@@
 The overall uncertainty area decreased across the region.
-Average error improved to $10.04$ cm.
+The average error improved to $10.04$ cm.
 

### 2015-05-22 reports/paper/system.tex
paper update. emailed

@@
 The micro-controller used in this project is \emph{teensy 3.1}, which has $64$k RAM memory and the ADC is capable of sampling at $500$kHz~\cite{tdoa:micloc, sys:teensy}.
-In this project, the micro-controller collects microphone data on all three channels for $12$ millisecond and then sends the recorded data to a computer through the USB port for localization.
+In this project, the micro-controller collects microphone data on all three channels for $12$ milliseconds and then sends the recorded data to a computer through the USB port for localization.
 
@@
   \end{subfigure}
-  \caption{Localization arrays}
+  \caption{Localization system setup}
   \label{fig:setup_array}
@@
 
-To handle the uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation output~\ref{eq:gcc2} as a measure of the likelihood of different arrival time differences.
+To handle the uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation output(equation~\ref{eq:gcc2}) as a measure of the likelihood of different arrival time differences.
 For each microphone array, we build a heatmap for the region.
@@
   \end{subfigure}
-  \caption{Likelihood maps for localizing point at $(0.0,-0.3)$ m}
+  \caption{Likelihood maps for localization. The source is placed at $(0.0,-0.3)$ m}
   \label{fig:liklihood}
@@
 
-From a timing point of view, the micro-controller spends $12$ millisecond on sampling the microphone data before sending it to a computer for processing.
-Sending the data through the USB port takes another $15$ millisecond, and processing on the computer takes around $50$ millisecond.
-Therefore, the total time lag between sound source and localization is around $80$ millisecond.
+From a timing point of view, the micro-controller spends $12$ milliseconds on sampling the microphone data before sending it to a computer for processing.
+Sending the data through the USB port takes another $15$ milliseconds, and processing on the computer takes around $50$ milliseconds.
+Therefore, the total time lag between sound source and localization is around $80$ milliseconds.

### 2015-05-29 reports/paper/conclusion.tex
converted to thesis format

\section{Conclusion}
In this paper, we proposed and built an inexpensive, portable yet reasonably accurate sound localization system with two microphone arrays. We analyzed different array architectures and demonstrated that a two array architecture achieves better accuracy compared to a single array system of similar size. By using cross-correlation output as the likelihood for different arrival time differences, each array generates a likelihood map for each possible location in the area. We demonstrated that by merging likelihood maps from the two arrays, we were able to achieve good localization accuracy (less than $3$ cm average error) in a local one meter by one meter region.

This system can be used in HCI applications that incorporate sound position and movement information, such as drawing with music, or can be used to create AI games with physical pieces where the computer controls some of the pieces. This system can also be used to in Augmented Reality (AR) applications that allow user computer interaction with music tags.

For future directions, this system can be extended with another microphone on each array to perform 3D localization. Another possible direction is to investigate multi-source localization using digitally tagged audio sources.

### 2015-05-29 reports/paper/experiment.tex
converted to thesis format

@@
-\section{Experiment}
+\chapter{Experiment}
 We conducted two sets of experiments to evaluate the system's localization accuracy: one on point localization and the other on movement tracking.
 
-\subsection{Setup}
-\subsubsection{Point localization}
+\section{Setup}
+\subsection{Point localization}
 
@@
 
-\subsubsection{Movement tracking}
+\subsection{Movement tracking}
 To test how well the system tracks movement, we mounted a rotating disk $40$ centimeters in diameter onto the grid at $(x=0$ cm$, y=-30$ cm$)$.
@@
 To test how different sound sources impact localization quality, we conducted the experiments with three different sound sources:
-\begin{description}[\IEEEsetlabelwidth{long a label}\IEEEusemathlabelsep]
+\begin{description}%[\IEEEsetlabelwidth{long a label}\IEEEusemathlabelsep]
 
@@
 For each experiment conducted, two different movement filters were evaluated:
-\begin{description}[\IEEEsetlabelwidth{Very long label}\IEEEusemathlabelsep]
+\begin{description}%[\IEEEsetlabelwidth{Very long label}\IEEEusemathlabelsep]
 \item[Averaging filter] localization for past $0.5$ seconds were averaged and outputted as current estimate.
@@
 
-\subsection{Results}
-\subsubsection{Point localization}
+\section{Results}
+\subsection{Point localization}
 To test how accuracy varies with window size, the algorithm is fed with microphone data of different lengths, and the result is shown in fig~\ref{fig:accuracy_vs_window}.
@@
 
-\subsubsection{Movement tracking}
+\subsection{Movement tracking}
 

### 2015-05-29 reports/paper/prior.tex
converted to thesis format

@@
-\section{Background}
+\chapter{Background}
 Acoustic localization has been researched extensively in the literature.
@@
 
-\subsection{LTM}
+\section{LTM}
 In LTM based approaches, acoustic templates acquired from different locations are first stored in the system during a ``training'' phase.
@@
 
-\subsection{TDOA}
+\section{TDOA}
 TDOA approaches exploit the difference of arrival time between the acoustic source and two fixed microphones on the plane.
@@
 We focused on three ways of prefiltering the signal:
-\begin{description}[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]
+\begin{description}%[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]
 \item[GCC] $W(\omega) = 1$. No prefiltering is done. This is unfiltered normal cross correlation.

### 2015-05-29 reports/paper/simulation.tex
converted to thesis format

@@
-\section{Array Architecture}
+\chapter{Array Architecture}
 As was mentioned in the previous section, points with the same TDOA to two fixed locations form a hyperbola on a 2D plane.

### 2015-05-29 reports/paper/conclusion.tex
update thesis

@@
-\section{Conclusion}
+\chapter{Conclusion}
 In this paper, we proposed and built an inexpensive, portable yet reasonably accurate sound localization system with two microphone arrays.

### 2015-05-29 reports/paper/prior.tex
update thesis

@@
 Acoustic localization has been researched extensively in the literature.
-Localization techniques can be broadly categorized into Location Template Matching (LTM) and Time Difference of Arrival (TDOA) based approaches.
+Localization techniques can be broadly categorized into Location Interaural Level Difference (ILD), Template Matching (LTM), and Time Difference of Arrival (TDOA) based approaches.
+
+\section{ILD}
+a
 
@@
 \end{eqnarray}
-, where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transforms of $x_1(t)$ and $x_2(t)$.
+where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transforms of $x_1(t)$ and $x_2(t)$.
 $W(\omega)$ provides a way to prefilter signals passed to the cross correlation estimator.

### 2015-05-29 reports/paper/system.tex
converted to thesis format
update thesis

@@
-\section{System Setup}
+\chapter{System Setup}
 The end system has two arrays, each with three microphones mounted on the vertices of a $20$ cm equilateral triangle.
@@
 \end{eqnarray*}
-,where $R_{m_1,m_2}(\tau)$,$R_{m_1,m_3}(\tau)$, and $R_{m_2,m_3}(\tau)$ denote GCC output from microphone pairs $m_1m_2,m_1m_3,$ and $m_2m_3$.
+where $R_{m_1,m_2}(\tau)$,$R_{m_1,m_3}(\tau)$, and $R_{m_2,m_3}(\tau)$ denote GCC output from microphone pairs $m_1m_2,m_1m_3,$ and $m_2m_3$.
 
@@
 \end{equation}
-, where $L_1(x,y)$ and $L_2(x,y)$ represent the likelihood map from array $1$ and array $2$.
+where $L_1(x,y)$ and $L_2(x,y)$ represent the likelihood map from array $1$ and array $2$.
 

### 2015-06-01 reports/paper/conclusion.tex
paper update

@@
 \chapter{Conclusion}
-In this paper, we proposed and built an inexpensive, portable yet reasonably accurate sound localization system with two microphone arrays.
+In this paper, we described and built an inexpensive, portable yet reasonably accurate sound localization system with two microphone arrays.
 We analyzed different array architectures and demonstrated that a two array architecture achieves better accuracy compared to a single array system of similar size.

### 2015-06-01 reports/paper/introduction.tex
converted to thesis format
paper update

@@
-\section{Introduction}
+%paper \section{Introduction}
+\chapter{Introduction}
 
@@
 
-In this paper, we first discussed in Section II some prior relevant research and approaches in sound localization.
-In Section III, we evaluated different array architectures and their impact on localization accuracy.
+In this paper, we first discussed in Chapter II some prior relevant research and approaches in sound localization.
+In Chapter III, we evaluated different array architectures and their impact on localization accuracy.
 We demonstrated that a two array system outperforms a single array system of similar physical dimension.
-In Section IV, we presented the chosen architecture along with hardware details.
-Finally, experiment details and results were presented in Section V.
+In Chapter IV, we presented the chosen architecture along with hardware details.
+Finally, experiment details and results were presented in Chapter V.

### 2015-06-01 reports/paper/prior.tex
paper update

@@
 Acoustic localization has been researched extensively in the literature.
-Localization techniques can be broadly categorized into Location Interaural Level Difference (ILD), Template Matching (LTM), and Time Difference of Arrival (TDOA) based approaches.
+Localization techniques can be broadly categorized into Interaural Level Difference (ILD), Location Template Matching (LTM), and Time Difference of Arrival (TDOA) based approaches.
 
 \section{ILD}
-a
+ILD techniques rely on the observation that signal intensity decays as the distance to the microhone increases.
+A microphone closer to the signal source would receive the signal with higher intensity than a microphone farther away.
+With multiple microphones, it is possible to infer the source location by comparing the signal intensity received at different microphones.
+Human's auditory system has used ILD cues to infer source direction~\cite{ild:2}, and this technique is most effective to localize high frquency sources, because they don't diffract effectively around the listener's head and produce a significant intensity difference.
+
+Let $s(t)$ denote the audio signal. The received signal $\bar s(t) $ at microphone $i$ can be modeled as:
+\[
+\bar s_i(t) = \frac{s(t)}{d_i} + \delta_i(t)
+\]
+where $d_i$ is the distance between the audio source and microphone $i$, and $\delta_i(t)$ represents Gaussian noise.
+The energy received at microphone $i$ is:
+\begin{eqnarray}
+E_i & = & \int \bar s_i(t)^2 dt \\
+\label{eqn:ild1}
+& = & \frac{1}{d_i^2} E_s + E_{\sigma_i}
+\end{eqnarray}
+It shows that the energy declays with the square of the distance.
+With two microphones, not considering the noise $\delta_i(t)$, energy received at both microphones has to satisfy equation~\ref{eqn:ild1}:
+\begin{eqnarray}
+E_1d_1^2 &=& E_2d_2^2\\
+\label{eqn:ild}
+\frac{E_1}{E_2} &=& \frac{d_2^2}{d_1^2}
+\end{eqnarray}
+It can be shown that, on a 2D plan, points satisfy equation~\ref{eqn:ild} form a circle when $E_1 \ne E_2$ and form a line when $E_1 = E_2$~\cite{ild:1}.
+With two microphones all points on this curve generate the same energy ratio and can not be distinguised from each other.
+Multiple approaches have been investigated to eliminate this ambiguity.
+\cite{ild:1} employed multiple microphones and used the intersection of circles from each microphone pair to estimate the source location.
+\cite{ild:2} combined ILD with Interaural Time Difference to estimate source direction (i.e azimuth).
+
+ILD approach relies on the accurate measurement of the received energy ratio between microphone pairs.
+Any obstacle object between the sound source and any microphone would produce a significant distortion in the measured energy ratio.
+We find this approach to be too restrictive for our system, since in an interactive system we can not control whether or not the user places any obstacle between the sound source and the microphones.
 
@@
 \item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divided by its magnitude. Only phase information contributes to delay estimation.
-\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. Part of magnitude information is included in delay estimation.
+\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$.
+This is somewhere between GCC and GCC\_PHAT.
+Part of magnitude information is included in arrival time difference estimation.
 \end{description}
+
+To see the reasoning behind different prefitering approaches, we can seperate the magnitude and phase part of $X_1(\omega)$ and $X_2(\omega)$ in Equation~\ref{eq:gcc2}:
+\begin{eqnarray}
+\label{eqn:phat1}
+R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) |X_1(\omega)||X_2(\omega)| e^{j(\omega\tau - (\angle{X_2(\omega)} - \angle{X_1(\omega)}) } d\omega\\
+&=& \int_{-\infty}^\infty W(w)|X_1(\omega)| |X_2(\omega)| \cos(\omega\tau - (\angle X_2(\omega) - \angle X_1(\omega))) d\omega
+\end{eqnarray}
+We look at the real part of equation~\ref{eqn:phat1} only since both $x_1(t)$ and $x_2(t)$ are real valued signals.
+When $\dot \tau$ is the true arrive time difference between $x_1(t)$ and $x_2(t)$, $\omega \dot \tau - (\angle X_2(\omega) - \angle X_1(\omega)) = 0$, and $\cos(\omega \dot \tau - (\angle X_2(\omega) - \angle X_1(\omega))) = 1$.
+Therefore, $\cos(\omega\tau - (\angle X_2(\omega) - \angle X_1(\omega)))$ can be seen as a measure of the phase error, and $W(w)|X_1(\omega)||X_2(\omega)|$ describes how the error should be weighted at each frequency.
+The TDOA estimator essentially sums the weighted phase error at each frequncy.
+
+Without any prefitering (i.e $W(\omega)=1$), the estimator weighs the phase error at each frequency by the magnitude of the signal at that frequency.
+In this weighting scheme, phase error at frequencies with higher magnitudes are penalized more compared to frequencies with a lower magnitude.
+This weighting is approporiate if there is only one source present, since freqeuncies with higher intensities have higher Signal to Noise Ratio (SNR).
+It makes sense to place higher weights at frequencies with higher SNR.
+
+However, with multiple sources, the source with the highest volume will dominate the phase error estimation, but there is no particular reson to assign higher weights to signals with higher volume.
+All sources should contribute equally in phase error estimation.
+In GCC\_PHAT,  $W(\omega)$ is set to $\frac{1}{|X_1(\omega)||X_2(\omega)|}$.
+In effect it weighs phase error uniformly across freqiency.

### 2015-06-01 reports/paper/simulation.tex
update thesis
paper update

@@
   \centering
-  \begin{subfigure}[]{.23\textwidth}
+  \begin{subfigure}[]{.46\textwidth}
     \includegraphics[width=\textwidth]{sim/sim_2_1}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.23\textwidth}
+  \begin{subfigure}[]{.46\textwidth}
     \includegraphics[width=\textwidth]{sim/sim_2_2}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.23\textwidth}
-    \includegraphics[width=\textwidth]{sim/sim_2_3}
-    \caption{source at $(r=50$ cm, $\theta = 90$ degrees$)$}
-  \end{subfigure}
+%  \begin{subfigure}[]{.3\textwidth}
+%    \includegraphics[width=\textwidth]{sim/sim_2_3}
+%    \caption{source at $(r=50$ cm, $\theta = 90$ degrees$)$}
+%  \end{subfigure}
   \caption{Uncertainty region. Yellow dots represent microphones' locations and the white dot represents the location of the source.}
@@
 \]
-, then fig~\ref{fig:sim_2_5} shows the region $R$ where all points have difference of distance to two microphones close to $2a$ within $1$ cm:
+then fig~\ref{fig:sim_2_5} shows the region $R$ where all points have difference of distance to two microphones close to $2a$ within $1$ cm:
 \[
@@
   \centering
-  \begin{subfigure}[]{.3\textwidth}
+  \begin{subfigure}[]{.46\textwidth}
     \includegraphics[width=\textwidth]{sim/result_20cm_0_degrees}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.3\textwidth}
-    \includegraphics[width=\textwidth]{sim/result_20cm_45_degrees}
-    \caption{45 degrees}
-  \end{subfigure}
-  \begin{subfigure}[]{.3\textwidth}
-    \includegraphics[width=\textwidth]{sim/result_20cm_90_degrees}
-    \caption{90 degrees}
-  \end{subfigure}
-  \begin{subfigure}[]{.3\textwidth}
+%  \begin{subfigure}[]{.3\textwidth}
+%    \includegraphics[width=\textwidth]{sim/result_20cm_45_degrees}
+%    \caption{45 degrees}
+%  \end{subfigure}
+%  \begin{subfigure}[]{.3\textwidth}
+%    \includegraphics[width=\textwidth]{sim/result_20cm_90_degrees}
+%    \caption{90 degrees}
+%  \end{subfigure}
+  \begin{subfigure}[]{.46\textwidth}
     \includegraphics[width=\textwidth]{sim/result_20cm_135_degrees}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.3\textwidth}
-    \includegraphics[width=\textwidth]{sim/result_20cm_180_degrees}
-    \caption{180 degrees}
-  \end{subfigure}
+%  \begin{subfigure}[]{.3\textwidth}
+%    \includegraphics[width=\textwidth]{sim/result_20cm_180_degrees}
+%    \caption{180 degrees}
+%  \end{subfigure}
   \caption{Uncertainty region. Microphones are at the vertices of a $20$cm equilateral triangle. The source is $20$cm away from the array.}
@@
   \centering
-  \begin{subfigure}[]{.3\textwidth}
+  \begin{subfigure}[]{.46\textwidth}
     \includegraphics[width=\textwidth]{sim/result_80cm_0_degrees}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.3\textwidth}
-    \includegraphics[width=\textwidth]{sim/result_80cm_45_degrees}
-    \caption{45 degrees}
-  \end{subfigure}
-  \begin{subfigure}[]{.3\textwidth}
-    \includegraphics[width=\textwidth]{sim/result_80cm_90_degrees}
-    \caption{90 degrees}
-  \end{subfigure}
-  \begin{subfigure}[]{.3\textwidth}
+%  \begin{subfigure}[]{.3\textwidth}
+%    \includegraphics[width=\textwidth]{sim/result_80cm_45_degrees}
+%    \caption{45 degrees}
+%  \end{subfigure}
+%  \begin{subfigure}[]{.3\textwidth}
+%    \includegraphics[width=\textwidth]{sim/result_80cm_90_degrees}
+%    \caption{90 degrees}
+%  \end{subfigure}
+  \begin{subfigure}[]{.46\textwidth}
     \includegraphics[width=\textwidth]{sim/result_80cm_135_degrees}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.3\textwidth}
-    \includegraphics[width=\textwidth]{sim/result_80cm_180_degrees}
-    \caption{180 degrees}
-  \end{subfigure}
+%  \begin{subfigure}[]{.3\textwidth}
+%    \includegraphics[width=\textwidth]{sim/result_80cm_180_degrees}
+%    \caption{180 degrees}
+%  \end{subfigure}
   \caption{Uncertainty region. Microphones are at the vertices of a $20$cm equilateral triangle. The source is $80$cm away from the array.}
@@
 An audio source is placed at $20$ cm away from the center of the array.
-Fig~\ref{fig:sim_3_2} shows the intersection of regions for $5$ different placement of the sound source.
+Fig~\ref{fig:sim_3_2} shows the intersection of regions for $2$ different placement of the sound source.
 It can be seen that accuracy decreases when sound source becomes close to the line connecting any two microphones.

### 2015-06-01 reports/paper/prior.tex
paper update

@@
 
-To see the reasoning behind different prefitering approaches, we can seperate the magnitude and phase part of $X_1(\omega)$ and $X_2(\omega)$ in Equation~\ref{eq:gcc2}:
+To see the reasoning behind different prefitering approaches, we can seperate the magnitude part from the phase part of $X_1(\omega)$ and $X_2(\omega)$ in Equation~\ref{eq:gcc2}:
 \begin{eqnarray}
@@
 \end{eqnarray}
-We look at the real part of equation~\ref{eqn:phat1} only since both $x_1(t)$ and $x_2(t)$ are real valued signals.
+We can look at the real part of equation~\ref{eqn:phat1} only since both $x_1(t)$ and $x_2(t)$ are real valued signals.
 When $\dot \tau$ is the true arrive time difference between $x_1(t)$ and $x_2(t)$, $\omega \dot \tau - (\angle X_2(\omega) - \angle X_1(\omega)) = 0$, and $\cos(\omega \dot \tau - (\angle X_2(\omega) - \angle X_1(\omega))) = 1$.
@@
 In this weighting scheme, phase error at frequencies with higher magnitudes are penalized more compared to frequencies with a lower magnitude.
-This weighting is approporiate if there is only one source present, since freqeuncies with higher intensities have higher Signal to Noise Ratio (SNR).
+This weighting is approporiate if there is only one source present, since freqeuncies with higher magnitude have higher Signal to Noise Ratio (SNR).
 It makes sense to place higher weights at frequencies with higher SNR.
 
-However, with multiple sources, the source with the highest volume will dominate the phase error estimation, but there is no particular reson to assign higher weights to signals with higher volume.
+However, with multiple sources, the source with the highest volume will dominate the phase error estimation, but there is no particular reson to assign a higher weight to the signal with the highest volume.
 All sources should contribute equally in phase error estimation.
@@
 In effect it weighs phase error uniformly across freqiency.
+Since phase error at all frequencies are weighted equally, this technique will suffer from error accumulation if the source has a lot of low power regions in the frequency domain.
+
+In GCC\_PHAT\_SQRT, $W(\omega)$ is set to $\frac{1}{(|X_1(\omega)||X_2(\omega)|)^{0.5}}$.
+Phase error weighting at each frequency still depends on the signal strength at that frequency, but the dependency is much weaker than that in unfiltered GCC.
+On the other hand, this weighting scheme doesn't go to the other extreme of completely ignoring signal strength information as does in GCC\_PHAT.
+This approach represents a balance between unfilterred GCC and GCC\_PHAT.

### 2015-06-01 reports/paper/simulation.tex
paper update

@@
 \chapter{Array Architecture}
-As was mentioned in the previous section, points with the same TDOA to two fixed locations form a hyperbola on a 2D plane.
+As was mentioned in the previous chapter, points with the same TDOA to two fixed locations form a hyperbola on a 2D plane.
 However, in practical systems we can only measure TDOA up to a precision.
@@
 This $\epsilon$ represents accuracy on the measurement of difference of distances, and in practice it is related to sampling rate and estimation techniques.
-In this section we evaluate the impact of difference of distance estimation on localization accuracy.
+In this chapter we evaluate the impact of difference of distance estimation on localization accuracy.
 
@@
 \]
-then fig~\ref{fig:sim_2_5} shows the region $R$ where all points have difference of distance to two microphones close to $2a$ within $1$ cm:
+then the TDOA is $2a/340$ seconds. Assuming a sampling rate of $34$ KHz, fig~\ref{fig:sim_2_5} shows the region $R$ where all points have TDOA close to $2a/340$ s within one sample difference:
 \[
-R=\{\hat P: |(\hat P M_1 - \hat P M_2) - 2a|< 1 \mbox{ cm}\}
+R=\{\hat P: |(\hat P M_1 - \hat P M_2) - 2a|< \frac{1}{2} \mbox{ samples}\}
 \]
@@
   \end{subfigure}
-  \caption{Error heatmap for different array configurations}
+  \caption{Error heatmap for different array configurations. The heatmap scale is the intersection area measured in $cm^2$}.
   \label{fig:sim_hm}

### 2015-06-01 reports/paper/experiment.tex
paper update

@@
   \centering
-  \begin{subfigure}[]{.2\textwidth}
+  \begin{subfigure}[]{.46\textwidth}
     \includegraphics[width=\textwidth]{setup_1.JPG}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.2\textwidth}
+  \begin{subfigure}[]{.46\textwidth}
     \includegraphics[width=\textwidth]{setup_2.JPG}
@@
 
+To test the limit of the system and to evaluate the accuracy when the source moves outside of the one meter by one meter region, we measured the localization error by placing the source at $10$ locations along $y$ axis ranging from $(0$ cm$, -10$ cm$)$ to $(0$ cm$, -200$ cm$)$.
+The result is presented in fig~\ref{fig:error_along_y}.
+The localization error is within $3$ cm when the source is within $100$ cm from the arrays.
+The error increases to about $5$ cm when the source distance increases to $150$ cm and the error exceeds $10$ cm after the source distance reaches $200$ cm.
+
 \begin{figure}[]
+\centering
 \includegraphics[width=0.5\textwidth]{average_error_window_size.eps}
@@
 \centering
-\includegraphics[width=0.4\textwidth]{calculation_time.eps}
+\includegraphics[width=0.5\textwidth]{calculation_time.eps}
 \caption{Computation time versus window size}
@@
 \centering
-\includegraphics[width=0.3\textwidth]{error_distribution.eps}
+\includegraphics[width=0.5\textwidth]{error_distribution.eps}
 \caption{error distribution in the grid. Arrays are placed at $(-0.5$ m$, 0$ m$)$ and $(0.5$ m$, 0$ m$)$}
@@
 \centering
-  \begin{subfigure}[]{.2\textwidth}
-    \includegraphics[width=\textwidth]{setup_ring.JPG}
-  \end{subfigure}
-  \begin{subfigure}[]{.2\textwidth}
+\includegraphics[width=0.5\textwidth]{error_along_y_axis.eps}
+\caption{Localization error as the distance between the source and the microphone arrays increases. The source is placed on the $y$ axis.}
+\label{fig:error_along_y}
+\end{figure}
+
+\begin{figure}[]
+  \centering
+%  \begin{subfigure}[]{.2\textwidth}
+%    \includegraphics[width=\textwidth]{setup_ring.JPG}
+%  \end{subfigure}
+  \begin{subfigure}[]{.5\textwidth}
     \includegraphics[width=\textwidth]{setup_rotating_disk.JPG}
@@
 \centering
-\includegraphics[width=0.4\textwidth]{error_window_size_movement.eps}
+\includegraphics[width=0.5\textwidth]{error_window_size_movement.eps}
 \caption{Localization error versus window size}
@@
 Fig~\ref{fig:wn} gives an intuitive representation of how accuracy changes with window size.
-When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate TDOA, which results in noisy localization.
+When window size is small ($1.02$ millisecond), the audio does not contain enough information to reliably estimate TDOA, which results in noisy localization.
 As window size increases, localization converges to the shape of the ground truth circle.

### 2015-06-01 reports/paper/prior.tex
paper update

@@
 \section{ILD}
-ILD techniques rely on the observation that signal intensity decays as the distance to the microhone increases.
+ILD techniques rely on the observation that signal intensity decays as the distance to the microphone increases.
 A microphone closer to the signal source would receive the signal with higher intensity than a microphone farther away.
 With multiple microphones, it is possible to infer the source location by comparing the signal intensity received at different microphones.
-Human's auditory system has used ILD cues to infer source direction~\cite{ild:2}, and this technique is most effective to localize high frquency sources, because they don't diffract effectively around the listener's head and produce a significant intensity difference.
+Human's auditory system has used ILD cues to infer source direction~\cite{ild:2}, and this technique is most effective to localize high frequency sources, because they don't diffract effectively around the listener's head and produce a significant intensity difference.
 
@@
 \end{eqnarray}
-It shows that the energy declays with the square of the distance.
+It shows that the energy decays with the square of the distance.
 With two microphones, not considering the noise $\delta_i(t)$, energy received at both microphones has to satisfy equation~\ref{eqn:ild1}:
@@
 \end{eqnarray}
-It can be shown that, on a 2D plan, points satisfy equation~\ref{eqn:ild} form a circle when $E_1 \ne E_2$ and form a line when $E_1 = E_2$~\cite{ild:1}.
-With two microphones all points on this curve generate the same energy ratio and can not be distinguised from each other.
+It can be shown that, on a 2D plane, points satisfy equation~\ref{eqn:ild} form a circle when $E_1 \ne E_2$ and form a line when $E_1 = E_2$~\cite{ild:1}.
+With two microphones all points on this curve generate the same energy ratio and can not be distinguished from each other.
 Multiple approaches have been investigated to eliminate this ambiguity.
 \cite{ild:1} employed multiple microphones and used the intersection of circles from each microphone pair to estimate the source location.
-\cite{ild:2} combined ILD with Interaural Time Difference to estimate source direction (i.e azimuth).
+\cite{ild:2} combined ILD with Interaural Time Difference to estimate the source direction (i.e azimuth).
 
@@
 
-To see the reasoning behind different prefitering approaches, we can seperate the magnitude part from the phase part of $X_1(\omega)$ and $X_2(\omega)$ in Equation~\ref{eq:gcc2}:
+To see the reasoning behind different prefiltering approaches, we separate the magnitude part from the phase part of $X_1(\omega)$ and $X_2(\omega)$ in Equation~\ref{eq:gcc2}:
 \begin{eqnarray}
@@
 R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) |X_1(\omega)||X_2(\omega)| e^{j(\omega\tau - (\angle{X_2(\omega)} - \angle{X_1(\omega)}) } d\omega\\
-&=& \int_{-\infty}^\infty W(w)|X_1(\omega)| |X_2(\omega)| \cos(\omega\tau - (\angle X_2(\omega) - \angle X_1(\omega))) d\omega
+&=& \int_{-\infty}^\infty \underbrace{W(w)|X_1(\omega)| |X_2(\omega)|}_{\mbox{weighting}} \underbrace{\cos(\omega\tau - (\angle X_2(\omega) - \angle X_1(\omega)))}_{\mbox{phase error}} d\omega
 \end{eqnarray}
 We can look at the real part of equation~\ref{eqn:phat1} only since both $x_1(t)$ and $x_2(t)$ are real valued signals.
-When $\dot \tau$ is the true arrive time difference between $x_1(t)$ and $x_2(t)$, $\omega \dot \tau - (\angle X_2(\omega) - \angle X_1(\omega)) = 0$, and $\cos(\omega \dot \tau - (\angle X_2(\omega) - \angle X_1(\omega))) = 1$.
+When $\dot \tau$ is the true arrival time difference between $x_1(t)$ and $x_2(t)$, $\omega \dot \tau - (\angle X_2(\omega) - \angle X_1(\omega)) = 0$, and $\cos(\omega \dot \tau - (\angle X_2(\omega) - \angle X_1(\omega))) = 1$.
+When $\dot \tau$ differs from the true arrival time difference, $\cos(\omega \dot \tau - (\angle X_2(\omega) - \angle X_1(\omega))) < 1$.
 Therefore, $\cos(\omega\tau - (\angle X_2(\omega) - \angle X_1(\omega)))$ can be seen as a measure of the phase error, and $W(w)|X_1(\omega)||X_2(\omega)|$ describes how the error should be weighted at each frequency.
-The TDOA estimator essentially sums the weighted phase error at each frequncy.
+The TDOA estimator essentially sums the weighted phase error at each frequency.
 
-Without any prefitering (i.e $W(\omega)=1$), the estimator weighs the phase error at each frequency by the magnitude of the signal at that frequency.
+Without any prefiltering (i.e $W(\omega)=1$), the estimator weighs the phase error at each frequency by the magnitude of the signal at that frequency.
 In this weighting scheme, phase error at frequencies with higher magnitudes are penalized more compared to frequencies with a lower magnitude.
-This weighting is approporiate if there is only one source present, since freqeuncies with higher magnitude have higher Signal to Noise Ratio (SNR).
-It makes sense to place higher weights at frequencies with higher SNR.
+This weighting is appropriate if there is only one source present, since frequencies with higher magnitude have higher Signal to Noise Ratio (SNR).
+It makes sense to place higher weights at frequencies with higher SNR, since low SNR regions can be dominated by noise.
 
-However, with multiple sources, the source with the highest volume will dominate the phase error estimation, but there is no particular reson to assign a higher weight to the signal with the highest volume.
-All sources should contribute equally in phase error estimation.
+However, with multiple sources, the source with the highest magnitude will dominate the phase error estimation, but there is no particular reason to assign a higher weight to the source with the highest volume.
+All sources should contribute equally in the phase error estimation.
 In GCC\_PHAT,  $W(\omega)$ is set to $\frac{1}{|X_1(\omega)||X_2(\omega)|}$.
-In effect it weighs phase error uniformly across freqiency.
-Since phase error at all frequencies are weighted equally, this technique will suffer from error accumulation if the source has a lot of low power regions in the frequency domain.
+In effect it ignores the signal magnitude and weighs phase errors uniformly across frequencies.
+Since the phase error at every frequency is weighted equally, this technique will suffer from error accumulation if the source has a lot of low power regions in the frequency domain.
 

### 2015-06-01 reports/paper/simulation.tex
paper update

@@
 A test sound source is emitted at point $P$ which is $50$ centimeters away from the origin $(0,0)$.
-Let $2a$ denote the difference of distance between $P$ and two microphones:
+Let $2a$ denote the TDOA between $P$ and two microphones:
 \[
-2a = P M_1 - P M_2 
+2a = \frac{P M_1 - P M_2}{340\mbox{ m/s}}
 \]
-then the TDOA is $2a/340$ seconds. Assuming a sampling rate of $34$ KHz, fig~\ref{fig:sim_2_5} shows the region $R$ where all points have TDOA close to $2a/340$ s within one sample difference:
+where $340$ m/s is used as the speed of sound. Assuming a sampling rate of $34$ KHz, fig~\ref{fig:sim_2_5} shows the region $R$ where all points have TDOA close to $2a$ s within one sample difference:
 \[
-R=\{\hat P: |(\hat P M_1 - \hat P M_2) - 2a|< \frac{1}{2} \mbox{ samples}\}
+R=\{\hat P: |\frac{(\hat P M_1 - \hat P M_2)}{340\mbox{ m/s}} - 2a|< \frac{1}{2} \mbox{ samples}\}
 \]
-Intuitively, points in $R$ have difference of distance to two microphones very similar to each other.
+Intuitively, points in $R$ have TDOA to two microphones very similar to each other.
 Looking at fig~\ref{fig:sim_2_5}, we can still see that $R$ has the shape of a hyperbola, but with an uncertainty region around it.
 The thickness of the uncertainty region is not uniform around the hyperbola, the farther away the point is, the larger the uncertainty region becomes.
-This indicates for the same delta distance movement it will generate smaller difference of distance change when the source is farther away from the array.
+This indicates for the same delta distance movement it will generate smaller TDOA change when the source is farther away from the array.
 The size of the uncertainty region is also angle dependent: points closer to the line connecting microphones have larger region compared to points close to the line bisecting microphones.
@@
 
-To evaluate how adding one microphone(without increasing the array size) improves accuracy, another microphone is added to the array at $(0,0)$.
+To evaluate how adding one microphone (without increasing the array size) improves accuracy, another microphone is added to the array at $(0,0)$.
 Result is presented in fig~\ref{fig:sim_hm_3_p1}.
@@
 
-To avoid the need to accurately place microphones at far distances(as required by fig~\ref{fig:sim_hm_4}), we explored configuration with two arrays.
+To avoid the need to accurately place microphones at far distances (as required by fig~\ref{fig:sim_hm_4}), we explored configuration with two arrays.
 Two $3$ microphone array are placed $1$ meter apart and the result is presented in fig~\ref{fig:sim_hm_2_array}.

### 2015-06-16 reports/paper/experiment.tex
thesis update

@@
   \centering
-  \begin{subfigure}[]{.46\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{setup_1.JPG}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.46\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{setup_2.JPG}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{average_error_window_size.eps}
+\includegraphics[width=1.0\textwidth]{average_error_window_size.eps}
 \caption{Localization error versus window size}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{calculation_time.eps}
+\includegraphics[width=1.0\textwidth]{calculation_time.eps}
 \caption{Computation time versus window size}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{error_distribution.eps}
+\includegraphics[width=1.0\textwidth]{error_distribution.eps}
 \caption{error distribution in the grid. Arrays are placed at $(-0.5$ m$, 0$ m$)$ and $(0.5$ m$, 0$ m$)$}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{error_along_y_axis.eps}
+\includegraphics[width=1.0\textwidth]{error_along_y_axis.eps}
 \caption{Localization error as the distance between the source and the microphone arrays increases. The source is placed on the $y$ axis.}
@@
 %  \end{subfigure}
-  \begin{subfigure}[]{.5\textwidth}
+  \begin{subfigure}[]{1.0\textwidth}
     \includegraphics[width=\textwidth]{setup_rotating_disk.JPG}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{error_window_size_movement.eps}
+\includegraphics[width=1.0\textwidth]{error_window_size_movement.eps}
 \caption{Localization error versus window size}
@@
 \centering
-\includegraphics[width=.8\textwidth]{trace_output_circle_wnn.eps}
+\includegraphics[width=1.0\textwidth]{trace_output_circle_wnn.eps}
 \caption{white noise ($10$ cm per second)}
@@
 \centering
-\includegraphics[width=.8\textwidth]{trace_output_circle_man.eps}
+\includegraphics[width=1.0\textwidth]{trace_output_circle_man.eps}
 \caption{music A ($10$ cm per second)}
@@
 \centering
-\includegraphics[width=.8\textwidth]{trace_output_circle_mbn.eps}
+\includegraphics[width=1.0\textwidth]{trace_output_circle_mbn.eps}
 \caption{music B ($10$ cm per second)}
@@
 \centering
-\includegraphics[width=.8\textwidth]{trace_output_circle_wnf.eps}
+\includegraphics[width=1.0\textwidth]{trace_output_circle_wnf.eps}
 \caption{white noise ($20$ cm per second)}
@@
 \centering
-  \includegraphics[width=.8\textwidth]{trace_output_circle_maf.eps}
+  \includegraphics[width=1.0\textwidth]{trace_output_circle_maf.eps}
   \caption{music A ($20$ cm per second)}
@@
 \centering
-  \includegraphics[width=.8\textwidth]{trace_output_circle_mbf.eps}
+  \includegraphics[width=1.0\textwidth]{trace_output_circle_mbf.eps}
   \caption{music B ($20$ cm per second)}

### 2015-06-16 reports/paper/prior.tex
thesis update

@@
 Cross-correlation has also been used to measure signal arrival time differences\cite{tdoa:mit2, tdoa:micloc, tdoa:3}.
+Cross correlation is a measure of similarity between two signals.
+For real valued signals $x_1(t)$, $x_2(t)$ a time shift $\tau$:
+\begin{eqnarray}\label{eqn:gcc0}
+\mbox{xcorr}_{x_1,x_2}(\tau) = \int_{-\infty}^{\infty} x_1(t) x_2(t+\tau)dt
+\end{eqnarray}
+We can take Fourier Transform on both sides of equation~\ref{eqn:gcc0}:
+\begin{eqnarray}\label{eqn:gcc1}
+\mathcal{F}\{\mbox{xcorr}_{x_1,x_2}(\tau)\} & = & X_1(\omega) X_2(-\omega)\\
+& =& X_1(\omega) X_2(\omega)^*
+\end{eqnarray}
+Where $X_1(\omega)$ and $X_2(\omega)$ are the Fourier Transforms of $x_1(t)$ and $x_2(t)$. We can retrieve the cross-correlation result in time domain by taking inverse Fourier transform:
+\begin{eqnarray}\label{eqn:gcc2}
+\mbox{xcorr}_{x_1,x_2}(\tau) = \int_{-\infty}^\infty X_1(\omega)X_2(\omega)^*e^{j\omega\tau} d\omega
+\end{eqnarray}
+The arrival time difference $t_0$ is the time shift $\tau$ that maximizes~\ref{eqn:gcc2}:
+\begin{eqnarray}\label{eqn:gcc3}
+t_0 = \arg\max_\tau\mbox{xcorr}_{x_1,x_2}(\tau) 
+\end{eqnarray}
+
+The benefits of calculating cross-correlation in the frequency domain as shown in equation~\ref{eqn:gcc2} are two folds.
+The first benefit is to speedup the calculation.
+Calculating cross-correlation using equation~\ref{eqn:gcc0} requires multiplying and summing the two signal vectors for each time shift $\tau$.
+With discrete signals of length $n$, it will take $O(n^2)$ number of calculations.
+Doing the same calculation in frequency domain, we need to transform the signal into frequency domain, multiple and sum the two transformed signal vectors once and then transform the result back to the time domain.
+Transforming a signal from time domain to frequency domain and back can be done efficiently with Fast Fourier Transform (FFT) and Inverse Fast Fourier Transform (IFFT), and the amount of calculation needed is $O(n\log n)$.
+Multiply and sum the transformed signal vectors takes another $O(n)$.
+Therefore, the total calculation required to calculate cross-correlation using Fourier transform is $O(n\log n)$, which is asymptotically faster than calculating in the time domain.
+
+The second benefit of formulating cross correlation in frequency domain is that it provides a unified framework to prefilter the signals.
 Cross-correlation with prefiltering is known as \emph{generalized cross correlation (GCC)}.
 Different prefiltering approaches have been investigated to improve arrival time difference estimation~\cite{tdoa:gcc1,tdoa:gcc2,tdoa:gcc3}.
-
-Under the GCC framework, the arrival time difference $t_0$ between two signals $x_1(t)$ and $x_2(t)$ can be estimated as:
+Under the GCC framework, the arrival time difference $t_0$ between two signals $x_1(t)$ and $x_2(t)$ is estimated as:
 \begin{eqnarray} \label{eq:gcc}
@@
 \end{eqnarray}
-where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transforms of $x_1(t)$ and $x_2(t)$.
-$W(\omega)$ provides a way to prefilter signals passed to the cross correlation estimator.
-We focused on three ways of prefiltering the signal:
+$W(\omega)$ provides a way to prefilter signals passed to the cross correlation estimator. We focused on three ways of prefiltering the signal:
 \begin{description}%[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]

### 2015-06-16 reports/paper/conclusion.tex
paper update

--- 2015-05-29 reports/paper/conclusion.tex
@@
 \chapter{Conclusion}
-In this paper, we proposed and built an inexpensive, portable yet reasonably accurate sound localization system with two microphone arrays.
-We analyzed different array architectures and demonstrated that a two array architecture achieves better accuracy compared to a single array system of similar size.
+In this thesis, we described and built an inexpensive, portable yet reasonably accurate sound localization system with two microphone arrays.
+We analyzed different array architectures and showed that the localization accuracy varies with the array geometry and the location of the sound source (both the distance and the angle).
+In particular, regions close to the line connecting two microphones are generally more difficult to localize than regions close the line bisecting two microphones.
+This problem can be alleviated by employing multiple microphones that are placed at large distance from each other, but such arrangement affects the portability of the system.
+As a trade off between accuracy and portability, we demonstrated that a two array architecture achieves better accuracy compared to a single array system of similar size.
 By using cross-correlation output as the likelihood for different arrival time differences, each array generates a likelihood map for each possible location in the area.
 We demonstrated that by merging likelihood maps from the two arrays, we were able to achieve good localization accuracy (less than $3$ cm average error) in a local one meter by one meter region.
+Since two arrays generate their likelihood maps individually, this system does not require synchronized clock between the arrays, which made the overall system easier to design and more portable.
 
-This system can be used in HCI applications that incorporate sound position and movement information, such as drawing with music, or can be used to create AI games with physical pieces where the computer controls some of the pieces.
-This system can also be used to in Augmented Reality (AR) applications that allow user computer interaction with music tags.
+This system can be used in HCI applications that incorporate sound position and movement information.
+One can use this system to build virtual drawing applications where the user can draw with music, without physically touching the computer.
+One can also use this system to design interactive AI games.
+For example, a chess game with physical pieces can be developed where each piece is equipped with a motor and a music tag.
+Since the computer knows where all the pieces are, it knows which piece the user has moved and can make its corresponding move.
+A similar example is AI toy car racing game, where the player controls one car and the computer controls another car.
+With real time location information, the computer can control one car to compete with the player on a racing track.
+This system can also be used to in Augmented Reality (AR) applications.
+For example, users can wear a music tag on the finger, and then the system would be able to track finger movement.
+It can be used as virtual mouse, and applications can be developed to detect finger gesture.
 
-For future directions, this system can be extended with another microphone on each array to perform 3D localization.
-Another possible direction is to investigate multi-source localization using digitally tagged audio sources.
+For future directions, one can extend the system with another microphone on each array to perform 3D localization.
+However, the number of grid points increases exponentially with the number of dimension, which means the calculation time also increases exponentially.
+Therefore, it is important to research more efficient ways to find the intersection of hyperbolic cones the in 3D space.
+Since the likelihood calculation on each grid point is independent of each other, one straightforward way to speedup calculation is to employ hardware accelerated parallel computing technologies such as GPU or FPGA.
+Another useful direction is to investigate multi-source localization using digitally tagged audio sources.

### 2015-06-16 reports/paper/experiment.tex
paper update

@@
 \chapter{Experiment}
+\section{Setup}
+\input{./system}
+
 We conducted two sets of experiments to evaluate the system's localization accuracy: one on point localization and the other on movement tracking.
 
-\section{Setup}
 \subsection{Point localization}

### 2015-06-16 reports/paper/prior.tex
paper update

@@
 
-Let $s(t)$ denote the audio signal. The received signal $\bar s(t) $ at microphone $i$ can be modeled as:
+For a point sound source in a direct field, the signal intensity decays proportional to the square of the distance between the source and the microphone.
+Let $I_i$ denote the received signal intensity at microphone $i$:
 \[
-\bar s_i(t) = \frac{s(t)}{d_i} + \delta_i(t)
+I_i \propto \frac{1}{d_i^2} I_s
 \]
-where $d_i$ is the distance between the audio source and microphone $i$, and $\delta_i(t)$ represents Gaussian noise.
-The energy received at microphone $i$ is:
+where $d_i$ is the distance between the audio source and microphone $i$, and $I_s$ is the sound intensity at the source.
+With two microphones, the signal intensity received at both microphones has to satisfy equation~\ref{eqn:ild1}:
 \begin{eqnarray}
-E_i & = & \int \bar s_i(t)^2 dt \\
-\label{eqn:ild1}
-& = & \frac{1}{d_i^2} E_s + E_{\sigma_i}
+I_1d_1^2 &=& I_2d_2^2\\
+\label{eqn:ild}
+\frac{I_1}{I_2} &=& \frac{d_2^2}{d_1^2}
 \end{eqnarray}
-It shows that the energy decays with the square of the distance.
-With two microphones, not considering the noise $\delta_i(t)$, energy received at both microphones has to satisfy equation~\ref{eqn:ild1}:
-\begin{eqnarray}
-E_1d_1^2 &=& E_2d_2^2\\
-\label{eqn:ild}
-\frac{E_1}{E_2} &=& \frac{d_2^2}{d_1^2}
-\end{eqnarray}
-It can be shown that, on a 2D plane, points satisfy equation~\ref{eqn:ild} form a circle when $E_1 \ne E_2$ and form a line when $E_1 = E_2$~\cite{ild:1}.
-With two microphones all points on this curve generate the same energy ratio and can not be distinguished from each other.
+It can be shown that, on a 2D plane, points satisfy equation~\ref{eqn:ild} form a circle when $I_1 \ne I_2$ and form a line when $I_1 = I_2$~\cite{ild:1}.
+With two microphones all points on this curve generate the same intensity ratio and can not be distinguished from each other.
 Multiple approaches have been investigated to eliminate this ambiguity.
 \cite{ild:1} employed multiple microphones and used the intersection of circles from each microphone pair to estimate the source location.
-\cite{ild:2} combined ILD with Interaural Time Difference to estimate the source direction (i.e azimuth).
+\cite{ild:3} combined ILD with Interaural Time Difference (ITD) to estimate the source direction (i.e azimuth).
+Instead of solving the intersection of circles, \cite{ild:4} employed machine learning techniques to automatically learn the mapping from ILD and ITD features to location coordinate.
+It uses four microphones and requires a training phase, during which the sound source is manually placed at predetermined locations for the system to learn the parameters that map from feature space to the sound location.
 
-ILD approach relies on the accurate measurement of the received energy ratio between microphone pairs.
-Any obstacle object between the sound source and any microphone would produce a significant distortion in the measured energy ratio.
+ILD approach relies on the accurate measurement of the received intensity ratio between microphone pairs.
+Any obstacle object between the sound source and any microphone would produce a significant distortion in the measured intensity ratio.
 We find this approach to be too restrictive for our system, since in an interactive system we can not control whether or not the user places any obstacle between the sound source and the microphones.
@@
 R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) |X_1(\omega)||X_2(\omega)| e^{j(\omega\tau - (\angle{X_2(\omega)} - \angle{X_1(\omega)}) } d\omega\\
-&=& \int_{-\infty}^\infty \underbrace{W(w)|X_1(\omega)| |X_2(\omega)|}_{\mbox{weighting}} \underbrace{\cos(\omega\tau - (\angle X_2(\omega) - \angle X_1(\omega)))}_{\mbox{phase error}} d\omega
+&=& \int_{-\infty}^\infty \underbrace{W(w)|X_1(\omega)| |X_2(\omega)|}_{\mbox{weighting}} \cos(\Theta_\epsilon) d\omega
 \end{eqnarray}
+Where $\Theta_{e}$ is the phase error:
+\[
+\Theta_\epsilon = \omega\tau - (\angle X_2(\omega) - \angle X_1(\omega))
+\]
 We can look at the real part of equation~\ref{eqn:phat1} only since both $x_1(t)$ and $x_2(t)$ are real valued signals.
-When $\dot \tau$ is the true arrival time difference between $x_1(t)$ and $x_2(t)$, $\omega \dot \tau - (\angle X_2(\omega) - \angle X_1(\omega)) = 0$, and $\cos(\omega \dot \tau - (\angle X_2(\omega) - \angle X_1(\omega))) = 1$.
-When $\dot \tau$ differs from the true arrival time difference, $\cos(\omega \dot \tau - (\angle X_2(\omega) - \angle X_1(\omega))) < 1$.
-Therefore, $\cos(\omega\tau - (\angle X_2(\omega) - \angle X_1(\omega)))$ can be seen as a measure of the phase error, and $W(w)|X_1(\omega)||X_2(\omega)|$ describes how the error should be weighted at each frequency.
+When $\dot \tau$ is the true arrival time difference between $x_1(t)$ and $x_2(t)$, phase error $\Theta_\epsilon  = 0$, and $\cos(\Theta_\epsilon) = 1$.
+When $\dot \tau$ differs from the true arrival time difference, $\cos(\Theta_\epsilon) < 1$.
+Therefore, $\cos(\Theta_\epsilon)$ can be seen as a measure of the phase error, and $W(w)|X_1(\omega)||X_2(\omega)|$ describes how the error should be weighted at each frequency.
 The TDOA estimator essentially sums the weighted phase error at each frequency.

### 2015-06-16 reports/paper/conclusion.tex
thesis

@@
 For future directions, one can extend the system with another microphone on each array to perform 3D localization.
-However, the number of grid points increases exponentially with the number of dimension, which means the calculation time also increases exponentially.
-Therefore, it is important to research more efficient ways to find the intersection of hyperbolic cones the in 3D space.
-Since the likelihood calculation on each grid point is independent of each other, one straightforward way to speedup calculation is to employ hardware accelerated parallel computing technologies such as GPU or FPGA.
-Another useful direction is to investigate multi-source localization using digitally tagged audio sources.
+However, the number of grid points increases exponentially with the number of dimensions, which translates into exponential increase in computation time.
+To keep the localization in real time (time lag of less than $0.1$ seconds), it is important to research more efficient ways of finding the intersection of hyperbolic cones the in the 3D space.
+Since the likelihood calculation on each grid point is simply a look up on the cross-correlation output from each microphone pair, the computation at each grid point is independent of each other.
+Therefore, One straightforward way to speedup calculation is to employ hardware accelerated parallel computing technologies such as GPU or FPGA.
+
+Another direction is to investigate multi-source localization.
+One approach is to allocate different frequency bands to difference sound sources.
+However, the number of frequency bands are limited in the audible audio spectrum.
+The arrival time difference estimation is also more immune to noise for wider band audio sources.
+A different approach is to digitally encode and tag the audio source.
+Different sources can be differentiated if each source uses a waveform that has a different digital signature.

### 2015-06-16 reports/paper/experiment.tex
thesis

@@
 The lowest error achieved is $2.53$ centimeters, which occurred when the window size is set to 12 millisecond and GCC\_PHAT is used for TDOA estimation.
+The performance differences among GCC, GCC\_PHAT and GCC\_PHAT\_SQRT is small.
 
@@
 When window size is small ($1.02$ millisecond), the audio does not contain enough information to reliably estimate TDOA, which results in noisy localization.
-As window size increases, localization converges to the shape of the ground truth circle.
+As window size increases, the TDOA estimation becomes more accurate and the localization converges to the shape of the ground truth circle.
 Fig~\ref{fig:err_win_circle} shows how the error changes with window size.
@@
 Fig~\ref{fig:circle_wnn} to \ref{fig:circle_musicbn} shows results for experiments at normal speed, and Fig~\ref{fig:circle_wnf} to~\ref{fig:circle_musicbf} shows results at fast speed.
-By comparing localization error for each audio source between normal movement speed and fast movement speed, we find that localization error does not depend on how fast the sound source is moving.
+By comparing localization error for each audio source between normal movement speed and fast movement speed, we find that localization error does not increase when the source movement speed increased from $10$ cm to $20$ cm.
 For example fig~\ref{fig:circle_musican} and fig~\ref{fig:circle_musicaf} shows that localization error is $1.289$ cm at normal movement speed and $1.291$ cm at fast movement speed.
 
-For normal speed movement tracking, localization error is $0.9$ cm for white noise, $1.29$ cm for Music A, and $2.9$ cm for Music B.
+For normal speed movement tracking, localization error is $0.9$ cm for white noise (Fig~\ref{fig:circle_wnn}), $1.29$ cm for Music A (Fig~\ref{fig:circle_musican}), and $2.9$ cm for Music B (Fig~\ref{fig:circle_musicbn}).
 Localization accuracy is the best for white noise source, and worst for Music B.
@@
 It also shows that raw detection has the most amount of jiggling.
-Kalman filter reduces the amount of jiggling from raw detection.
+Kalman filter reduces the amount of jiggling from raw detection by combining past estimates with current estimates.
 Averaging filter has the least amount of jiggling.

### 2015-06-16 reports/paper/introduction.tex
thesis

@@
 Accurate indoor localization allows creation of novel applications with surrounding awareness that uses position and movement information as input.
-One application is to allow users to draw with music without physically touching the computer.
-Another example is to build AI games with physical pieces such as toy car racing where the computer controls some of the toy cars.
-In this work, we aim to build such a source localization system that is portable, inexpensive, yet reasonably accurate for localization in a small area.
 
@@
 
-In this project, we have built a localization system with reasonably high precision for small area using microphone arrays that localize typical audio sources.
+In this work, we have built a source localization system that is portable, inexpensive, yet reasonably accurate for localization in a small area.
 Our system is built with inexpensive electret microphones mounted on portable frames.
 Users can interact with our system using any device that has audio output such as a mobile phone.
+
+This system can be used in Human-Computer Interaction (HCI) applications that incorporate sound position and movement information.
+One can use this system to build virtual drawing applications where the user can draw with music without physically touching the computer.
+One can also use this system to design interactive AI games with physical pieces.
+For example, a chess game with physical pieces can be developed where each piece is equipped with a motor and a music tag.
+Since the computer knows where all the pieces are and where the user has moved, it can make its corresponding move.
+Another similar example is to build an AI toy car racing game, where the player controls one car and the computer controls another.
+With real time location information, the computer can compete with the player on a racing track.
+
+This system can also be used to in Augmented Reality (AR) applications.
+For example, the user can wear a music tag on the finger, and then the system would be able to track the finger movement.
+this particular setup can be used as a virtual mouse for wearable technologies such as Google Glass.
 

### 2015-06-16 reports/paper/prior.tex
thesis

@@
 Let $I_i$ denote the received signal intensity at microphone $i$:
-\[
+\begin{eqnarray}\label{eqn:ild1}
 I_i \propto \frac{1}{d_i^2} I_s
-\]
+\end{eqnarray}
 where $d_i$ is the distance between the audio source and microphone $i$, and $I_s$ is the sound intensity at the source.
@@
 Cross correlation is a measure of similarity between two signals.
-For real valued signals $x_1(t)$, $x_2(t)$ a time shift $\tau$:
+For real valued signals $x_1(t)$ and $x_2(t)$, the cross-correlation between them at a particular time shift $\tau$ can be calculated as:
 \begin{eqnarray}\label{eqn:gcc0}
-\mbox{xcorr}_{x_1,x_2}(\tau) = \int_{-\infty}^{\infty} x_1(t) x_2(t+\tau)dt
+ R_{x_1,x_2}(\tau) = \int_{-\infty}^{\infty} x_1(t) x_2(t+\tau)dt
 \end{eqnarray}
@@
 \begin{eqnarray}\label{eqn:gcc1}
-\mathcal{F}\{\mbox{xcorr}_{x_1,x_2}(\tau)\} & = & X_1(\omega) X_2(-\omega)\\
+\mathcal{F}\{ R_{x_1,x_2}(\tau)\} & = & X_1(\omega) X_2(-\omega)\\
 & =& X_1(\omega) X_2(\omega)^*
@@
 \begin{eqnarray}\label{eqn:gcc2}
-\mbox{xcorr}_{x_1,x_2}(\tau) = \int_{-\infty}^\infty X_1(\omega)X_2(\omega)^*e^{j\omega\tau} d\omega
+ R_{x_1,x_2}(\tau) = \int_{-\infty}^\infty X_1(\omega)X_2(\omega)^*e^{j\omega\tau} d\omega
 \end{eqnarray}
@@
 \begin{eqnarray}\label{eqn:gcc3}
-t_0 = \arg\max_\tau\mbox{xcorr}_{x_1,x_2}(\tau) 
+t_0 = \arg\max_\tau R_{x_1,x_2}(\tau) 
 \end{eqnarray}
@@
 We can look at the real part of equation~\ref{eqn:phat1} only since both $x_1(t)$ and $x_2(t)$ are real valued signals.
-When $\dot \tau$ is the true arrival time difference between $x_1(t)$ and $x_2(t)$, phase error $\Theta_\epsilon  = 0$, and $\cos(\Theta_\epsilon) = 1$.
-When $\dot \tau$ differs from the true arrival time difference, $\cos(\Theta_\epsilon) < 1$.
+When $\tau$ is the true arrival time difference between $x_1(t)$ and $x_2(t)$, phase error $\Theta_\epsilon  = 0$, and $\cos(\Theta_\epsilon) = 1$.
+When $\tau$ differs from the true arrival time difference, $\cos(\Theta_\epsilon) < 1$.
 Therefore, $\cos(\Theta_\epsilon)$ can be seen as a measure of the phase error, and $W(w)|X_1(\omega)||X_2(\omega)|$ describes how the error should be weighted at each frequency.

### 2015-06-16 reports/paper/simulation.tex
thesis update
thesis

@@
   \centering
-  \begin{subfigure}[]{.46\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{sim/sim_2_1}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.46\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{sim/sim_2_2}
@@
   \centering
-  \begin{subfigure}[]{.46\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{sim/result_20cm_0_degrees}
@@
 %  \end{subfigure}
-  \begin{subfigure}[]{.46\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{sim/result_20cm_135_degrees}
@@
   \centering
-  \begin{subfigure}[]{.46\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{sim/result_80cm_0_degrees}
@@
 %  \end{subfigure}
-  \begin{subfigure}[]{.46\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{sim/result_80cm_135_degrees}
@@
   \centering
-  \begin{subfigure}[]{.3\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_center}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.3\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_p1}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.3\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_2x}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.3\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_line}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.3\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_4corner}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.3\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_2a}
@@
 Regions near lines connecting microphones still have significantly larger uncertainty region.
+This result suggests that adding more microphones in a small microphone array (without increasing the array size) is not an effective method to improve the localization accuracy.
 
@@
 The average error improved to $10.04$ cm.
+By comparing with the previous result, it shows that increasing array size is more effective in improving accuracy than adding more microphones.
 
 In fig~\ref{fig:sim_hm_3_line}, three microphones are placed $10$ cm apart from each other on the x-axis.
-Error heatmap shows high uncertainty on the x axis, and the overall accuracy is not as good as that with three microphones placed in a triangle.
+Error heatmap shows high uncertainty on the x axis, and the overall accuracy much worse than that with three microphones placed in a triangle.
 The average error is $55.05$ cm.

### 2015-06-17 reports/paper/introduction.tex
paper update

@@
 Global Positioning System (GPS) is the prevailing technology used for outdoor localization.
+A set of 32 satellites equipped with synchronized atomic clock orbit around the earth and broadcast their time and position information at fixed intervals.
+A GPS receiver listens to at least 4 satellites and use their broadcasted timing information to infer the location.
 Commercial grade GPS has an average error of a few meters, depending on the size and quality of the receiver~\cite{intro:gps}.
@@
 For example, a chess game with physical pieces can be developed where each piece is equipped with a motor and a music tag.
-Since the computer knows where all the pieces are and where the user has moved, it can make its corresponding move.
+Since the computer knows where all the pieces are and which one the user has moved, it can make its corresponding move.
 Another similar example is to build an AI toy car racing game, where the player controls one car and the computer controls another.
-With real time location information, the computer can compete with the player on a racing track.
+With real time location information, the AI engine can compete with the player on a racing track.
 
 This system can also be used to in Augmented Reality (AR) applications.
-For example, the user can wear a music tag on the finger, and then the system would be able to track the finger movement.
-this particular setup can be used as a virtual mouse for wearable technologies such as Google Glass.
+For example, the user can wear a music tag on one finger, and then the system would be able to track the finger movement.
+This particular setup can be used as a virtual mouse for wearable technologies such as Google Glass.
 
@@
 We demonstrated that a two array system outperforms a single array system of similar physical dimension.
-In Chapter IV, we presented the chosen architecture along with hardware details.
-Finally, experiment details and results were presented in Chapter V.
+In Chapter IV, we first presented the chosen architecture along with hardware details, followed by experiment details and results were presented.

### 2015-06-17 reports/paper/introduction.tex
thesis update

@@
 A set of 32 satellites equipped with synchronized atomic clock orbit around the earth and broadcast their time and position information at fixed intervals.
-A GPS receiver listens to at least 4 satellites and use their broadcasted timing information to infer the location.
+A GPS receiver listens to at least 4 satellites and uses the broadcasted timing information to infer the location.
 Commercial grade GPS has an average error of a few meters, depending on the size and quality of the receiver~\cite{intro:gps}.
 While accuracy in this range is good for many applications including driving navigation and vehicle tracking, it does not provide enough precision for local movement tracking.
-Ultrasound based indoor localization approaches on the other hand, have achieved sub-centimeter accuracy~\cite{intro:ultra}.
-However, ultrasound systems require the use of expensive transducers.
+
+Ultrasound based indoor localization approaches, on the other hand, have achieved sub-centimeter accuracy~\cite{intro:ultra}.
+These systems employ multiple ultrasound receivers and use the arrival time difference between the sound source and receivers to infer the source location.
+Although obtaining good accuracy, ultrasound systems require the use of expensive transducers.
 
 Bluetooth and Wi-Fi based technologies have gained popularity in indoor positioning recently, mainly due to the widespread deployment of bluetooth tags and Wi-Fi stations in public spaces.
+The fact that modern consumer devices such as mobile phones and tablets are commonly equipped with Wi-Fi and bluetooth modules also made this approach particular attractive because no hardware needs to be installed on the user end.
+Recent applications including serving advertisements and coupons to consumers in shopping centers based on the customer's location.
 In these systems, signal strength received from different base stations are used for the estimation of the device location.

### 2015-06-18 reports/paper/conclusion.tex
thesis update

@@
 For example, users can wear a music tag on the finger, and then the system would be able to track finger movement.
-It can be used as virtual mouse, and applications can be developed to detect finger gesture.
+This particular setup can be used as a virtual mouse for wearable technologies such as Google Glass.
 
@@
 One approach is to allocate different frequency bands to difference sound sources.
-However, the number of frequency bands are limited in the audible audio spectrum.
-The arrival time difference estimation is also more immune to noise for wider band audio sources.
+However, the number of frequency bands are limited in the audible audio spectrum, and our results showed that TDOA estimation is more accurate for a wide band signal compared to a narrow band signal.
 A different approach is to digitally encode and tag the audio source.

### 2015-06-18 reports/paper/prior.tex
thesis update

@@
 Localization can be performed by comparing the incoming waveform with the stored templates, and the location with the best matching template is chosen as the output.
+Authors in \cite{extended:tusi} has built a virtual piano keyboard application where different musical notes are associated with different locations on a surface.
+A user can then play by tapping at different locations on the surface.
 Different ways of extracting templates from raw acoustic source and different similarity measures have been investigated in the past.
@@
 The requirement of having a template for each location to be detected makes this approach too restrictive for our project, since we want the localization to be continuous in a 2D region.
+\cite{ltm:tusi2} has investigated contiguous localization by linearly extrapolating between stored locations, but result has high error.
 Moreover, the need to recalibrate all locations during setup is too cumbersome for the end users in a portable system.
@@
 The algorithm then used extracted parameters to predict arrival time with a second order polynomial, the parameters of which were fitted during calibration at fixed locations.
+Authors in \cite{tdoa:mit5} have used similar techniques and built an interactive window by placing four microphones on four corners of a glass pane.
+The glass window was installed in shopping centers.
+A projector projects product information onto the glass where consumers can browse between pages by tapping on the window.
 
@@
 Therefore, the total calculation required to calculate cross-correlation using Fourier transform is $O(n\log n)$, which is asymptotically faster than calculating in the time domain.
+This calculation speedup is especially beneficial in real-time interactive systems since significant time lag would produce unsmooth user experience.
 
@@
 Since the phase error at every frequency is weighted equally, this technique will suffer from error accumulation if the source has a lot of low power regions in the frequency domain.
+This weighting is also beneficial if the source signal is white noise, since white noise should contain all frequency components with equal magnitude.
 

### 2015-06-18 reports/paper/simulation.tex
thesis update

@@
 
-In fig~\ref{fig:sim_hm_3_line}, three microphones are placed $10$ cm apart from each other on the x-axis.
-Error heatmap shows high uncertainty on the x axis, and the overall accuracy much worse than that with three microphones placed in a triangle.
+In fig~\ref{fig:sim_hm_3_line}, we experimented with an extreme arrangement where all three microphones are placed on a line, $10$ cm apart from each other.
+Error heatmap shows high uncertainty on the x axis, and the overall accuracy is much worse than that with three microphones placed in a triangle.
 The average error is $55.05$ cm.

### 2015-06-18 reports/paper/experiment.tex
thesis update

@@
 A total of $32$ testing locations were chosen uniformly in this grid.
+Testing is done by placing the audio source at each grid location, and turning on the audio source with white noise.
 We reported the error as the average distance between our placement of the audio source and the location estimated from the arrays.
@@
 A sound source is placed on the edge of the rotating disk and the arrays track the sound source as it rotates in a circle.
+In this experiment we used GCC\_PHAT as the prefiltering for cross-correlation because in the point localization experiment we found out GCC\_PHAT gives the best result.
 In this experiment, we evaluated how accuracy changes with:
@@
 
+To test how localization accuracy varies with different window size of audio signal received, we conducted the experiment with audio signals with varying length from 1.02 ms to 12 ms.
+
 To test how different sound sources impact localization quality, we conducted the experiments with three different sound sources:
@@
 
-\item[White Noise] A recording of white noise.
+\item[White Noise] A recording of white noise. We expect GCC\_PHAT works best with white noise. The white noise is generated by sampling uniform randomly between $-1$ and $1$.
 
-\item[Music A] A randomly picked music that has normal audio amplitude throughout the experiment period. \emph{Honest Eyes} by \emph{Black Tide} was the music used.
+\item[Music A] A randomly picked music that has non-zero audio amplitude throughout the experiment period. \emph{Honest Eyes} by \emph{Black Tide} was the music used.
 

### 2015-06-18 reports/paper/introduction.tex
thesis update

@@
 In this work, we have built a source localization system that is portable, inexpensive, yet reasonably accurate for localization in a small area.
+To test the accuracy of our localization system, we have acoustic localization in an one meter by one meter area.
+The experiment area size is adequate for our purposes because human arm movement are constrained in this area.
+Our experiment show that our system reports reasonably good accuracy in this area.
 Our system is built with inexpensive electret microphones mounted on portable frames.

### 2015-06-18 reports/paper/prior.tex
thesis update

@@
 \chapter{Background}
+
+\section{Point localization}
 Acoustic localization has been researched extensively in the literature.
@@
 
-\section{ILD}
+\subsection{ILD}
 ILD techniques rely on the observation that signal intensity decays as the distance to the microphone increases.
@@
 
-\section{LTM}
+\subsection{LTM}
 In LTM based approaches, acoustic templates acquired from different locations are first stored in the system during a ``training'' phase.
@@
 
-\section{TDOA}
+\subsection{TDOA}
 TDOA approaches exploit the difference of arrival time between the acoustic source and two fixed microphones on the plane.
@@
 This approach represents a balance between unfilterred GCC and GCC\_PHAT.
+
+\section{Movement Tracking}
+a
+\subsection{Averaging filtering}
+a
+\subsection{Kalman filtering}
+a

### 2015-06-18 reports/paper/system.tex
paper update
thesis update

@@
-\chapter{System Setup}
 The end system has two arrays, each with three microphones mounted on the vertices of a $20$ cm equilateral triangle.
+We used omni-directional foil electret condenser microphones due to their low cost.
+The microphone has a frequency range of $100$ to $10K$ Hz, and a minimum SNR of 58dB[x].
+We used operational amplifer OPA344 from Texas Instrument to amplify the microphone output by a factor of 100.
 A micro-controller is attached to one of the vertices.
 Fig~\ref{fig:setup_array} shows a picture of the array setup.
-The micro-controller used in this project is \emph{teensy 3.1}, which has $64$k RAM memory and the ADC is capable of sampling at $500$kHz~\cite{tdoa:micloc, sys:teensy}.
+The micro-controller baord in this project is \emph{teensy 3.1}.
+It contains a which has a onboard RAM of $64$k.
+ADC is capable of sampling at $500$kHz~\cite{tdoa:micloc, sys:teensy}.
 In this project, the micro-controller collects microphone data on all three channels for $12$ milliseconds and then sends the recorded data to a computer through the USB port for localization.
@@
   \centering
-  \begin{subfigure}[]{.23\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{array_close.JPG}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.23\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{array.JPG}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.23\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{setup_2.JPG}
@@
   \centering
-  \begin{subfigure}[]{.23\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{left.png}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.23\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{right.png}
@@
   \end{subfigure}
-  \begin{subfigure}[]{.23\textwidth}
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{combined.png}

### 2015-06-19 reports/paper/experiment.tex
thesis update

@@
 \chapter{Experiment}
+\section{System}
+\input{./system}
+\clearpage
+
 \section{Setup}
-\input{./system}
 
 We conducted two sets of experiments to evaluate the system's localization accuracy: one on point localization and the other on movement tracking.
+For the point localization experiment we looked into using different window size of audio signals and different prefiltering methods.
+The size of the window limits how far apart the microphone arrays can be.
+If the time delay from one microphone array to another exceeds the size of the window, the location of the source can never be estimated.
+A large window size gives a more complete acoustic signal which makes cross-correlation less prone to noise.
+However, the larger the window size the more time it takes to compute the cross-correlation.
+This is a trade off that we will address.
+
+For the movement tracking experiment, on top of the varying window size, we also looked into varying the movement speed of the audio source, applying different movement filters, and using different types of audio source.
+By applying an increasing movement speed of the audio source, we can test how fast the system can track an audible object moving in real time.
+We experimented with two types of music as our audio source for movement tracking, one with no low volume throughout, and the other with intermittent low or no volume.
+We want to test how the system performs when the audio source isn't continuously outputting sound.
+Furthermore, we applied different movement filters to help reduce noise and smooth out the path of the moving audio source.
 
@@
 
-\begin{figure}[]
+\begin{figure}[h!]
   \centering
@@
 \subsection{Movement tracking}
+
+\begin{figure}[h!]
+  \centering
+%  \begin{subfigure}[]{.2\textwidth}
+%    \includegraphics[width=\textwidth]{setup_ring.JPG}
+%  \end{subfigure}
+  \begin{subfigure}[]{1.0\textwidth}
+    \includegraphics[width=\textwidth]{setup_rotating_disk.JPG}
+  \end{subfigure}
+  \caption{Setup for movement tracking evaluation}
+  \label{fig:setup_circle}
+\end{figure}
+
 To test how well the system tracks movement, we mounted a rotating disk $40$ centimeters in diameter onto the grid at $(x=0$ cm$, y=-30$ cm$)$.
@@
 To test how the movement speed of sound source affects localization quality, each experiment was conducted at two different speeds:
+
 \begin{description}
@@
 Therefore, the error is reported as the distance between the localized point to its closest point on the ground truth circle.
+\clearpage
 
@@
 \subsection{Point localization}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=1.0\textwidth]{average_error_window_size.eps}
+\caption{Localization error versus window size}
+\label{fig:accuracy_vs_window}
+\end{figure}
 To test how accuracy varies with window size, the algorithm is fed with microphone data of different lengths, and the result is shown in fig~\ref{fig:accuracy_vs_window}.
@@
 
-Although accuracy improves with the window size, computation time also increases with it.
+\begin{figure}[h!]
+\centering
+\includegraphics[width=1.0\textwidth]{calculation_time.eps}
+\caption{Computation time versus window size}
+\label{fig:speed_vs_window}
+\end{figure}
+As was mentioned before, although accuracy improves with the window size, computation time also increases with it.
 The part of calculation that depends on the window size is using cross correlation for TDOA estimation.
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=1.0\textwidth]{error_distribution.eps}
+\caption{error distribution in the grid. Arrays are placed at $(-0.5$ m$, 0$ m$)$ and $(0.5$ m$, 0$ m$)$}
+\label{fig:error_distribution}
+\end{figure}
 We also calculated the localization error for each tested point in the grid.
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=1.0\textwidth]{error_along_y_axis.eps}
+\caption{Localization error as the distance between the source and the microphone arrays increases. The source is placed on the $y$ axis.}
+\label{fig:error_along_y}
+\end{figure}
 To test the limit of the system and to evaluate the accuracy when the source moves outside of the one meter by one meter region, we measured the localization error by placing the source at $10$ locations along $y$ axis ranging from $(0$ cm$, -10$ cm$)$ to $(0$ cm$, -200$ cm$)$.
@@
 
-\begin{figure}[]
-\centering
-\includegraphics[width=1.0\textwidth]{average_error_window_size.eps}
-\caption{Localization error versus window size}
-\label{fig:accuracy_vs_window}
-\end{figure}
-
-\begin{figure}[]
-\centering
-\includegraphics[width=1.0\textwidth]{calculation_time.eps}
-\caption{Computation time versus window size}
-\label{fig:speed_vs_window}
-\end{figure}
-
-\begin{figure}[]
-\centering
-\includegraphics[width=1.0\textwidth]{error_distribution.eps}
-\caption{error distribution in the grid. Arrays are placed at $(-0.5$ m$, 0$ m$)$ and $(0.5$ m$, 0$ m$)$}
-\label{fig:error_distribution}
-\end{figure}
-
-\begin{figure}[]
-\centering
-\includegraphics[width=1.0\textwidth]{error_along_y_axis.eps}
-\caption{Localization error as the distance between the source and the microphone arrays increases. The source is placed on the $y$ axis.}
-\label{fig:error_along_y}
-\end{figure}
-
-\begin{figure}[]
-  \centering
-%  \begin{subfigure}[]{.2\textwidth}
-%    \includegraphics[width=\textwidth]{setup_ring.JPG}
-%  \end{subfigure}
-  \begin{subfigure}[]{1.0\textwidth}
-    \includegraphics[width=\textwidth]{setup_rotating_disk.JPG}
-  \end{subfigure}
-  \caption{Setup for movement tracking evaluation}
-  \label{fig:setup_circle}
-\end{figure}
-
-\begin{figure*}[]
+\clearpage
+\subsection{Movement tracking}
+\begin{figure*}[h!]
 \centering
@@
 
-\begin{figure}[]
+\begin{figure}[h!]
 \centering
@@
 \end{figure}
-
-\subsection{Movement tracking}
-
 Fig~\ref{fig:wn} gives an intuitive representation of how accuracy changes with window size.
@@
 
-\begin{figure*}[]
+\begin{figure*}[h!]
 \centering
@@
 
-\begin{figure*}[]
+Fig~\ref{fig:circle_wnn} shows the result when white noise is used as the sound source and the source is rotated at $10$ cm per second.
+The top right plot in this figures shows the raw detection output with the ground truth circle overlayed on top.
+It shows that the array's raw output matches the underneath ground truth circle reasonably well.
+The average error between the localization output with its closest point on the ground truth circle is $0.9$ cm.
+The top left plot in the figure shows the average error with different movement filters applied as a function of window size.
+The error decreases as window size increases and until the window size exceeds around $10$ ms.
+When the window size is $12$ cm, the localization accuracy is similar among raw output, Kalman filtering and averaging filtering.
+The bottom left plot shows the tracked movement for Averaging filter and the bottom right filter shows the tracked movement for Kalman filtering.
+Kalman filtering result is smoother while the averaging tracking stays closer to the ground truth circle.
+Compared to the experiment in fig~\ref{fig:circle_wnf}, where the same sound source is played but at a faster rotating speed of $20$ cm per second, we find that the performance is equally well between these two movement speeds.
+
+\begin{figure*}[h!]
 \centering
@@
 
-\begin{figure*}[]
+Fig~\ref{fig:circle_man} shows the result when Music A is used as the sound source.
+The top right plot shows that the array still tracks the movement well, but has a slightly larger deviation compared to that in fig~\ref{fig:circle_wnn}.
+The average localization error for Music A is $1.289$ cm.
+From the top left plot, it can be seen that the error still decreases with the window size.
+The performance improvement from raw output to Kalman filtering output is bigger than that with white noise (top left plot of fig~\ref{circle_wnn}).
+Averaging filtering still produces lowest average error.
+Fig~\ref{fig:circle_maf} shows the result for the same sound source moved at twice the speed ($20$ cm per second).
+The average localization error at the faster speed is $1.291$ cm.
+Compared to the experiment where the same sound source is placed but with a faster movement speed (fig~\ref{fig:circle_maf}, we found that the performance does not degrade as the movement speed increases from $10$ cm per second to $20$ cm per second.
+
+\begin{figure*}[h!]
 \centering
@@
 
-\begin{figure*}[]
+Fig~\ref{fig:circle_mbn} shows the result when Music B is used as the sound source.
+The low amplitude intervals in Music B affects the performance significantly.
+The average localization error is $2.9$.
+The ``blank'' regions in the music can also be visually seen from the top right plot.
+The top left plot shows that the Kalman filtering still performs better than of the white noise, and behaves similar to the case with Music A.
+Averaging filtering produces lowest average error.
+Compared to the faster movement experiment of the same sound source, we find that the localization error is similar between these two movement speeds.
+
+\begin{figure*}[h!]
 \centering
@@
 
+\subsection{Discussion}
 It also shows that raw detection has the most amount of jiggling.

### 2015-06-19 reports/paper/simulation.tex
thesis update

@@
 \chapter{Array Architecture}
+
+In this chapter, we first look at how localization accuracy is affected by TDOA accuracy in a two microphone setup.
+We then explore different microphone array configurations and select one that gives a good localization accuracy while being portable.
+
+\section{Two Microphones}
 As was mentioned in the previous chapter, points with the same TDOA to two fixed locations form a hyperbola on a 2D plane.
@@
 
+\section{Three Microphone Array}
 With more than two microphones, each pair of microphones generates a hyperbolic region and localization becomes finding the intersection of hyperbolic regions.
@@
 
-\begin{figure*}[]
+Each microphone pair generates a hyperbolic region, and the source location is in the intersection of these regions.
+The area of the intersection region is a measure of the localization accuracy: the smaller the area, the more certain we are about the source location.
+Different array configuration gives different intersection area size.
+To evaluate an array's accuracy in a region, we can place sound source at predetermined grid points in the region and look at the size of the intersection area for each tested point in the grid.
+The smaller the intersection area size, the better the configuration.
+Another measure we can use to evaluate an array configuration is the error distance between the centre of the intersection area to the actual source location.
+The smaller the error difference on average, the better the configuration.
+We used both the intersection area size and error distance as measures for evaluating the following configurations.
+
+\subsection{Linear Configuration}
+
+\begin{figure}[h!]
+  \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_line}
+  \caption{$3$ microphones are placed in a line, $10$ cm apart from each other. The average error is $55.05$ cm}
+  \label{fig:sim_hm_3_line}
+\end{figure}
+
+In fig~\ref{fig:sim_hm_3_line}, we experimented with an extreme arrangement where all three microphones are placed on a line, $10$ cm apart from each other.
+Looking at the figure above, we see that the intersection area is greater than $3000 \mathrm{cm}^2$ for points along the x axis.
+The average error distance is $55.05$ cm, which is not accurate enough for sub-meter localization.
+
+\subsection{Equilateral Triangular Configuration}
+
+\begin{figure*}[h!]
 \centering
@@
   \begin{subfigure}[]{.48\textwidth}
-    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_p1}
-    \caption{Another microphone is added at origin. The average error is $17.1$ cm}
-    \label{fig:sim_hm_3_p1}
-  \end{subfigure}
-  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_2x}
@@
   \end{subfigure}
+\end{figure*}
+
+Fig~\ref{fig:sim_hm_3} shows the accuracy when microphones are placed at the three vertices of a $20$ cm equilateral triangle.
+The region inside the array has good accuracy.
+However, for regions along the line connecting any two microphones, the accuracy drops significantly.
+The average distance error across the region is $18.6$ cm.
+
+To evaluate the array size's impact on accuracy, the size of the original array (as in fig~\ref{fig:sim_hm_3}) is increased by a factor of $2$.
+The result is presented in fig~\ref{fig:sim_hm_3_2x}.
+The overall uncertainty area decreased across the region.
+The average error distance improved to $10.04$ cm.
+By comparing with the previous result, it shows that increasing array size is effective in improving the overall accuracy.
+
+\section{More than three}
+
+\begin{figure*}[h!]
+  \centering
   \begin{subfigure}[]{.48\textwidth}
-    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_line}
-    \caption{$3$ microphones are placed in a line, $10$ cm apart from each other. The average error is $55.05$ cm}
-    \label{fig:sim_hm_3_line}
+    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_p1}
+    \caption{Another microphone is added at origin. The average error is $17.1$ cm}
+    \label{fig:sim_hm_3_p1}
   \end{subfigure}
@@
 
-Each microphone pair generates a hyperbolic region, and the source location is in the intersection of these regions.
-The area of the intersection region is a measure of the localization accuracy: the smaller the area, the more certain we are about the source location.
-To evaluate an array's accuracy in a region, we can place sound source at predetermined grid points in the region and look at the intersection area for each tested point in the grid.
-The center location of the intersection region can be used as the localization estimate to calculate localization error.
-Results for a few different microphone array configurations are presented in fig~\ref{fig:sim_hm}.
-
-Fig~\ref{fig:sim_hm_3} shows the accuracy when microphones are placed at the three vertices of a $20$ cm equilateral triangle.
-The region inside the array has good accuracy.
-However, for regions along the line connecting any two microphones, the accuracy drops significantly.
-The average error across the region is $18.6$ cm.
-
-To evaluate how adding one microphone (without increasing the array size) improves accuracy, another microphone is added to the array at $(0,0)$.
+To evaluate how adding one microphone (without increasing the array size) improves accuracy, another microphone is added at $(0,0)$ for the array described in fig~\ref{sim:hm_3}.
 Result is presented in fig~\ref{fig:sim_hm_3_p1}.
 Addition of the new microphone only slightly improved the accuracy around the array region.
-The average error dropped from $18.6$ cm to $17.1$ cm.
-Regions near lines connecting microphones still have significantly larger uncertainty region.
-This result suggests that adding more microphones in a small microphone array (without increasing the array size) is not an effective method to improve the localization accuracy.
-
-To evaluate the array size's impact on accuracy, the size of the original array (as in fig~\ref{fig:sim_hm_3}) is increased by a factor of $2$.
-The result is presented in fig~\ref{fig:sim_hm_3_2x}.
-The overall uncertainty area decreased across the region.
-The average error improved to $10.04$ cm.
-By comparing with the previous result, it shows that increasing array size is more effective in improving accuracy than adding more microphones.
-
-In fig~\ref{fig:sim_hm_3_line}, we experimented with an extreme arrangement where all three microphones are placed on a line, $10$ cm apart from each other.
-Error heatmap shows high uncertainty on the x axis, and the overall accuracy is much worse than that with three microphones placed in a triangle.
-The average error is $55.05$ cm.
+The average distance error dropped from $18.6$ cm to $17.1$ cm.
+Regions near lines connecting microphones still have significantly large intersection area.
+This result suggests that adding more microphones in a small microphone array (without increasing the array size) is not an effective method for improving the localization accuracy.
 
@@
 With this configuration, accuracy is consistently good across the region.
-The average error is $0.05$ cm.
+The average distance error is $0.05$ cm.
 However, placing microphones far apart at corners of the region requires accurate placement of all four individual microphones.
@@
 The result indicates that this configuration has good accuracy when source is close to the arrays.
-Accuracy decreases as sound source moves outside of the one meter by one meter region.
+We observe that the localization accuracy decreases as the sound source moves outside of the region one meter by one meter away from the two arrays.
 The average error is $2.60$ cm.
 
-With the simulation results, we decided to build the two array system as described in fig~\ref{fig:sim_hm_2_array}.
-The setup is reasonably portable (compared to fig~\ref{fig:sim_hm_4}), while at the same time having significantly better accuracy compared to one array systems.
+\section{Discussion}
+
+Looking at the simulation results on localization accuracy for different microphone array size and configurations, we see that the further away the microphones are placed from each other the higher the localization accuracy.
+However, as we increase the distance between each microphone pairs, we size of the overall array increases and the system becomes less portable.
+In the end we decided to build the two array system as described in fig~\ref{fig:sim_hm_2_array}.
+The setup is reasonably portable (compared to fig~\ref{fig:sim_hm_4}), yet it can achieve on average high localization accuracy within a region one meter by one meter close the setup.
+Since our target application is HCI, an one meter by one meter region will be adequate for this purpose and it will be used for all the following experiments.

### 2015-06-19 reports/paper/experiment.tex
thesis update

@@
 The top left plot in the figure shows the average error with different movement filters applied as a function of window size.
-The error decreases as window size increases and until the window size exceeds around $10$ ms.
-When the window size is $12$ cm, the localization accuracy is similar among raw output, Kalman filtering and averaging filtering.
-The bottom left plot shows the tracked movement for Averaging filter and the bottom right filter shows the tracked movement for Kalman filtering.
+The error decreases as window size increases until the window size exceeds around $10$ ms.
+When the window size is $12$ cm, the localization accuracy is similar among raw output, Kalman filtering output and averaging filtering output.
+The bottom left plot shows the tracked movement for averaging filter and the bottom right filter shows the tracked movement for Kalman filtering.
 Kalman filtering result is smoother while the averaging tracking stays closer to the ground truth circle.
-Compared to the experiment in fig~\ref{fig:circle_wnf}, where the same sound source is played but at a faster rotating speed of $20$ cm per second, we find that the performance is equally well between these two movement speeds.
+Compared to the experiment in fig~\ref{fig:circle_wnf}, where the same sound source is played but rotated at a faster linear speed of $20$ cm per second, we find that the performance is good well between these two movement speeds.
 
@@
 The low amplitude intervals in Music B affects the performance significantly.
-The average localization error is $2.9$.
+The average localization error is $2.9$ cm.
 The ``blank'' regions in the music can also be visually seen from the top right plot.
-The top left plot shows that the Kalman filtering still performs better than of the white noise, and behaves similar to the case with Music A.
+The top left plot shows that the Kalman filtering still performs better than raw output.
+The performance improvement of Kalman filtering is similar to that with Music A.
 Averaging filtering produces lowest average error.
-Compared to the faster movement experiment of the same sound source, we find that the localization error is similar between these two movement speeds.
+Compared to the faster movement experiment of the same sound source (shown in fig~\ref{fig:circle_mbf}), we find that the localization error is similar between these two movement speeds.
 
@@
 
-Fig~\ref{fig:circle_wnn} to \ref{fig:circle_musicbn} shows results for experiments at normal speed, and Fig~\ref{fig:circle_wnf} to~\ref{fig:circle_musicbf} shows results at fast speed.
-By comparing localization error for each audio source between normal movement speed and fast movement speed, we find that localization error does not increase when the source movement speed increased from $10$ cm to $20$ cm.
-For example fig~\ref{fig:circle_musican} and fig~\ref{fig:circle_musicaf} shows that localization error is $1.289$ cm at normal movement speed and $1.291$ cm at fast movement speed.
+\subsection{Discussion}
 
-For normal speed movement tracking, localization error is $0.9$ cm for white noise (Fig~\ref{fig:circle_wnn}), $1.29$ cm for Music A (Fig~\ref{fig:circle_musican}), and $2.9$ cm for Music B (Fig~\ref{fig:circle_musicbn}).
-Localization accuracy is the best for white noise source, and worst for Music B.
-This is consistent with our expectation because low amplitude regions in Music B would cause arrays to lose track of where the source is.
-This can also be seen from the ``blank'' regions in fig~\ref{fig:circle_musicbn}.
+Different prefiltering options produce very similar result. GCC\_PHAT gives slightly better accuracy, but the different among unfiltered GCC, GCC\_PHAT and GCC\_PHAT\_SQRT is very small.
 
-\subsection{Discussion}
-It also shows that raw detection has the most amount of jiggling.
+By comparing experiments at normal speed (fig~\ref{fig:circle_wnn} to \ref{fig:circle_musicbn}) with experiments at fast speed (fig~\ref{fig:circle_wnf} to~\ref{fig:circle_musicbf}) fast speed, we found that the localization error does increase when the movement speed increased from $10$ cm per second to $20$ cm per second.
+.
+
+When white noise is used as the audio source, Kalman filtering and raw localization produces very similar accuracy, and averaging filter gives slightly better accuracy.
+However, when the audio source is Music A or Music B, Kalman filtering produces better accuracy compared to raw detection, but Averaging filter still gives the best result.
+In terms of smoothness between movement filters, it can be seen that raw detection has the most amount of jiggling.
 Kalman filter reduces the amount of jiggling from raw detection by combining past estimates with current estimates.
 Averaging filter has the least amount of jiggling.
-However, averaging filter averages detection outputs from past 0.5 seconds, which makes the filtered output lag the real movement.
+However, averaging filter averages detection outputs from past $0.5$ seconds, which makes the filtered output lag the real movement.

### 2015-06-19 reports/paper/experiment.tex
thesis update

@@
 
-Fig~\ref{fig:circle_man} shows the result when Music A is used as the sound source.
+Fig~\ref{fig:circle_musican} shows the result when Music A is used as the sound source.
 The top right plot shows that the array still tracks the movement well, but has a slightly larger deviation compared to that in fig~\ref{fig:circle_wnn}.
@@
 From the top left plot, it can be seen that the error still decreases with the window size.
-The performance improvement from raw output to Kalman filtering output is bigger than that with white noise (top left plot of fig~\ref{circle_wnn}).
+The performance improvement from raw output to Kalman filtering output is bigger than that with white noise (top left plot of fig~\ref{fig:circle_wnn}).
 Averaging filtering still produces lowest average error.
-Fig~\ref{fig:circle_maf} shows the result for the same sound source moved at twice the speed ($20$ cm per second).
+Fig~\ref{fig:circle_musicaf} shows the result for the same sound source moved at twice the speed ($20$ cm per second).
 The average localization error at the faster speed is $1.291$ cm.
-Compared to the experiment where the same sound source is placed but with a faster movement speed (fig~\ref{fig:circle_maf}, we found that the performance does not degrade as the movement speed increases from $10$ cm per second to $20$ cm per second.
+Compared to the experiment where the same sound source is placed but with a faster movement speed (fig~\ref{fig:circle_musicaf}, we found that the performance does not degrade as the movement speed increases from $10$ cm per second to $20$ cm per second.
 
@@
 
-Fig~\ref{fig:circle_mbn} shows the result when Music B is used as the sound source.
+Fig~\ref{fig:circle_musicbn} shows the result when Music B is used as the sound source.
 The low amplitude intervals in Music B affects the performance significantly.
@@
 Averaging filtering produces lowest average error.
-Compared to the faster movement experiment of the same sound source (shown in fig~\ref{fig:circle_mbf}), we find that the localization error is similar between these two movement speeds.
+Compared to the faster movement experiment of the same sound source (shown in fig~\ref{fig:circle_musicbf}), we find that the localization error is similar between these two movement speeds.
 
@@
 By comparing experiments at normal speed (fig~\ref{fig:circle_wnn} to \ref{fig:circle_musicbn}) with experiments at fast speed (fig~\ref{fig:circle_wnf} to~\ref{fig:circle_musicbf}) fast speed, we found that the localization error does increase when the movement speed increased from $10$ cm per second to $20$ cm per second.
-.
 
@@
 However, averaging filter averages detection outputs from past $0.5$ seconds, which makes the filtered output lag the real movement.
+
+\begin{figure}[h!]
+  \centering
+  \begin{subfigure}[]{.48\textwidth}
+    \includegraphics[width=\textwidth]{show_l.png}
+    \caption{drawing letter `L'}
+  \end{subfigure}
+  \begin{subfigure}[]{.48\textwidth}
+    \includegraphics[width=\textwidth]{show_m.png}
+    \caption{drawing letter `M'}
+  \end{subfigure}
+
+  \begin{subfigure}[]{.48\textwidth}
+    \includegraphics[width=\textwidth]{show_r.png}
+    \caption{drawing letter `R'}
+  \end{subfigure}
+  \begin{subfigure}[]{.48\textwidth}
+    \includegraphics[width=\textwidth]{show_c.png}
+    \caption{drawing letter `C'}
+  \end{subfigure}
+  \begin{subfigure}[]{.48\textwidth}
+    \includegraphics[width=\textwidth]{show_n.png}
+    \caption{drawing letter `N'}
+  \end{subfigure}
+  \begin{subfigure}[]{.48\textwidth}
+    \includegraphics[width=\textwidth]{show_b.png}
+    \caption{drawing letter `B'}
+  \end{subfigure}
+  \caption{Drawing letters. Green dots represent raw localization output and the red line is the output after averaging filtering.}
+  \label{fig:show_case}
+\end{figure}
+
+In figure~\ref{fig:show_case}, a few examples of drawing letters with music are presented.
+Green dots on the plots represent the raw localization output and the red line shows the movement output with averaging filtering (window size of $0.5$ seconds is used).
+The demonstrated letters are drawn with free hand, without a track guiding the movement.
+The accuracy is reasonably good, and each letter can be easily recognized.
+There is a bit of jiggling in the tracked movement.
+Part of the jiggling comes from the system, and the rest comes from the hand movement.

### 2015-06-19 reports/paper/prior.tex
thesis update

@@
 \section{Movement Tracking}
-a
-\subsection{Averaging filtering}
-a
-\subsection{Kalman filtering}
-a
+In the previous section, we discussed various ways of localizing acoustic source.
+For each received microphone data, the algorithm produces a point estimate of the source location.
+If we have multiple estimates of the source location, they can be intelligently combined to make better location prediction.
+
+In this section we describe two ways of combining past estimates to produce better estimates: Finite impulse response (FIR) filter and Kalman filter.
+
+\subsection{FIR filter}
+The impulse response is of finite duration for FIR filters.
+In our system, we only have access to location estimates from the past, we look at a special category of FIR filters: causal discrete-time FIR filters.
+In such systems, the output $y[n]$ is a linear weighted combination of past $N+1$ estimates from input $x[n]$:
+\begin{eqnarray}
+y[n] & = & b_0x[n] + b_1x[n-1] + \cdots + b_Nx[n-N]\\
+& = & \sum_{i=0}^N b_i x[n-i]\\
+\end{eqnarray}
+where
+\begin{itemize}
+\item $y[n]$ is the output sequence
+\item $x[n]$ is the input sequence
+\item $N$ is the filter order
+\item $b_i$ is the impulse response
+\end{itemize}
+
+By controlling the impulse response $b_i$, we specify how the past few data should be blended to produce the output.
+
+When $b_i=\frac{1}{N+1}$, each of the previous $N+1$ localization estimate contributes equally to the output.
+In this case, the filter becomes a simple averaging filter (also called rolling mean).
+If the source does not move during the past $N+1$ estimates, and assuming each estimate is an independent estimate of the true location with an additive Gaussian noise.
+Then it can be shown that the output after filter is an unbiased estimation of the true source location.
+However, if the source has moved during the past $N+1$ estimates, the filter would output the mean location for the past $N+1$ estimated locations, which results in a ``lagging'' effect between the true source location and the system output.
+After the stopped moving, the filter output catches up with the source location.
+
+To reduce the ``lagging'' effect ehxibited by the averaging filter, we can assign higher weights to more recent estimates.
+Recent estimates gets a higher contribution to the output location.
+However, if the localization system
+
+\subsection{Kalman filter}

### 2015-06-19 reports/paper/prior.tex
thesis update

@@
 To reduce the ``lagging'' effect ehxibited by the averaging filter, we can assign higher weights to more recent estimates.
-Recent estimates gets a higher contribution to the output location.
-However, if the localization system
+Recent estimates gets a higher contribution to the output location, which makes the filtered out tracks more closely with the sound source.
+However, if the localization system is noisy with large estimation variance, then the error in the most recent estimate dominates the filtered output and might produces a large error.
+
+As a compromise between balancing the tradeoff between ``lagging'' problem from assigning equal weights to all estimates in the window and the error problem with assigning more weights.
+We will aim to design the system so that it produces estimates at a fast rate, so that the source would not move a large distance and it is safe to average the past a few estimates.
 
 \subsection{Kalman filter}
+The Kalman filter is a recursive filter where input data can be efficiently combined to produce online prediction.
+If all noise are Gaussian, the Kalman filter is a statistically optimum filter that minimizes squared error of the estimated parameters.
+Even if the noise is not Gaussian, given only first two statistics of noise, Kalman filter is the best linear estimator[x].
+Due to its statitical optimality and its recursive nature that enables online prediction, the Kalman filter has found applications in a wide range of areas.
+It has been used to track aircraft using RADAR, and to track Robot with sensors and beacons.
+
+The Kalman filter update equations are:
+\begin{eqnarray}
+\hat{x}_k^- & = & A\hat{x}_{k-1} + Bu_k\\
+P_k^- & = & AP_{k-1}A^T + Q\\
+K_k & = & P_k^- H^T(HP_k^-H^T+R)^{-1}\\
+\hat{x}_k & = & \hat{x}_k^- + K_k(z_k-H\hat{x}_k^-) \\
+P_k & = & (1-K_kH)P_k^-
+\end{eqnarray}
+where:
+\begin{itemize}
+\item $\hat{x}_k^-$ is
+\item $P_k^-$ is
+\item $K_k$ is
+\item $hat{x}_k$ is
+\item $P_k$ is
+\end{itemize}

### 2015-06-19 reports/paper/experiment.tex
update

@@
 
-In the movement tracking experiments described above, we have the ground truth location of the circle, but not the exact location of the audio source at each moment during the movement.
+In the movement tracking experiments described above, we know the ground truth location of the circle, but not the exact location of the audio source at each moment during the movement.
 Therefore, the error is reported as the distance between the localized point to its closest point on the ground truth circle.
@@
 The error decreases as window size increases and plateaus after the window size exceeds around $10$ millisecond.
-The lowest error achieved is $2.53$ centimeters, which occurred when the window size is set to 12 millisecond and GCC\_PHAT is used for TDOA estimation.
+The lowest error achieved is $2.53$ centimeters, which occurred when the window size is set to 12 millisecond and when GCC\_PHAT is used for TDOA estimation.
 The performance differences among GCC, GCC\_PHAT and GCC\_PHAT\_SQRT is small.
@@
 Kalman filtering result is smoother while the averaging tracking stays closer to the ground truth circle.
-Compared to the experiment in fig~\ref{fig:circle_wnf}, where the same sound source is played but rotated at a faster linear speed of $20$ cm per second, we find that the performance is good well between these two movement speeds.
+Compared to the experiment in fig~\ref{fig:circle_wnf}, where the same sound source is played but rotated at a faster linear speed of $20$ cm per second, we find that the performance is equally good between these two movement speeds.
 
@@
 Fig~\ref{fig:circle_musican} shows the result when Music A is used as the sound source.
-The top right plot shows that the array still tracks the movement well, but has a slightly larger deviation compared to that in fig~\ref{fig:circle_wnn}.
+The top right plot shows that the array still tracks the movement well, but has a slightly larger deviation compared to that in fig~\ref{fig:circle_wnn} when white noise was used.
 The average localization error for Music A is $1.289$ cm.
@@
 The average localization error at the faster speed is $1.291$ cm.
-Compared to the experiment where the same sound source is placed but with a faster movement speed (fig~\ref{fig:circle_musicaf}, we found that the performance does not degrade as the movement speed increases from $10$ cm per second to $20$ cm per second.
+Comparing the two experiments where music A was used as the sound source but with different movement speed, we find that the performance does not degrade as the movement speed increases from $10$ cm per second to $20$ cm per second.
 
@@
 Averaging filtering produces lowest average error.
-Compared to the faster movement experiment of the same sound source (shown in fig~\ref{fig:circle_musicbf}), we find that the localization error is similar between these two movement speeds.
+Comparing to the faster movement experiment of the same sound source (shown in fig~\ref{fig:circle_musicbf}), we find that the localization error is similar between these two movement speeds.
 
@@
 
-Different prefiltering options produce very similar result. GCC\_PHAT gives slightly better accuracy, but the different among unfiltered GCC, GCC\_PHAT and GCC\_PHAT\_SQRT is very small.
+Different prefiltering options produce very similar result. GCC\_PHAT gives slightly better accuracy, but the difference among unfiltered GCC, GCC\_PHAT and GCC\_PHAT\_SQRT is very small.
 
-By comparing experiments at normal speed (fig~\ref{fig:circle_wnn} to \ref{fig:circle_musicbn}) with experiments at fast speed (fig~\ref{fig:circle_wnf} to~\ref{fig:circle_musicbf}) fast speed, we found that the localization error does increase when the movement speed increased from $10$ cm per second to $20$ cm per second.
+By comparing experiments at normal speed (fig~\ref{fig:circle_wnn} to \ref{fig:circle_musicbn}) with experiments at fast speed (fig~\ref{fig:circle_wnf} to~\ref{fig:circle_musicbf}), we find that the localization error does not increase when the movement speed increased from $10$ cm per second to $20$ cm per second.
 
-When white noise is used as the audio source, Kalman filtering and raw localization produces very similar accuracy, and averaging filter gives slightly better accuracy.
-However, when the audio source is Music A or Music B, Kalman filtering produces better accuracy compared to raw detection, but Averaging filter still gives the best result.
-In terms of smoothness between movement filters, it can be seen that raw detection has the most amount of jiggling.
+When white noise is used as the audio source, Kalman filtering and raw localization produces very similar accuracy, and averaging filter gives slightly higher accuracy.
+However, when the audio source is changed to Music A or Music B, Kalman filtering produces better accuracy compared to raw detection, but Averaging filter still gives the best result.
+Looking at the smoothness of the movement path after applying different movement filters, it can be seen that raw detection has the most amount of jiggling.
 Kalman filter reduces the amount of jiggling from raw detection by combining past estimates with current estimates.
@@
 There is a bit of jiggling in the tracked movement.
-Part of the jiggling comes from the system, and the rest comes from the hand movement.
+Part of the jiggling can be attributed to the noise from system output, and the rest comes from the hand movement.

### 2015-06-19 reports/paper/system.tex
thesis update
update

@@
 The end system has two arrays, each with three microphones mounted on the vertices of a $20$ cm equilateral triangle.
-We used omni-directional foil electret condenser microphones due to their low cost.
+We used omni-directional foil electret condenser microphone due to its low cost and small size.
 The microphone has a frequency range of $100$ to $10K$ Hz, and a minimum SNR of 58dB[x].
-We used operational amplifer OPA344 from Texas Instrument to amplify the microphone output by a factor of 100.
-A micro-controller is attached to one of the vertices.
+We also used operational amplifer OPA344 from Texas Instrument to amplify the microphone output by a factor of 100, so the received signal can be easily picked up by the analog-to-digital converter (ADC) module installed on the micro-controller.
+The micro-controller board used in this project is \emph{teensy 3.1}, and it is attached to one of the vertices of the triangle.
 Fig~\ref{fig:setup_array} shows a picture of the array setup.
-The micro-controller baord in this project is \emph{teensy 3.1}.
-It contains a which has a onboard RAM of $64$k.
-ADC is capable of sampling at $500$kHz~\cite{tdoa:micloc, sys:teensy}.
-In this project, the micro-controller collects microphone data on all three channels for $12$ milliseconds and then sends the recorded data to a computer through the USB port for localization.
+The micro-controller contains an onboard RAM of $64$k, and an ADC module capable of sampling at $500$kHz~\cite{tdoa:micloc, sys:teensy}.
+In this project, the micro-controller collects microphone data on all three channels for a duration $12$ milliseconds and then sends the recorded data to a computer through the USB port for localization.
 
-\begin{figure}[]
+\begin{figure}[h!]
   \centering
@@
 To handle the uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation output(equation~\ref{eq:gcc2}) as a measure of the likelihood of different arrival time differences.
-For each microphone array, we build a heatmap for the region.
-The intensity at each point represents the likelihood of it being the source.
-For each point on the grid, the theoretical TDOA to each microphone pair can be precomputed.
-Then the heatmap can be generated by going through points on the grid and performing a lookup using equation~\ref{eq:gcc2}.
+Each index $i$ from the cross-correlation vector denotes the time delay across the two microphones receiving the acoustic signal, and the cross-correlation value $v$[i] at each index $i$ denotes the likelihood of the time delay being $i$.
+
+For each microphone array, we build a heatmap of likelihood for the region.
+The intensity at each point on the heatmap represents the likelihood of it being the source.
+To generate the likelihood heatmap for an microphone array, we apply the following algorithm.
+For each point ($x$,$y$) on the grid, the theoretical TDOA to each microphone pair can be precomputed using:
+\[
+ D_{m_1,m_2}(x,y) =  \frac{((x-x_1)^2 + (y-y_1)^2)^{0.5} - ((x-x_2)^2 + (y-y_2)^2)^{0.5}}{v}
+\]
+where ($x_1, y_1$) and ($x_2, y_2$) are the locations of the microphone pair and $v$ is the speed of sound.
+Then the heatmap can be generated by going through all the points on the grid and performing a lookup using equation~\ref{eq:gcc2}.
 With three microphones $m_1,m_2,$ and $m_3$, there are three microphone pairs: $m_1m_2,m_1m_3,$ and $m_2m_3$.
@@
 
-\begin{figure}[]
+To see the effect of accuracy improvement using multiple arrays, fig~\ref{fig:liklihood} shows a real life localization where the source is placed at $(0$ cm$,-30$ cm$)$.
+The top two figures show the individual likelihood maps produced by single microphone arrays.
+We can see that the individual arrays can give accurate angle estimate, but have high uncertainty in distance estimate.
+The bottom figure shows the combined likelihood map according to equation~\ref{eq:combine_l}.
+The combined likelihood map demonstrated that by merging estimates from two arrays the system is able to perform more accurate localization.
+
+\begin{figure}[h!]
   \centering
@@
 
-To see the effect of accuracy improvement using multiple arrays, fig~\ref{fig:liklihood} shows a real life localization where the source is placed at $(0$ cm$,-30$ cm$)$.
-It shows the individual likelihood map from each array and also the combined likelihood map according to equation~\ref{eq:combine_l}.
-Individual arrays give accurate angle estimate, but have high uncertainty in distance estimate.
-The combined likelihood map demonstrated that by merging estimates from two arrays the system is able to perform more accurate localization.
-
 From a timing point of view, the micro-controller spends $12$ milliseconds on sampling the microphone data before sending it to a computer for processing.

### 2015-06-19 reports/paper/prior.tex
thesis update

@@
 
-The Kalman filter update equations are:
+The Kalman filter has a prdict step:
 \begin{eqnarray}
-\hat{x}_k^- & = & A\hat{x}_{k-1} + Bu_k\\
-P_k^- & = & AP_{k-1}A^T + Q\\
-K_k & = & P_k^- H^T(HP_k^-H^T+R)^{-1}\\
-\hat{x}_k & = & \hat{x}_k^- + K_k(z_k-H\hat{x}_k^-) \\
-P_k & = & (1-K_kH)P_k^-
+\hat{x}_{k+1}^- & = &F_k \hat{x}_k + B_ku_k\\
+P_{k+1}^- & = & F_kP_kF_k^T + Q_k\\
+\end{eqnarray}
+and an update step:
+\begin{eqnarray}
+y_k & = & z_k - H_k\hat{x}_k^-\\
+S_k & = & H_kP_k^-H_k^T + R_k \\
+K_k & = & P_k^-H_k^TS_k^{-1}\\
+\hat{x}_k & = & \hat{x}_k^- + K_ky\\
+P_k & = & (I-K_kH_k)P_k^-\\
 \end{eqnarray}
@@
 \begin{itemize}
-\item $\hat{x}_k^-$ is
-\item $P_k^-$ is
-\item $K_k$ is
-\item $hat{x}_k$ is
-\item $P_k$ is
+\item $x$ is the state variable
+\item $P$ is the covariance on the state variable $x$
+\item $F_k \hat{x}_k + B_k u_k$ describes the station transition form time $k$ to $k+1$.
+The Kalman filter is a general framework which allows not only tracking but also controlling the object.
+I these situations $u_k$ is the control input and $B$ describes the control input model.
+For our application, we are only tracking and not controlling the movement, $B_ku_k = 0$.
+\item $z$ is the measurement.
+\item $y$ is the residue in the measurement sapce. 
+\item $H$ is the measurement function that transforms from state space to measurement space.
+\item $R$ is the measurement noise matrix that models sensor noise as a covariance matrix
+\item $Q$ is the unmodeled noise in the process.
 \end{itemize}

### 2015-06-19 reports/paper/prior.tex
update

@@
 If all noise are Gaussian, the Kalman filter is a statistically optimum filter that minimizes squared error of the estimated parameters.
-Even if the noise is not Gaussian, given only first two statistics of noise, Kalman filter is the best linear estimator[x].
+Even if the noise is not Gaussian, given only the first two statistics of noise, Kalman filter is the best linear estimator[x].
 Due to its statitical optimality and its recursive nature that enables online prediction, the Kalman filter has found applications in a wide range of areas.
 It has been used to track aircraft using RADAR, and to track Robot with sensors and beacons.
+
+Kalman filter use observed variables to infer hidden variables and use that to help predict the next state.
+In our system, observed variables are the $x$ and $y$ coordinates of the system, and we also model unobserved motion variables velocity $v_x, v_y$ and acceleration variables $a_x, a_y$.
+Then the state variables we are tracking ca be represented as:
+\[
+\left[\begin{array}{c}
+x\\
+y\\
+a_x\\
+a_y\\
+\end{array}\right]
+\]
 
@@
 \item $R$ is the measurement noise matrix that models sensor noise as a covariance matrix
-\item $Q$ is the unmodeled noise in the process.
+\item $Q$ models the process noise
 \end{itemize}
+
+In our system,

### 2015-06-19 reports/paper/prior.tex
update

--- 2015-06-19 reports/paper/prior.tex
@@
 If all noise are Gaussian, the Kalman filter is a statistically optimum filter that minimizes squared error of the estimated parameters.
-Even if the noise is not Gaussian, given only first two statistics of noise, Kalman filter is the best linear estimator[x].
-Due to its statitical optimality and its recursive nature that enables online prediction, the Kalman filter has found applications in a wide range of areas.
+Even if the noise is not Gaussian, given only the first two statistics of noise, Kalman filter is the best linear estimator[x].
+Due to its statistical optimality and its recursive nature that enables online prediction, the Kalman filter has found applications in a wide range of areas.
 It has been used to track aircraft using RADAR, and to track Robot with sensors and beacons.
 
-The Kalman filter has a prdict step:
+Kalman filter uses observed variables to infer hidden variables and use them to help predict the next state.
+In our system, observed variables $z$ are the $(x,y)$ coordinates of the localized acoustic source:
+\[
+z = \left[\begin{array}{c}
+x\\
+y\\
+\end{array}\right]
+\]
+We can also model unobserved motion variables such as velocity $(\dot{x}, \dot{y})$ and acceleration $(\ddot{x}, \ddot{y})$.
+Then the state variables we are tracking $\mathbf{x}$ can be represented as:
+\[
+\mathbf{x} = \left[x, y, \dot{x}, \dot{y}, \ddot{x}, \ddot{y}\right]^T
+\]
+Note that by modeling up to acceleration, we are implicitly assuming higher order motion variables are constant (such as jerk).
+Kalman filter includes a process noise term $Q$ that can be used to partially reduce this effect, but there will exist systematic tracking bias there is acceleration change.
+
+At any time instant, Kalman filter can use current state information to infer the predicted next state $\mathbf{x}^-$ and the predicted uncertainty $P^-$:
 \begin{eqnarray}
-\hat{x}_{k+1}^- & = &F_k \hat{x}_k + B_ku_k\\
-P_{k+1}^- & = & F_kP_kF_k^T + Q_k\\
+\mathbf{x}^- & = &F \mathbf{x} + B_ku_k\\
+P^- & = & FPF^T + Q\\
 \end{eqnarray}
+where
+\begin{itemize}
+\item F is the state transition matrix. In our system, the next coordinate output can be computed with laws of physics using current position, velocity and acceleration:
+\[
+F  =  \left(\begin{array}{cccccc}
+1 & 0 & \frac{1}{2} & 0 & \frac{1}{2} & 0\\
+1 & 0 & \frac{1}{2} & 0 & \frac{1}{2} & 0 \\
+\end{array}\right)
+\]
+\item $F_k \hat{x}_k + B_k u_k$ describes the station transition form time $k$ to $k+1$.
+The Kalman filter is a general framework which allows not only tracking but also controlling the object.
+I these situations $u_k$ is the control input and $B$ describes the control input model.
+For our application, we are only tracking and not controlling the movement, $B_ku_k = 0$.
+\end{itemize}
 and an update step:
@@
 \item $P$ is the covariance on the state variable $x$
-\item $F_k \hat{x}_k + B_k u_k$ describes the station transition form time $k$ to $k+1$.
-The Kalman filter is a general framework which allows not only tracking but also controlling the object.
-I these situations $u_k$ is the control input and $B$ describes the control input model.
-For our application, we are only tracking and not controlling the movement, $B_ku_k = 0$.
 \item $z$ is the measurement.

### 2015-06-19 reports/paper/conclusion.tex
update

@@
 In this thesis, we described and built an inexpensive, portable yet reasonably accurate sound localization system with two microphone arrays.
-We analyzed different array architectures and showed that the localization accuracy varies with the array geometry and the location of the sound source (both the distance and the angle).
+We have analyzed different array architectures and showed that the localization accuracy varies with the array geometry and the location of the sound source (in terms of both the distance and the angle).
 In particular, regions close to the line connecting two microphones are generally more difficult to localize than regions close the line bisecting two microphones.
@@
 
-This system can be used in HCI applications that incorporate sound position and movement information.
+This system we have built can be installed in HCI applications that incorporate sound position and movement information.
 One can use this system to build virtual drawing applications where the user can draw with music, without physically touching the computer.

### 2015-06-19 reports/paper/introduction.tex
update

--- 2015-06-17 reports/paper/introduction.tex
@@
 The fact that modern consumer devices such as mobile phones and tablets are commonly equipped with Wi-Fi and bluetooth modules also made this approach particular attractive because no hardware needs to be installed on the user end.
-Recent applications including serving advertisements and coupons to consumers in shopping centers based on the customer's location.
-In these systems, signal strength received from different base stations are used for the estimation of the device location.
-However, their reported accuracy are in the range of $1$ to $5$ meters~\cite{intro:blue, intro:loc}, which is not enough for local movement tracking.
+Relevant commercial user case includes applications that serve advertisements and coupons to consumers in shopping centers based on the customer's location.
+In these systems, device signal strength received at different base stations are used for the estimation of the user device location.
+However, their reported accuracy are in the range of $1$ to $5$ meters~\cite{intro:blue, intro:loc}, which is not accurate enough for local movement tracking.
 
 In this work, we have built a source localization system that is portable, inexpensive, yet reasonably accurate for localization in a small area.
+To test the accuracy of our localization system, we perform acoustic localization experiments in an one meter by one meter area.
+The experiment area size is adequate for our purposes because human arm movement will generally be constrained in this area.
+Our experiment show that our system localizes both static and moving acoustic source with good accuracy in this area.
 Our system is built with inexpensive electret microphones mounted on portable frames.
@@
 One can use this system to build virtual drawing applications where the user can draw with music without physically touching the computer.
-One can also use this system to design interactive AI games with physical pieces.
+One can also use this system to design interactive artificial intelligence (AI) games with audible physical game pieces.
 For example, a chess game with physical pieces can be developed where each piece is equipped with a motor and a music tag.
@@
 Another similar example is to build an AI toy car racing game, where the player controls one car and the computer controls another.
-With real time location information, the AI engine can compete with the player on a racing track.
+With real time location information of both cars, the AI engine can compete with the player on a racing track.
 
 This system can also be used to in Augmented Reality (AR) applications.
-For example, the user can wear a music tag on one finger, and then the system would be able to track the finger movement.
+For example, the user can wear a music tag on one finger, and then the system would be able to track the user's finger movements.
 This particular setup can be used as a virtual mouse for wearable technologies such as Google Glass.
@@
 We demonstrated that a two array system outperforms a single array system of similar physical dimension.
-In Chapter IV, we first presented the chosen architecture along with hardware details, followed by experiment details and results were presented.
+In Chapter IV, we first presented the chosen architecture along with hardware details, followed by experiment details and results.

### 2015-06-19 reports/paper/prior.tex
update

--- 2015-06-19 reports/paper/prior.tex
@@
 
-\section{Movement Tracking}
-In the previous section, we discussed various ways of localizing acoustic source.
-For each received microphone data, the algorithm produces a point estimate of the source location.
-If we have multiple estimates of the source location, they can be intelligently combined to make better location prediction.
-
-In this section we describe two ways of combining past estimates to produce better estimates: Finite impulse response (FIR) filter and Kalman filter.
-
-\subsection{FIR filter}
-The impulse response is of finite duration for FIR filters.
-In our system, we only have access to location estimates from the past, we look at a special category of FIR filters: causal discrete-time FIR filters.
-In such systems, the output $y[n]$ is a linear weighted combination of past $N+1$ estimates from input $x[n]$:
-\begin{eqnarray}
-y[n] & = & b_0x[n] + b_1x[n-1] + \cdots + b_Nx[n-N]\\
-& = & \sum_{i=0}^N b_i x[n-i]\\
-\end{eqnarray}
-where
-\begin{itemize}
-\item $y[n]$ is the output sequence
-\item $x[n]$ is the input sequence
-\item $N$ is the filter order
-\item $b_i$ is the impulse response
-\end{itemize}
-
-By controlling the impulse response $b_i$, we specify how the past few data should be blended to produce the output.
-
-When $b_i=\frac{1}{N+1}$, each of the previous $N+1$ localization estimate contributes equally to the output.
-In this case, the filter becomes a simple averaging filter (also called rolling mean).
-If the source does not move during the past $N+1$ estimates, and assuming each estimate is an independent estimate of the true location with an additive Gaussian noise.
-Then it can be shown that the output after filter is an unbiased estimation of the true source location.
-However, if the source has moved during the past $N+1$ estimates, the filter would output the mean location for the past $N+1$ estimated locations, which results in a ``lagging'' effect between the true source location and the system output.
-After the stopped moving, the filter output catches up with the source location.
-
-To reduce the ``lagging'' effect ehxibited by the averaging filter, we can assign higher weights to more recent estimates.
-Recent estimates gets a higher contribution to the output location, which makes the filtered out tracks more closely with the sound source.
-However, if the localization system is noisy with large estimation variance, then the error in the most recent estimate dominates the filtered output and might produces a large error.
-
-As a compromise between balancing the tradeoff between ``lagging'' problem from assigning equal weights to all estimates in the window and the error problem with assigning more weights.
-We will aim to design the system so that it produces estimates at a fast rate, so that the source would not move a large distance and it is safe to average the past a few estimates.
-
-\subsection{Kalman filter}
-The Kalman filter is a recursive filter where input data can be efficiently combined to produce online prediction.
-If all noise are Gaussian, the Kalman filter is a statistically optimum filter that minimizes squared error of the estimated parameters.
-Even if the noise is not Gaussian, given only first two statistics of noise, Kalman filter is the best linear estimator[x].
-Due to its statitical optimality and its recursive nature that enables online prediction, the Kalman filter has found applications in a wide range of areas.
-It has been used to track aircraft using RADAR, and to track Robot with sensors and beacons.
-
-The Kalman filter has a prdict step:
-\begin{eqnarray}
-\hat{x}_{k+1}^- & = &F_k \hat{x}_k + B_ku_k\\
-P_{k+1}^- & = & F_kP_kF_k^T + Q_k\\
-\end{eqnarray}
-and an update step:
-\begin{eqnarray}
-y_k & = & z_k - H_k\hat{x}_k^-\\
-S_k & = & H_kP_k^-H_k^T + R_k \\
-K_k & = & P_k^-H_k^TS_k^{-1}\\
-\hat{x}_k & = & \hat{x}_k^- + K_ky\\
-P_k & = & (I-K_kH_k)P_k^-\\
-\end{eqnarray}
-where:
-\begin{itemize}
-\item $x$ is the state variable
-\item $P$ is the covariance on the state variable $x$
-\item $F_k \hat{x}_k + B_k u_k$ describes the station transition form time $k$ to $k+1$.
-The Kalman filter is a general framework which allows not only tracking but also controlling the object.
-I these situations $u_k$ is the control input and $B$ describes the control input model.
-For our application, we are only tracking and not controlling the movement, $B_ku_k = 0$.
-\item $z$ is the measurement.
-\item $y$ is the residue in the measurement sapce. 
-\item $H$ is the measurement function that transforms from state space to measurement space.
-\item $R$ is the measurement noise matrix that models sensor noise as a covariance matrix
-\item $Q$ is the unmodeled noise in the process.
-\end{itemize}
+\input{prior_movement.tex}

### 2015-06-19 reports/paper/prior_movement.tex
update

\section{Movement Tracking}
In the previous section, we discussed various ways of localizing acoustic source. For each received microphone data, the algorithm produces a point estimate of the source location. If we have multiple estimates of the source location, they can be intelligently combined to make better location prediction.

In this section we describe two ways of combining past estimates to produce better estimates: Finite impulse response (FIR) filter and Kalman filter.

\subsection{FIR filter}
The impulse response is of finite duration for FIR filters. In our system, we only have access to location estimates from the past, we look at a special category of FIR filters: causal discrete-time FIR filters. In such systems, the output $y[n]$ is a linear weighted combination of past $N+1$ estimates from input $x[n]$:
\begin{eqnarray}
y[n] & = & b_0x[n] + b_1x[n-1] + \cdots + b_Nx[n-N]\\
& = & \sum_{i=0}^N b_i x[n-i]\\
\end{eqnarray}
where
\begin{itemize}
\item $y[n]$ is the output sequence
\item $x[n]$ is the input sequence
\item $N$ is the filter order
\item $b_i$ is the impulse response
\end{itemize}

By controlling the impulse response $b_i$, we specify how the past few data should be blended to produce the output.

When $b_i=\frac{1}{N+1}$, each of the previous $N+1$ localization estimate contributes equally to the output. In this case, the filter becomes a simple averaging filter (also called rolling mean). If the source does not move during the past $N+1$ estimates, and assuming each estimate is an independent estimate of the true location with an additive Gaussian noise. Then it can be shown that the output after filter is an unbiased estimation of the true source location. However, if the source has moved during the past $N+1$ estimates, the filter would output the mean location for the past $N+1$ estimated locations, which results in a ``lagging'' effect between the true source location and the system output. After the stopped moving, the filter output catches up with the source location.

To reduce the ``lagging'' effect ehxibited by the averaging filter, we can assign higher weights to more recent estimates. Recent estimates gets a higher contribution to the output location, which makes the filtered out tracks more closely with the sound source. However, if the localization system is noisy with large estimation variance, then the error in the most recent estimate dominates the filtered output and might produces a large error.

As a compromise between balancing the tradeoff between ``lagging'' problem from assigning equal weights to all estimates in the window and the error problem with assigning more weights. We will aim to design the system so that it produces estimates at a fast rate, so that the source would not move a large distance and it is safe to average the past a few estimates.

\subsection{Kalman filter}
The Kalman filter is a recursive filter where input data can be efficiently combined to produce online prediction. If all noise are Gaussian, the Kalman filter is a statistically optimum filter that minimizes squared error of the estimated parameters. Even if the noise is not Gaussian, given only the first two statistics of noise, Kalman filter is the best linear estimator[x]. Due to its statistical optimality and its recursive nature that enables online prediction, the Kalman filter has found applications in a wide range of areas. It has been used to track aircraft using RADAR, and to track Robot with sensors and beacons.

Kalman filter uses observed variables to infer hidden variables and use them to help predict the next state. In our system, observed variables $z$ are the $(x,y)$ coordinates of the localized acoustic source:
\[
z = \left[\begin{array}{c}
x\\
y\\
\end{array}\right]
\]
We can also model unobserved motion variables such as velocity $(\dot{x}, \dot{y})$ and acceleration $(\ddot{x}, \ddot{y})$. Then the state variables we are tracking $\mathbf{x}$ can be represented as:
\[
\mathbf{x} = \left[x, y, \dot{x}, \dot{y}, \ddot{x}, \ddot{y}\right]^T
\]
Note that by modeling up to acceleration, we are implicitly assuming higher order motion variables are constant (such as jerk). Kalman filter includes a process noise term $Q$ that can be used to partially reduce this effect, but there will exist systematic tracking bias there is acceleration change.

At any time instant, Kalman filter can use current state information to infer the predicted next state $\mathbf{x}^-$ and the predicted uncertainty $P^-$:
\begin{eqnarray}
\mathbf{x}^- & = &F \mathbf{x} + B_ku_k\\
P^- & = & FPF^T + Q\\
\end{eqnarray}
where
\begin{itemize}
\item F is the state transition matrix. In our system, the next coordinate output can be computed with laws of physics using current position, velocity and acceleration:
\[
F  =  \left(\begin{array}{cccccc}
1 & 0 & \frac{1}{2} & 0 & \frac{1}{2} & 0\\
1 & 0 & \frac{1}{2} & 0 & \frac{1}{2} & 0 \\
\end{array}\right)
\]
\item $F_k \hat{x}_k + B_k u_k$ describes the station transition form time $k$ to $k+1$. The Kalman filter is a general framework which allows not only tracking but also controlling the object. I these situations $u_k$ is the control input and $B$ describes the control input model. For our application, we are only tracking and not controlling the movement, $B_ku_k = 0$.
\end{itemize}
and an update step:
\begin{eqnarray}
y_k & = & z_k - H_k\hat{x}_k^-\\
S_k & = & H_kP_k^-H_k^T + R_k \\
K_k & = & P_k^-H_k^TS_k^{-1}\\
\hat{x}_k & = & \hat{x}_k^- + K_ky\\
P_k & = & (I-K_kH_k)P_k^-\\
\end{eqnarray}
where:
\begin{itemize}
\item $x$ is the state variable
\item $P$ is the covariance on the state variable $x$
\item $z$ is the measurement.
\item $y$ is the residue in the measurement sapce. 
\item $H$ is the measurement function that transforms from state space to measurement space.
\item $R$ is the measurement noise matrix that models sensor noise as a covariance matrix
\item $Q$ models the process noise
\end{itemize}

In our system,

### 2015-06-19 reports/paper/prior.tex
update

@@
 \section{Point localization}
-Acoustic localization has been researched extensively in the literature.
+Acoustic localization techniques have been researched extensively in the literature.
 Localization techniques can be broadly categorized into Interaural Level Difference (ILD), Location Template Matching (LTM), and Time Difference of Arrival (TDOA) based approaches.
@@
 With multiple microphones, it is possible to infer the source location by comparing the signal intensity received at different microphones.
-Human's auditory system has used ILD cues to infer source direction~\cite{ild:2}, and this technique is most effective to localize high frequency sources, because they don't diffract effectively around the listener's head and produce a significant intensity difference.
+Human's auditory system has used ILD cues to infer source direction~\cite{ild:2}, and this technique is most effective when used to localize high frequency sources, because these sources don't diffract significantly around the listener's head and produce a significant intensity difference.
 
@@
 It can be shown that, on a 2D plane, points satisfy equation~\ref{eqn:ild} form a circle when $I_1 \ne I_2$ and form a line when $I_1 = I_2$~\cite{ild:1}.
-With two microphones all points on this curve generate the same intensity ratio and can not be distinguished from each other.
+With two microphones all points on this curve generate the same intensity ratio and they can not be distinguished from each other.
 Multiple approaches have been investigated to eliminate this ambiguity.
-\cite{ild:1} employed multiple microphones and used the intersection of circles from each microphone pair to estimate the source location.
-\cite{ild:3} combined ILD with Interaural Time Difference (ITD) to estimate the source direction (i.e azimuth).
+The authors in \cite{ild:1} employed multiple microphones and used the intersection of circles from each microphone pair to estimate the source location.
+While the authors in \cite{ild:3} combined ILD with Interaural Time Difference (ITD) to estimate the source direction (i.e azimuth).
 Instead of solving the intersection of circles, \cite{ild:4} employed machine learning techniques to automatically learn the mapping from ILD and ITD features to location coordinate.
-It uses four microphones and requires a training phase, during which the sound source is manually placed at predetermined locations for the system to learn the parameters that map from feature space to the sound location.
+Their technique uses four microphones and requires a training phase, during which the sound source is manually placed at predetermined locations for the system to learn the parameters that map from feature space to the sound location.
 
@@
 Any obstacle object between the sound source and any microphone would produce a significant distortion in the measured intensity ratio.
-We find this approach to be too restrictive for our system, since in an interactive system we can not control whether or not the user places any obstacle between the sound source and the microphones.
+We find this approach to be too restrictive for our purposes, since in an interactive system we do not control whether or not the user places any obstacle between the sound source and the microphones.
 
@@
 
-\cite{extended:tusi} and \cite{ltm:pham} investigated using max value from cross-correlation as a similarity measure to localize user taps on interactive surfaces.
-\cite{ltm:lpc} used L2 distance in the Linear Predictive Coding coefficient space as a similarity measure to localize taps on surfaces.
-\cite{ltm:tusi2} further explored accuracy improvement by using multiple templates for each location and speed improvement by merging multiple templates into one representative template.
+Authors in \cite{extended:tusi} and \cite{ltm:pham} have investigated using max value from cross-correlation as a similarity measure to localize user taps on interactive surfaces.
+In \cite{ltm:lpc}, the authors used L2 distance in the Linear Predictive Coding coefficient space as a similarity measure to localize taps on surfaces.
+Authors in \cite{ltm:tusi2} further explored accuracy improvement by using multiple templates for each location and speed improvement by merging multiple templates into one representative template.
 
 The requirement of having a template for each location to be detected makes this approach too restrictive for our project, since we want the localization to be continuous in a 2D region.
-\cite{ltm:tusi2} has investigated contiguous localization by linearly extrapolating between stored locations, but result has high error.
+The authors in \cite{ltm:tusi2} have also investigated into contiguous localization by linearly extrapolating between stored locations, but result has high error.
 Moreover, the need to recalibrate all locations during setup is too cumbersome for the end users in a portable system.
@@
 In \cite{tdoa:ppp}, authors used eight microphones mounted on the corners of a ping pong table to localize points where the ball hits the table.
-They used a threshold to determine the arrival time of acoustic signal.
+They used a pre-set threshold to determine the arrival time of acoustic signal.
 This approach works well in noise free environment but the performance degrades with background noise.
@@
 To make it more robust, authors in \cite{tdoa:mit3} and \cite{tdoa:mit4} extracted descriptive parameters for each significant peak(e.g., peak height, width, mean arrival time).
-The algorithm then used extracted parameters to predict arrival time with a second order polynomial, the parameters of which were fitted during calibration at fixed locations.
+Their algorithm then used the extracted parameters to predict arrival time with a second order polynomial, the parameters of which were fitted at fixed locations during a calibration phase.
 Authors in \cite{tdoa:mit5} have used similar techniques and built an interactive window by placing four microphones on four corners of a glass pane.
@@
 \end{eqnarray}
-The arrival time difference $t_0$ is the time shift $\tau$ that maximizes~\ref{eqn:gcc2}:
+The arrival time difference $t_0$ is the time shift $\tau$ that maximizes~(\ref{eqn:gcc2}):
 \begin{eqnarray}\label{eqn:gcc3}
@@
 
-The benefits of calculating cross-correlation in the frequency domain as shown in equation~\ref{eqn:gcc2} are two folds.
+The benefits of calculating cross-correlation in the frequency domain as shown in equation~(\ref{eqn:gcc2}) are two folds.
 The first benefit is to speedup the calculation.
@@
 With discrete signals of length $n$, it will take $O(n^2)$ number of calculations.
-Doing the same calculation in frequency domain, we need to transform the signal into frequency domain, multiple and sum the two transformed signal vectors once and then transform the result back to the time domain.
+Doing the same calculation in frequency domain, we first need to transform the signal into frequency domain, then multiply and sum the two transformed signal vectors once and then transform the result back to the time domain.
 Transforming a signal from time domain to frequency domain and back can be done efficiently with Fast Fourier Transform (FFT) and Inverse Fast Fourier Transform (IFFT), and the amount of calculation needed is $O(n\log n)$.
@@
 Therefore, the total calculation required to calculate cross-correlation using Fourier transform is $O(n\log n)$, which is asymptotically faster than calculating in the time domain.
-This calculation speedup is especially beneficial in real-time interactive systems since significant time lag would produce unsmooth user experience.
+This calculation speedup is particularly beneficial in real-time interactive systems since significant time lag would make the system less interactive to the user.
 
@@
 \]
-We can look at the real part of equation~\ref{eqn:phat1} only since both $x_1(t)$ and $x_2(t)$ are real valued signals.
+We can look at the real part of equation~(\ref{eqn:phat1}) only since both $x_1(t)$ and $x_2(t)$ are real valued signals.
 When $\tau$ is the true arrival time difference between $x_1(t)$ and $x_2(t)$, phase error $\Theta_\epsilon  = 0$, and $\cos(\Theta_\epsilon) = 1$.

### 2015-06-19 reports/paper/prior_movement.tex

@@
 \]
+Internally, the Kalman filter also keeps track of the uncertainty on the state variables. It is represented as a covariance matrix $P$ on state variables.
+
 Note that by modeling up to acceleration, we are implicitly assuming higher order motion variables are constant (such as jerk).
@@
 \begin{eqnarray}
-\mathbf{x}^- & = &F \mathbf{x} + B_ku_k\\
+\mathbf{x}^- & = &F \mathbf{x} + Bu\\
 P^- & = & FPF^T + Q\\
@@
 \begin{itemize}
-\item F is the state transition matrix. In our system, the next coordinate output can be computed with laws of physics using current position, velocity and acceleration:
+\item F is the state transition matrix. In our system, next coordinates $(x,y)$ can be computed with laws of physics using current position, velocity and acceleration:
 \[
 F  =  \left(\begin{array}{cccccc}
-1 & 0 & \frac{1}{2} & 0 & \frac{1}{2} & 0\\
-1 & 0 & \frac{1}{2} & 0 & \frac{1}{2} & 0 \\
+1 & 0 & \delta t & 0 & \frac{1}{2}\delta t^2 & 0 \\
+0 & 1 & 0 & \delta t & 0 & \frac{1}{2} \delta t^2 \\
+0 & 0 & 1 & 0 & \delta t & 0\\
+0 & 0 & 0 & 1 & 0 & \delta t\\
+0 & 0 & 0 & 0 & 1 & 0 \\
+0 & 0 & 0 & 0 & 0 & 1 \\
 \end{array}\right)
-\]
-\item $F_k \hat{x}_k + B_k u_k$ describes the station transition form time $k$ to $k+1$.
-The Kalman filter is a general framework which allows not only tracking but also controlling the object.
-I these situations $u_k$ is the control input and $B$ describes the control input model.
-For our application, we are only tracking and not controlling the movement, $B_ku_k = 0$.
-\end{itemize}
-and an update step:
-\begin{eqnarray}
-y_k & = & z_k - H_k\hat{x}_k^-\\
-S_k & = & H_kP_k^-H_k^T + R_k \\
-K_k & = & P_k^-H_k^TS_k^{-1}\\
-\hat{x}_k & = & \hat{x}_k^- + K_ky\\
-P_k & = & (I-K_kH_k)P_k^-\\
-\end{eqnarray}
-where:
-\begin{itemize}
-\item $x$ is the state variable
-\item $P$ is the covariance on the state variable $x$
-\item $z$ is the measurement.
-\item $y$ is the residue in the measurement sapce. 
-\item $H$ is the measurement function that transforms from state space to measurement space.
-\item $R$ is the measurement noise matrix that models sensor noise as a covariance matrix
+\], where $\delta t$ represents the time difference between current time and last state update. 
+\item The Kalman filter is a general framework which allows not only tracking but also controlling the object.
+In these situations $u$ is the control input and $B$ describes the control input model.
+For our application, we are only tracking and not controlling the movement, $Bu = 0$.
 \item $Q$ models the process noise
@@
 
-In our system,
+After the observation for the next coordinates are made, then the Kalman filter calculates the residue between the prediction $\mathbf{x}^-$ and the measurement $z$:
+state $z$:
+\[
+y = z - H\mathbf{x}^-
+\]
+where $H$ is the measurement function that transforms from state space to measurement space. In our example, the transformation is simply taking the location $(x,y)$ from the state space:
+\[
+H = \left(\begin{array}{cccccc}
+1 & 0 & 0 & 0 & 0 & 0 \\
+0 & 1 & 0 & 0 & 0 & 0 \\
+\end{array}\right)
+\]
+
+Then the Kalman filter updates the states estimates $\mathbf{x}$ and states uncertainty estimates $P$:
+\begin{eqnarray}
+\mathbf{x} & = & \mathbf{x}^- + Ky\\
+P & = & (I-KH)P^-
+\end{eqnarray}
+where $K$ is the Kalman gain:
+\begin{eqnarray}
+K & = & P^-H^TS^{-1}\\
+S & = & HP^-H^T + R
+\end{eqnarray}
+$R$ is the measurement noise matrix that models the localization system's output noise as a covariance matrix.

### 2015-06-19 reports/paper/prior_movement.tex
update

@@
 The impulse response is of finite duration for FIR filters.
-In our system, we only have access to location estimates from the past, we look at a special category of FIR filters: causal discrete-time FIR filters.
+In our system, since we only have access to location estimates from the past, we will look at a special category of FIR filters: causal discrete-time FIR filters.
 In such systems, the output $y[n]$ is a linear weighted combination of past $N+1$ estimates from input $x[n]$:
@@
 
-By controlling the impulse response $b_i$, we specify how the past few data should be blended to produce the output.
+By controlling the impulse response $b_i$, we specify how the past few data should be weighted to produce the desired output.
 
@@
 Then it can be shown that the output after filter is an unbiased estimation of the true source location.
-However, if the source has moved during the past $N+1$ estimates, the filter would output the mean location for the past $N+1$ estimated locations, which results in a ``lagging'' effect between the true source location and the system output.
-After the stopped moving, the filter output catches up with the source location.
+However, if the source has moved during the past $N+1$ estimates, the filter would only output the mean location for the past $N+1$ estimated locations, which results in a ``lagging'' effect between the true source location and the system output.
+After the source has stopped moving, the filter output catches up with the source location.
 
-To reduce the ``lagging'' effect ehxibited by the averaging filter, we can assign higher weights to more recent estimates.
+To reduce the ``lagging'' effect exhibited by the averaging filter, we can assign higher weights to more recent estimates.
 Recent estimates gets a higher contribution to the output location, which makes the filtered out tracks more closely with the sound source.
-However, if the localization system is noisy with large estimation variance, then the error in the most recent estimate dominates the filtered output and might produces a large error.
+However, if the localization system is noisy with large estimation variance, then the error in the most recent estimate dominates the filtered output will be prone to noise and exhibit large variance.
 
-As a compromise between balancing the tradeoff between ``lagging'' problem from assigning equal weights to all estimates in the window and the error problem with assigning more weights.
-We will aim to design the system so that it produces estimates at a fast rate, so that the source would not move a large distance and it is safe to average the past a few estimates.
+As a compromise between balancing the tradeoff between ``lagging'' problem from assigning equal weights to all estimates in the window and the noise error problem with assigning more weights to most recent weights.
+We will aim to design a system that can produce location estimates at a fast rate, so that the source would not have enough time to move a significant distance before the next estimate.
+Then it will be safe to average just a few estimates from the past.
 
@@
 Kalman filter uses observed variables to infer hidden variables and use them to help predict the next state.
-In our system, observed variables $z$ are the $(x,y)$ coordinates of the localized acoustic source:
+In our system, observed variable $z$ is the $(x,y)$ coordinates of the localized acoustic source:
 \[
@@
 \]
-Internally, the Kalman filter also keeps track of the uncertainty on the state variables. It is represented as a covariance matrix $P$ on state variables.
+Internally, the Kalman filter also keeps track of the uncertainty of the state variables. It is represented as a covariance matrix $P$ on state variables.
 
 Note that by modeling up to acceleration, we are implicitly assuming higher order motion variables are constant (such as jerk).
-Kalman filter includes a process noise term $Q$ that can be used to partially reduce this effect, but there will exist systematic tracking bias there is acceleration change.
+Kalman filter includes a process noise term $Q$ that can be used to partially reduce this effect, but there will exist systematic tracking bias if there is jerk change.
 
@@
 \begin{itemize}
-\item F is the state transition matrix. In our system, next coordinates $(x,y)$ can be computed with laws of physics using current position, velocity and acceleration:
+\item F is the state transition matrix. In our system, next coordinates $(x,y)$ can be computed with law of physics using current position, velocity and acceleration:
 \[
@@
 \], where $\delta t$ represents the time difference between current time and last state update. 
-\item The Kalman filter is a general framework which allows not only tracking but also controlling the object.
+\item The Kalman filter is a general framework which allows not only object tracking but also controlling the motion of the object.
 In these situations $u$ is the control input and $B$ describes the control input model.

### 2015-06-19 reports/paper/prior.tex
update

@@
 The authors in \cite{ild:1} employed multiple microphones and used the intersection of circles from each microphone pair to estimate the source location.
-While the authors in \cite{ild:3} combined ILD with Interaural Time Difference (ITD) to estimate the source direction (i.e azimuth).
-Instead of solving the intersection of circles, \cite{ild:4} employed machine learning techniques to automatically learn the mapping from ILD and ITD features to location coordinate.
+Authors in \cite{ild:3} combined ILD with Interaural Time Difference (ITD) to estimate the source direction (i.e azimuth).
+Instead of solving the intersection of circles, authors in \cite{ild:4} employed machine learning techniques to automatically learn the mapping from ILD and ITD features to location coordinate.
 Their technique uses four microphones and requires a training phase, during which the sound source is manually placed at predetermined locations for the system to learn the parameters that map from feature space to the sound location.

### 2015-06-19 reports/paper/prior_movement.tex
update

@@
 The impulse response is of finite duration for FIR filters.
-In our system, since we only have access to location estimates from the past, we will look at a special category of FIR filters: causal discrete-time FIR filters.
+In our system, since only location estimates from the past are available, we look at a specific category of FIR filters: causal discrete-time FIR filters.
 In such systems, the output $y[n]$ is a linear weighted combination of past $N+1$ estimates from input $x[n]$:
@@
 To reduce the ``lagging'' effect exhibited by the averaging filter, we can assign higher weights to more recent estimates.
-Recent estimates gets a higher contribution to the output location, which makes the filtered out tracks more closely with the sound source.
-However, if the localization system is noisy with large estimation variance, then the error in the most recent estimate dominates the filtered output will be prone to noise and exhibit large variance.
+Recent estimates gets a higher contribution to the output location, which makes the filtered output tracks more closely with the sound source.
+However, if the localization system is noisy with large estimation variance, then the error in the most recent estimate dominates the filtered output, which makes the filtered output prone to noise and exhibit large variance.
 
-As a compromise between balancing the tradeoff between ``lagging'' problem from assigning equal weights to all estimates in the window and the noise error problem with assigning more weights to most recent weights.
-We will aim to design a system that can produce location estimates at a fast rate, so that the source would not have enough time to move a significant distance before the next estimate.
+To overcome the ''lagging'' problem while making the system robust to noise, we can design a system that produce location estimates at a fast rate.
+This way we can make sure that the source would not have enough time to move a significant distance before the next estimate.
 Then it will be safe to average just a few estimates from the past.
@@
 The Kalman filter is a recursive filter where input data can be efficiently combined to produce online prediction.
-If all noise are Gaussian, the Kalman filter is a statistically optimum filter that minimizes squared error of the estimated parameters.
-Even if the noise is not Gaussian, given only the first two statistics of noise, Kalman filter is the best linear estimator[x].
+If all noise are Gaussian, the Kalman filter is a statistically optimum filter that minimizes squared error of the estimated parameters\cite{kalman:1, kalman:3}.
+Even if the noise is not Gaussian, given only the first two statistics of noise, Kalman filter is the best linear estimator\cite{kalman:2}.
 Due to its statistical optimality and its recursive nature that enables online prediction, the Kalman filter has found applications in a wide range of areas.
-It has been used to track aircraft using RADAR, and to track Robot with sensors and beacons.
+It has been used to track aircraft using RADAR, and to track Robot with sensors and beacons\cite{kalman:4}.
 
@@
 We can also model unobserved motion variables such as velocity $(\dot{x}, \dot{y})$ and acceleration $(\ddot{x}, \ddot{y})$.
-Then the state variables we are tracking $\mathbf{x}$ can be represented as:
+Then the state variables $\mathbf{x}$ that we are tracking can be represented as:
 \[
@@
 \end{array}\right)
-\], where $\delta t$ represents the time difference between current time and last state update. 
+\], where $\delta t$ represents the time difference between current time and the time of last state update. 
 \item The Kalman filter is a general framework which allows not only object tracking but also controlling the motion of the object.
@@
 After the observation for the next coordinates are made, then the Kalman filter calculates the residue between the prediction $\mathbf{x}^-$ and the measurement $z$:
-state $z$:
 \[

### 2015-06-19 reports/paper/simulation.tex
update

@@
   \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_line}
-  \caption{$3$ microphones are placed in a line, $10$ cm apart from each other. The average error is $55.05$ cm}
+  \caption{Error heatmap for different array configurations.
+The heatmap scale is the intersection area measured in $cm^2$.
+$3$ microphones are placed in a line, $10$ cm apart from each other.
+The average error is $55.05$ cm}
   \label{fig:sim_hm_3_line}
@@
   \end{subfigure}
+  \caption{Error heatmap for different array configurations. The heatmap scale is the intersection area measured in $cm^2$}.
 \end{figure*}
@@
   \caption{Error heatmap for different array configurations. The heatmap scale is the intersection area measured in $cm^2$}.
-  \label{fig:sim_hm}
 \end{figure*}
 
-To evaluate how adding one microphone (without increasing the array size) improves accuracy, another microphone is added at $(0,0)$ for the array described in fig~\ref{sim:hm_3}.
+To evaluate how adding one microphone (without increasing the array size) improves accuracy, another microphone is added at $(0,0)$ for the array described in fig~\ref{fig:sim_hm_3}.
 Result is presented in fig~\ref{fig:sim_hm_3_p1}.

### 2015-06-19 reports/paper/system.tex
update

@@
 We used omni-directional foil electret condenser microphone due to its low cost and small size.
-The microphone has a frequency range of $100$ to $10K$ Hz, and a minimum SNR of 58dB[x].
+The microphone has a frequency range of $100$ to $10K$ Hz, and a minimum SNR of 58dB\cite{sys:1}.
 We also used operational amplifer OPA344 from Texas Instrument to amplify the microphone output by a factor of 100, so the received signal can be easily picked up by the analog-to-digital converter (ADC) module installed on the micro-controller.
@@
 To handle the uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation output(equation~\ref{eq:gcc2}) as a measure of the likelihood of different arrival time differences.
-Each index $i$ from the cross-correlation vector denotes the time delay across the two microphones receiving the acoustic signal, and the cross-correlation value $v$[i] at each index $i$ denotes the likelihood of the time delay being $i$.
+Each index $i$ from the cross-correlation vector denotes the time delay across the two microphones receiving the acoustic signal, and the cross-correlation value $k$[i] at each index $i$ denotes the likelihood of the time delay being $i$.
 

### 2015-06-25 reports/paper/experiment.tex
thesis update
added controlling color with volume

@@
 Part of the jiggling can be attributed to the noise from system output, and the rest comes from the hand movement.
+
+\begin{figure}[h!]
+\centering
+  \begin{subfigure}[]{.48\textwidth}
+    \includegraphics[width=\textwidth]{color_1.png}
+    \caption{drawing a strip with maximum volume}
+    \label{fig:show_color_1_1}
+  \end{subfigure}
+  \begin{subfigure}[]{.48\textwidth}
+    \includegraphics[width=\textwidth]{color_2.png}
+    \caption{Drawing another strip with the volume decreased by $20\%$}
+    \label{fig:show_color_1_2}
+  \end{subfigure}
+  \begin{subfigure}[]{.48\textwidth}
+    \includegraphics[width=\textwidth]{color_3.png}
+    \caption{Drawing a third strip with the volume decreased by an additional $20\%$}
+    \label{fig:show_color_1_3}
+  \end{subfigure}
+  \caption{Drawing strips where the color of each strip is controlled by the volume of the white noise}
+  \label{fig:show_color}
+\end{figure}
+
+The color of the painted dots can be controlled by the volume of the sound.
+Fig~\ref{fig:show_color} presented painting three strips with different colors by varying the volume of the white noise.
+The color is generated in the Hue Saturation Value color space where the Saturation and Value are set to $0.9$ while the volume controls the hue.
+Figure~\ref{fig:show_color_1_1} shows the painted first strip where the volume is set to maximum.
+The color is in the red range.
+In fig~\ref{fig:show_color_1_2}, we painted another strip after the volume is decreased by $20$ percent.
+After the volume is decreased, the painted color is in the blue range.
+In fig~\ref{fig:show_color_1_3}, we painted a third strip after the volume is decreased by another $20$ percent.
+In the final strip, the color is in the green range.
+A video demonstration is available at~\cite{demo:color}.
+
+\begin{figure}[h!]
+  \centering
+    \includegraphics[width=\textwidth]{color_4.png}
+    \caption{Painting a heart shape with a pop music by free hand. The color of the heart changes with the volume of the song.}
+    \label{fig:show_color_2}
+\end{figure}
+
+If a music is used, then the volume naturally changes with the content in the music.
+In figure~\ref{fig:show_color_2}, we painted a heart shape by free hand with a pop song.
+The song used here is xx by xxx.
+As can be seen from the figure, the color of the heart changes at each time instant according to the volume of the song.
+A video demonstration is available at~\cite{demo:color2}.

### 2015-06-25 reports/paper/introduction.tex
thesis update
udpate

@@
 
-In this paper, we first discussed in Chapter II some prior relevant research and approaches in sound localization.
+\section{Contribution}
+In this thesis, we provided a detailed simulation of the impact of the arriving time estimation's accuracy on the localization error.
+We evaluated different array architectures, and showed that a two array architecture achieves better localization accuracy compared to a single array of similar physical size.
+We implemented and built the proposed sound localization system and compared differetn state of the art methods for localization and tracking.
+We also
+In this thesis, we first discussed in Chapter II some prior relevant research and approaches in sound localization.
 In Chapter III, we evaluated different array architectures and their impact on localization accuracy.

### 2015-06-26 reports/paper/experiment.tex
udpate
thesis update

@@
 In figure~\ref{fig:show_case}, a few examples of drawing letters with music are presented.
+Each letter took around $5$ to $10$ seconds to draw, depending on the size of the letter and the speed of the movement.
 Green dots on the plots represent the raw localization output and the red line shows the movement output with averaging filtering (window size of $0.5$ seconds is used).
@@
 
-The color of the painted dots can be controlled by the volume of the sound.
-Fig~\ref{fig:show_color} presented painting three strips with different colors by varying the volume of the white noise.
-The color is generated in the Hue Saturation Value color space where the Saturation and Value are set to $0.9$ while the volume controls the hue.
-Figure~\ref{fig:show_color_1_1} shows the painted first strip where the volume is set to maximum.
-The color is in the red range.
+The volume of the acoustic source can also be used to control the color of the dots painted in the above experiment.
+To show how the volume acts as a color control, a simple experiment is shown in Figure~\ref{fig:show_color}.
+In this experiment, three stripes are drawn with the same acoustic source but different volume.
+The color is generated in the Hue-Saturation-Value color space where the Saturation and Value are set to $0.9$ while the hue is controlled by the volume.
+Figure~\ref{fig:show_color_1_1} shows the first stripe where the volume is set to the maximum and the painted color is in the red range.
 In fig~\ref{fig:show_color_1_2}, we painted another strip after the volume is decreased by $20$ percent.
-After the volume is decreased, the painted color is in the blue range.
-In fig~\ref{fig:show_color_1_3}, we painted a third strip after the volume is decreased by another $20$ percent.
-In the final strip, the color is in the green range.
+The painted color changed to the blue range after the volume is decreased.
+In fig~\ref{fig:show_color_1_3}, we painted a third strip where we further decreased the volume by another $20$ percent.
+In this final stripe, the color changed to the green range.
 A video demonstration is available at~\cite{demo:color}.
@@
 
-If a music is used, then the volume naturally changes with the content in the music.
-In figure~\ref{fig:show_color_2}, we painted a heart shape by free hand with a pop song.
-The song used here is xx by xxx.
+Instead of manually varying the volume of the acoustic source, a natural extension is to use music as our acoustic source.
+Generally, the volume of music will vary with time and this change will be reflected by the colors drawn.
+In figure~\ref{fig:show_color_2}, we used a pop song as our acoustic source and painted a heart shape by free hand.
+The song used here is \emph{Don't Let It Break Your Heart} by \emph{Coldplay}.
 As can be seen from the figure, the color of the heart changes at each time instant according to the volume of the song.

### 2015-06-26 reports/paper/introduction.tex
thesis update

--- 2015-06-19 reports/paper/introduction.tex
@@
 
-In this paper, we first discussed in Chapter II some prior relevant research and approaches in sound localization.
+\section{Contribution}
+In this thesis, we provided a detailed simulation analyzing the impact of the arriving time estimation accuracy on localization error.
+We evaluated different array architectures, and showed that a two array architecture achieves good localization accuracy while maintaining good portability.
+We implemented and built the proposed sound localization system and compared different state of the art methods for localization and tracking.
+We also demonstrated that our system can be used as a tool for painting in space.
+In this thesis, we first discussed in Chapter II some prior relevant research and approaches in sound localization.
 In Chapter III, we evaluated different array architectures and their impact on localization accuracy.

### 2015-06-26 reports/paper/experiment.tex
added volume based scaling factor

@@
   \begin{subfigure}[]{.48\textwidth}
+    \includegraphics[width=\textwidth]{scaling_pos.png}
+    \caption{randomly going through the region to collect localized $(x,y)$ data and the received volume data. The volume is kept constant}
+    \label{fig:color_pos}
+  \end{subfigure}
+  \centering
+  \begin{subfigure}[]{.48\textwidth}
+    \includegraphics[width=\textwidth]{color_scaling.png}
+    \caption{calculate a scaling factor for each point in the region. The scaling factor is the average of volume of the 5 closest points from the plot on the left}
+    \label{fig:color_scale}
+  \end{subfigure}
+  \caption{Sound volume normalization}
+  \label{fig:color_norm}
+\end{figure}
+
+The volume of the acoustic source can also be used to control the color of the dots painted in the above experiment.
+However, the volume of the audio signal received is itself a function of the location of the audio source.
+Therefore, the received volume needs to be normalized.
+For the volume $V$ received at location $(x,y)$, we normalize it by a scaling factor $b_{xy}$.
+The normalized volume is $\bar{V}$:
+\[
+\bar{V} = \frac{V}{b_{xy}}
+\]
+The scaling factor $b_{xy}$ is location dependent.
+It is found by from calibration, which only needs to be carried out once.
+During calibration, a white noise sound source is moved randomly in the region with constant volume.
+As the source moves around in the region the calibration modules collects the source's location and received volume information.
+Then $b_{xy}$ is calculated by taking the volume average of $10$ closest samples from the collected data.
+Fig~\ref{fig:color_norm} shows and example of the volume control calibration procedure to calibrate volume information in a $60$ cm by $40$ cm region.
+Fig~\ref{fig:color_pos} shows how the white noise source with constant volume is moved in the region and fig~\ref{fig:color_scale} shows the scaling factor $b_{xy}$ for each point in the region by taking the volume average of $5$ closest points in fig~\ref{fig:color_pos}.
+
+\begin{figure}[h!]
+\centering
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{color_1.png}
@@
 
-The volume of the acoustic source can also be used to control the color of the dots painted in the above experiment.
-To show how the volume acts as a color control, a simple experiment is shown in Figure~\ref{fig:show_color}.
+To show how the volume acts as a color control.
+a simple experiment is shown in Figure~\ref{fig:show_color}.
 In this experiment, three stripes are drawn with the same acoustic source but different volume.

### 2015-06-26 reports/paper/experiment.tex
thesis update

--- 2015-06-26 reports/paper/experiment.tex
@@
   \begin{subfigure}[]{.48\textwidth}
+    \includegraphics[width=\textwidth]{scaling_pos.png}
+    \caption{randomly going through the region to collect localized $(x,y)$ data and the received power data. The volume of the source is kept constant}
+    \label{fig:color_pos}
+  \end{subfigure}
+  \centering
+  \begin{subfigure}[]{.48\textwidth}
+    \includegraphics[width=\textwidth]{color_scaling.png}
+    \caption{Calculate a scaling factor for each point in the region. The scaling factor is the average power of the $5$ closest points from the points on the left}
+    \label{fig:color_scale}
+  \end{subfigure}
+  \caption{Volume control normalization}
+  \label{fig:color_norm}
+\end{figure}
+
+The volume of the acoustic source can also be used to control the color of the dots painted in the above experiment. For received a signal $x[n]$, the power $P$ is an indicator of the source volume:
+\[
+P = \frac{1}{N} \sum_{i=1}^N x[n]^2
+\]
+However, the received power of the audio signal is itself a function of the location of the audio source (even as the source volume stays constant).
+Therefore, the received power needs to be normalized.
+For the power $P$ received at location $(x,y)$, we normalize it by a scaling factor $b_{xy}$.
+The normalized power $\bar{P}$:
+\[
+\bar{P} = \frac{P}{b_{xy}}
+\]
+The scaling factor $b_{xy}$ is location dependent.
+It is found from calibration, which needs to be carried out once.
+During the calibration, a white noise sound source is moved randomly in the region with constant volume.
+As the source moves around in the region the calibration modules collects the source's location and received power information.
+Then $b_{xy}$ is calculated by taking the average power of $10$ closest samples from the collected data.
+Fig~\ref{fig:color_norm} shows and example of the volume control calibration procedure to calibrate the power information in a $60$ cm by $40$ cm region.
+Fig~\ref{fig:color_pos} shows how the white noise source with constant volume is moved in the region and fig~\ref{fig:color_scale} shows the scaling factor $b_{xy}$ for each point in the region by taking the average power of $5$ closest points in fig~\ref{fig:color_pos}.
+
+\begin{figure}[h!]
+\centering
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{color_1.png}

### 2015-06-26 reports/paper/experiment.tex
thesis update

--- 2015-06-26 reports/paper/experiment.tex
@@
   \begin{subfigure}[]{.48\textwidth}
+    \includegraphics[width=\textwidth]{scaling_pos.png}
+    \caption{randomly traverse the region to collect location $(x,y)$ and received power data. The volume of the source is kept constant}
+    \label{fig:color_pos}
+  \end{subfigure}
+  \centering
+  \begin{subfigure}[]{.48\textwidth}
+    \includegraphics[width=\textwidth]{color_scaling.png}
+    \caption{calculate a scaling factor $b_{xy}$ for each point in the region. The scaling factor is the average power of $5$ closest points from the traversed points on the left}
+    \label{fig:color_scale}
+  \end{subfigure}
+  \caption{Volume control normalization. Two microphone arrays are placed at $(-50\mbox{ cm}, 0\mbox{ cm})$ and $(50\mbox{ cm}, 0\mbox{ cm})$.}
+  \label{fig:color_norm}
+\end{figure}
+
+The volume of the acoustic source can also be used to control the color of the dots painted in the above experiment.
+For signal $x_i[n]$ received at each microphone $i$, the power $P_i$ can be used as an indicator of the source volume:
+\[
+P_i = \frac{1}{N} \sum_{i=1}^N x_i[i]^2
+\]
+Since there are $6$ microphones in our system ($3$ in each array), we choose the maximum power $P$ received across all $6$ microphones as the indicator for the source volume:
+\[
+P = \max_i P_i 
+\]
+However, the received power of the audio signal is itself a function of the audio source location (since the received power decays with increasing distance between the source to receiver).
+Therefore, the received power needs to be normalized.
+For the power $P$ received at location $(x,y)$, we normalize it by a scaling factor $b_{xy}$.
+The normalized power $\bar{P}$ is:
+
+\[
+\bar{P} = \frac{P}{b_{xy}}
+\]
+
+The scaling factor $b_{xy}$ is location dependent.
+To calculate $b_{xy}$, we conduct an one time calibration step at the beginning.
+During the calibration step, a white noise sound source is used to randomly traverse the entire region with constant volume.
+As the source moves around in the region the calibration modules collects the source's location and the received power information.
+Then $b_{xy}$ can be calculated for each location $(x,y)$  by taking the average power of $5$ closest samples from that location.
+Fig~\ref{fig:color_norm} shows an example of the volume control calibration procedure for calibrating the location power information in a $60$ cm by $40$ cm region.
+Fig~\ref{fig:color_pos} shows how the white noise source with constant volume is traversed in the region and fig~\ref{fig:color_scale} shows the corresponding scaling factor $b_{xy}$ for each point in the region by taking the average power of $5$ closest points in fig~\ref{fig:color_pos}.
+As can be seen from the figure, for any fixed $y$ coordinate, the received power first decreases then increases along increase in the $x$ direction.
+This is because the arrays are placed at the top left and top right corners of the region.
+For a fixed $x$ coordinate, the received power decreases monotonically along the $y$ direction.
+
+\begin{figure}[h!]
+\centering
+  \begin{subfigure}[]{.48\textwidth}
     \includegraphics[width=\textwidth]{color_1.png}
-    \caption{drawing a strip with maximum volume}
+    \caption{drawing a stripe with maximum volume}
     \label{fig:show_color_1_1}
@@
     \includegraphics[width=\textwidth]{color_2.png}
-    \caption{Drawing another strip with the volume decreased by $20\%$}
+    \caption{Drawing another stripe with the volume decreased by $20\%$}
     \label{fig:show_color_1_2}
@@
     \includegraphics[width=\textwidth]{color_3.png}
-    \caption{Drawing a third strip with the volume decreased by an additional $20\%$}
+    \caption{Drawing a third stripe with the volume decreased by an additional $20\%$}
     \label{fig:show_color_1_3}
   \end{subfigure}
-  \caption{Drawing strips where the color of each strip is controlled by the volume of the white noise}
+  \caption{Drawing stripes where the color of each stripe is controlled by the volume of the white noise}
   \label{fig:show_color}
@@
 
-The volume of the acoustic source can also be used to control the color of the dots painted in the above experiment.
-To show how the volume acts as a color control, a simple experiment is shown in Figure~\ref{fig:show_color}.
+To show how the volume acts as a color control.
+A simple experiment is shown in Figure~\ref{fig:show_color}.
 In this experiment, three stripes are drawn with the same acoustic source but different volume.

### 2015-06-26 reports/paper/system.tex
add

@@
+\subsection{Hardware}
 The end system has two arrays, each with three microphones mounted on the vertices of a $20$ cm equilateral triangle.
@@
 \end{figure}
+
+\subsection{Software}
+
+On the software side, \emph{Python} is used as the main programming language since it has extensive libraries in both real time data handling and signal processing.
+\emph{zeroMQ} is an inter process messaging queue that is used in our system to pass data across different modules.
+Our system is made up of two data acquisition modules (each is used to receive raw date from microphone array output) and one localization module (listens to both data acquisition modules and perform localization using the raw microphone data).
+Using \emph{zeroMQ} as connections to different modules adds flexibility to our system, as we can design applications such as the drawing application we have used in our system to interface only with the localization module and disregard how the microphone data is collected.
+Other localization applications can be easily integrated into our system by connecting them with the localization module.
+Furthermore, we have built a recording module that interfaces with the data acquisition modules for offline analysis and parameter tuning.
 
@@
 
-To see the effect of accuracy improvement using multiple arrays, fig~\ref{fig:liklihood} shows a real life localization where the source is placed at $(0$ cm$,-30$ cm$)$.
-The top two figures show the individual likelihood maps produced by single microphone arrays.
-We can see that the individual arrays can give accurate angle estimate, but have high uncertainty in distance estimate.
-The bottom figure shows the combined likelihood map according to equation~\ref{eq:combine_l}.
-The combined likelihood map demonstrated that by merging estimates from two arrays the system is able to perform more accurate localization.
-
 \begin{figure}[h!]
@@
 
+To see the effect of accuracy improvement using multiple arrays, fig~\ref{fig:liklihood} shows a real life localization where the source is placed at $(0$ cm$,-30$ cm$)$.
+The top two figures show the individual likelihood maps produced by single microphone arrays.
+We can see that the individual arrays can give accurate angle estimate, but have high uncertainty in distance estimate.
+The bottom figure shows the combined likelihood map according to equation~\ref{eq:combine_l}.
+The combined likelihood map demonstrated that by merging estimates from two arrays the system is able to perform more accurate localization.
+
 From a timing point of view, the micro-controller spends $12$ milliseconds on sampling the microphone data before sending it to a computer for processing.

### 2015-06-26 reports/paper/introduction.tex
add
update

@@
 In this work, we have built a source localization system that is portable, inexpensive, yet reasonably accurate for localization in a small area.
+The purpose of our system is to give applications to designs that allow remote human computer interaction.
 To test the accuracy of our localization system, we perform acoustic localization experiments in an one meter by one meter area.
@@
 We also demonstrated that our system can be used as a tool for painting in space.
-In this thesis, we first discussed in Chapter II some prior relevant research and approaches in sound localization.
+
+We first discussed in Chapter II some prior relevant research and approaches in sound localization.
 In Chapter III, we evaluated different array architectures and their impact on localization accuracy.

### 2015-06-26 reports/paper/prior_movement.tex
update

@@
 Note that by modeling up to acceleration, we are implicitly assuming higher order motion variables are constant (such as jerk).
-Kalman filter includes a process noise term $Q$ that can be used to partially reduce this effect, but there will exist systematic tracking bias if there is jerk change.
+This assumption gives the system a tracking bias if there is a jerk change.
+Kalman filter also includes a term $Q$ that can be used to model the process noise.
 

### 2015-06-26 reports/paper/conclusion.tex
thesis update

@@
 Since the likelihood calculation on each grid point is simply a look up on the cross-correlation output from each microphone pair, the computation at each grid point is independent of each other.
-Therefore, One straightforward way to speedup calculation is to employ hardware accelerated parallel computing technologies such as GPU or FPGA.
+Therefore, one straightforward way to speedup calculation is to employ hardware accelerated parallel computing technologies such as GPU or FPGA.
 
 Another direction is to investigate multi-source localization.
-One approach is to allocate different frequency bands to difference sound sources.
-However, the number of frequency bands are limited in the audible audio spectrum, and our results showed that TDOA estimation is more accurate for a wide band signal compared to a narrow band signal.
-A different approach is to digitally encode and tag the audio source.
-Different sources can be differentiated if each source uses a waveform that has a different digital signature.
+One approach is to allocate different frequency bands to different sound sources.
+However, the number of frequency bands are limited in the audible spectrum, and our results showed that TDOA estimation is more accurate for a wide band signal compared to a narrow band signal.
+To allow signals to occupy the entire frequency band, one potential approach is to devise protocols that allow time division multiplex of source signals, where each source signal occupies a different time slot.
+Another approach is for each source to have a different audio signature.
+If the localization module knows all sources' audio signatures, it can use that information to recover each source signal from the received superimposed waveform.

### 2015-06-26 reports/paper/system.tex
thesis update

@@
 On the software side, \emph{Python} is used as the main programming language since it has extensive libraries in both real time data handling and signal processing.
-\emph{zeroMQ} is an inter process messaging queue that is used in our system to pass data across different modules.
+\emph{ZeroMQ} is an inter process messaging queue that is used in our system to pass data across different modules.
 Our system is made up of two data acquisition modules (each is used to receive raw date from microphone array output) and one localization module (listens to both data acquisition modules and perform localization using the raw microphone data).
-Using \emph{zeroMQ} as connections to different modules adds flexibility to our system, as we can design applications such as the drawing application we have used in our system to interface only with the localization module and disregard how the microphone data is collected.
+Using ZeroMQ as connections to different modules adds flexibility to our system, as we can design applications such as the drawing application we have used in our system to interface only with the localization module and disregard how the microphone data is collected.
 Other localization applications can be easily integrated into our system by connecting them with the localization module.

### 2015-06-26 reports/paper/experiment.tex
update
thesis update

@@
 \includegraphics[width=1.0\textwidth]{error_distribution.eps}
-\caption{error distribution in the grid. Arrays are placed at $(-0.5$ m$, 0$ m$)$ and $(0.5$ m$, 0$ m$)$}
+\caption{Error distribution in the grid. Arrays are placed at $(-0.5$ m$, 0$ m$)$ and $(0.5$ m$, 0$ m$)$}
 \label{fig:error_distribution}
@@
 The color is generated in the Hue-Saturation-Value color space where the Saturation and Value are set to $0.9$ while the hue is controlled by the volume.
-Figure~\ref{fig:show_color_1_1} shows the first stripe where the volume is set to the maximum and the painted color is in the red range.
+Figure~\ref{fig:show_color_1_1} shows the first stripe where the volume is set to the maximum and the painted color is in the red to orange range.
 In fig~\ref{fig:show_color_1_2}, we painted another strip after the volume is decreased by $20$ percent.
-The painted color changed to the blue range after the volume is decreased.
+The painted color changed to the blue to purple range after the volume is decreased.
 In fig~\ref{fig:show_color_1_3}, we painted a third strip where we further decreased the volume by another $20$ percent.
-In this final stripe, the color changed to the green range.
+In this final stripe, the color changed to the green to yellow range.
 A video demonstration is available at~\cite{demo:color}.

### 2015-06-27 reports/paper/introduction.tex
thesis update

@@
 \section{Contribution}
-In this thesis, we provided a detailed simulation analyzing the impact of the arriving time estimation accuracy on localization error.
-We evaluated different array architectures, and showed that a two array architecture achieves good localization accuracy while maintaining good portability.
-We implemented and built the proposed sound localization system and compared different state of the art methods for localization and tracking.
-We also demonstrated that our system can be used as a tool for painting in space.
+The main contributions of the thesis are:
+\begin{enumerate}
+\item We provided a detailed simulation analyzing the impact of the arriving time estimation accuracy on localization error. 
+\item We evaluated different array architectures, and showed that a two array architecture achieves good localization accuracy while maintaining good portability. 
+\item We implemented and built the proposed sound localization system and compared different state of the art methods for localization and tracking. 
+\end{enumerate}
+As a fun demonstration of the technology, we also demonstrated that our system can be used as a tool for painting in space.
 
+\section{Outline}
 We first discussed in Chapter II some prior relevant research and approaches in sound localization.

### 2015-06-27 reports/paper/introduction.tex
thesis update

@@
 \section{Outline}
-We first discussed in Chapter II some prior relevant research and approaches in sound localization.
+We first discussed in Chapter II prior relevant research and approaches in sound localization.
+We also discussed various ways of combining past localization estimates for movement tracking.
 In Chapter III, we evaluated different array architectures and their impact on localization accuracy.

### 2015-06-27 reports/paper/experiment.tex
thesis update

@@
 \]
+
 Since there are $6$ microphones in our system ($3$ in each array), we choose the maximum power $P$ received across all $6$ microphones as the indicator for the source volume:
@@
 \]
+
 However, the received power of the audio signal is itself a function of the audio source location (since the received power decays with increasing distance between the source to receiver).
@@
   \centering
+    \includegraphics[width=\textwidth]{color_5.png}
+    \caption{Painting an eclipse. The color of the heart changes with the volume of the song.}
+    \label{fig:show_color_3}
+\end{figure}
+
+\textcolor{red}{ Instead of stop and change the sound volume, we carried out a similar experiment where the volume is changed as we move the source.
+In this experiment, we drew an ellipse by free hand in a clockwise fashion.
+We started with the volume set to the maximum (which corresponds to the red color).
+Then we gradually decreases the volume as the move to the right.
+After the volume is decreased to $60\%$ of the maximum volume we started with, we start to gradually increase the volume until the maximum volume is reached again.
+Figure~\ref{fig:show_color_3} gives a visual presentation of the painted result.
+As can be seen from the figure, the color is red at the top left region of the ellipse.
+The color changes to the green yellow range and then back go red orange range as we completes the ellipse.
+A video demonstration of this experiment is available at~\cite{demo:color3} }
+
+\begin{figure}[h!]
+  \centering
     \includegraphics[width=\textwidth]{color_4.png}

### 2015-06-27 reports/paper/simulation.tex
update
thesis update

@@
 The average error is $2.60$ cm.
+\textcolor{red}{Generally, the clock used to sample microphone data has to be the same in order to compare arrival time differences.
+In the next Chapter, we will describe a method that does not require the clock to be synchronized across arrays.
+This approach would make the system easier to design since we don't need to consider clock synchronization issue}
 
@@
 Looking at the simulation results on localization accuracy for different microphone array size and configurations, we see that the further away the microphones are placed from each other the higher the localization accuracy.
-However, as we increase the distance between each microphone pairs, we size of the overall array increases and the system becomes less portable.
+However, as we increase the distance between each microphone pairs, the size of the overall array increases and the system becomes less portable.
 In the end we decided to build the two array system as described in fig~\ref{fig:sim_hm_2_array}.

### 2015-06-28 reports/paper/experiment.tex
thesis update

@@
 \chapter{Experiment}
+In this chapter, we first discuss the hardware equipments that are used to build the final system.
+Next we discuss the high level software architecture we designed and implemented in our system.
+We then detail the localization method we used in our system and how we circumvented the need to synchronize clocks across the two arrays.
+We then describe the experiments we have conducted using our system to test its localization accuracy and system responsiveness.
+These experiments include acoustic source location estimates and movement tracking.
+In the discussion section of this chapter we analyze the results from the experiments above.
+To conclude this chapter, we have included several interesting applications of our system demonstrating how it can be used to facilitate human computer interaction.
+
 \section{System}
@@
 
-\textcolor{red}{ Instead of stop and change the sound volume, we carried out a similar experiment where the volume is changed as we move the source.
-In this experiment, we drew an ellipse by free hand in a clockwise fashion.
-We started with the volume set to the maximum (which corresponds to the red color).
-Then we gradually decreases the volume as the move to the right.
-After the volume is decreased to $60\%$ of the maximum volume we started with, we start to gradually increase the volume until the maximum volume is reached again.
+\textcolor{red}{To test how sensitive the color change is to the different volumes in our system, we have conducted the following experiment.
+In this experiment, we moved the acoustic source clockwise in an elliptical fashion.
+As we are moving the source, we varied the source volume gradually and noted its color change.
+We started with the volume set to the maximum (the corresponding color is red).
+Then we gradually decreased the volume as the move to the right.
+After the volume is decreased to $60\%$ of the maximum, we started to gradually increase the volume until the maximum is reached again.
 Figure~\ref{fig:show_color_3} gives a visual presentation of the painted result.
-As can be seen from the figure, the color is red at the top left region of the ellipse.
-The color changes to the green yellow range and then back go red orange range as we completes the ellipse.
+As can be seen from the figure, the top left region of the drawn ellipse is red.
+The color then gradually changes to the green yellow range and then back to the red orange range as we complete the ellipse.
 A video demonstration of this experiment is available at~\cite{demo:color3} }

### 2015-06-28 reports/paper/simulation.tex
thesis update

--- 2015-06-19 reports/paper/simulation.tex
@@
 The result indicates that this configuration has good accuracy when source is close to the arrays.
-We observe that the localization accuracy decreases as the sound source moves outside of the region one meter by one meter away from the two arrays.
+We observe that the localization accuracy decreases as the sound source moves outside of the one meter by one meter region away from the two arrays.
 The average error is $2.60$ cm.
+In general, synchronized clocks must be used to sample microphone data to ensure accurate comparison of arrival time differences.
+In practice, this means clocks must be synchronized for all the microphones across the two arrays.
+In the next chapter, we will describe a method that does not require clocks to be synchronized across the two arrays.
+This approach would make the system easier to design, although clocks still needs to be synchronized for microphones within the same array.
 

### 2015-06-28 reports/paper/experiment.tex
thesis update

@@
 The error is below $3$ cm for most areas inside the grid.
-There is one error spike in the mid-left region and we contribute this to audio source misplacement because the error is fairly low and consistent around that spike.
+However, note that there is one error spike in the mid-left region in the grid.
+We attribute this inconsistency with the surrounding region to errors we have made in placing the audio source, since manually placing the audio source with centimeter precision is difficult.
+To confirm this speculation, we later reevaluated the errors in this region and found them to be close to the errors in the surrounding region.
 
@@
 The result is presented in fig~\ref{fig:error_along_y}.
-The localization error is within $3$ cm when the source is within $100$ cm from the arrays.
+The localization error is on average below $3$ cm when the source is within $100$ cm along the $y$ axis from the arrays.
 The error increases to about $5$ cm when the source distance increases to $150$ cm and the error exceeds $10$ cm after the source distance reaches $200$ cm.
+We can see that our system achieves good accuracy within the one meter by one meter region and gradually loses its localization accuracy as we move the source outside.
 
-\clearpage
+%\clearpage
 \subsection{Movement tracking}
@@
 By comparing experiments at normal speed (fig~\ref{fig:circle_wnn} to \ref{fig:circle_musicbn}) with experiments at fast speed (fig~\ref{fig:circle_wnf} to~\ref{fig:circle_musicbf}), we find that the localization error does not increase when the movement speed increased from $10$ cm per second to $20$ cm per second.
+However, it's conceivable that the localization error will increase with further speed increase, but since our system is designed for HCI, we feel $20$ cm per second covers most of the use cases.
 

### 2015-06-28 reports/paper/system.tex
thesis update

@@
 Other localization applications can be easily integrated into our system by connecting them with the localization module.
-Furthermore, we have built a recording module that interfaces with the data acquisition modules for offline analysis and parameter tuning.
+Furthermore, we have built a recording module that interfaces with the data acquisition modules for offline data analysis and parameter tuning.
 
-To handle the uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation output(equation~\ref{eq:gcc2}) as a measure of the likelihood of different arrival time differences.
+\subsection{Localization Method}
+In this section, we will describe the method we used for localization in our system and the advantages of using such an approach.
+In chapter two we described how to use the TDOA approach for point source location estimation using three microphones.
+This approach entails first finding the three arrival delays corresponding to equation ~\ref{eq:gcc} for each microphone pair.
+These three arrival delays can then be mapped to three distinct hyperbolas inside the search region.
+The intersection of these three hyperbolas can then be used as the estimate for the source location.
+However, this approach bears two uncertainties.
+First, there is no guarantee that the three hyperbolas will intersect with each other, and we cannot assign an estimate in such a case.
+Second, in our two array system, if each array gives a different point estimate, we have no way of knowing which array's estimate is more accurate or how to combine the two.
+
+To handle these two uncertainties, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation output(equation~\ref{eq:gcc2}) as a measure of the likelihood of different arrival time differences.
 Each index $i$ from the cross-correlation vector denotes the time delay across the two microphones receiving the acoustic signal, and the cross-correlation value $k$[i] at each index $i$ denotes the likelihood of the time delay being $i$.
@@
 Sending the data through the USB port takes another $15$ milliseconds, and processing on the computer takes around $50$ milliseconds.
-Therefore, the total time lag between sound source and localization is around $80$ milliseconds.
+Therefore, the total time lag between sound source and localization is around $80$ milliseconds, which translates to around $12.5$ localizations per second.

### 2015-06-28 reports/paper/experiment.tex
update

@@
 
-\textcolor{red}{To test how sensitive the color change is to the different volumes in our system, we have conducted the following experiment.
+To test how sensitive the color change is to the different volumes in our system, we have conducted the following experiment.
 In this experiment, we moved the acoustic source clockwise in an elliptical fashion.
@@
 The color then gradually changes to the green yellow range and then back to the red orange range as we complete the ellipse.
-A video demonstration of this experiment is available at~\cite{demo:color3} }
+A video demonstration of this experiment is available at~\cite{demo:color3}.
 

### 2015-06-28 reports/paper/experiment.tex
thesis update

@@
 By comparing experiments at normal speed (fig~\ref{fig:circle_wnn} to \ref{fig:circle_musicbn}) with experiments at fast speed (fig~\ref{fig:circle_wnf} to~\ref{fig:circle_musicbf}), we find that the localization error does not increase when the movement speed increased from $10$ cm per second to $20$ cm per second.
-However, it's conceivable that the localization error will increase with further speed increase, but since our system is designed for HCI, we feel $20$ cm per second covers most of the use cases.
+However, it's conceivable that the localization error will increase with further speed increase, but since our system is designed for HCI, a speed of $20$ cm per second covers most of the use cases.
 
@@
     \includegraphics[width=\textwidth]{color_5.png}
-    \caption{Painting an eclipse. The color of the heart changes with the volume of the song.}
+    \caption{Painting an ellipse. The color of the painting is controlled by the volume of the white noise.}
     \label{fig:show_color_3}

### 2015-06-28 reports/paper/introduction.tex
thesis update

@@
 Global Positioning System (GPS) is the prevailing technology used for outdoor localization.
-A set of 32 satellites equipped with synchronized atomic clock orbit around the earth and broadcast their time and position information at fixed intervals.
+A set of $32$ satellites equipped with synchronized atomic clock orbit around the earth and broadcast their time and position information at fixed intervals~\cite{intro:gps2}.
 A GPS receiver listens to at least 4 satellites and uses the broadcasted timing information to infer the location.
@@
 
-Ultrasound based indoor localization approaches, on the other hand, have achieved sub-centimeter accuracy~\cite{intro:ultra}.
+Ultrasound based indoor localization approaches, on the other hand, have achieved sub-centimeter accuracy~\cite{intro:ultra, intro:ultra2}.
 These systems employ multiple ultrasound receivers and use the arrival time difference between the sound source and receivers to infer the source location.
@@
 In these systems, device signal strength received at different base stations are used for the estimation of the user device location.
-However, their reported accuracy are in the range of $1$ to $5$ meters~\cite{intro:blue, intro:loc}, which is not accurate enough for local movement tracking.
+However, their reported accuracy are in the range of $1$ to $5$ meters~\cite{intro:blue, intro:loc, intro:wifi}, which is not accurate enough for local movement tracking.
 

### 2015-06-28 reports/paper/prior.tex
thesis update

@@
 With multiple microphones, it is possible to infer the source location by comparing the signal intensity received at different microphones.
-Human's auditory system has used ILD cues to infer source direction~\cite{ild:2}, and this technique is most effective when used to localize high frequency sources, because these sources don't diffract significantly around the listener's head and produce a significant intensity difference.
+Human's auditory system has used ILD cues to infer source direction~\cite{ild:2, ild:h, ild:5}, and this technique is most effective when used to localize high frequency sources, because these sources don't diffract significantly around the listener's head and produce a significant intensity difference.
 
@@
 This is somewhere between GCC and GCC\_PHAT.
-Part of magnitude information is included in arrival time difference estimation.
+Part of the magnitude information is included in arrival time difference estimation.
 \end{description}

### 2015-06-28 reports/paper/system.tex
thesis update

@@
 Fig~\ref{fig:setup_array} shows a picture of the array setup.
-The micro-controller contains an onboard RAM of $64$k, and an ADC module capable of sampling at $500$kHz~\cite{tdoa:micloc, sys:teensy}.
+The micro-controller contains an onboard RAM of $64$k, and an ADC module capable of sampling at $500$kHz~\cite{tdoa:micloc, sys:teensy, sys:teensy2}.
 In this project, the micro-controller collects microphone data on all three channels for a duration $12$ milliseconds and then sends the recorded data to a computer through the USB port for localization.
@@
 \emph{ZeroMQ} is an inter process messaging queue that is used in our system to pass data across different modules.
-Our system is made up of two data acquisition modules (each is used to receive raw date from microphone array output) and one localization module (listens to both data acquisition modules and perform localization using the raw microphone data).
+Our system is made up of two data acquisition modules (each is used to receive raw data from microphone array output) and one localization module (listens to both data acquisition modules and perform localization using the raw microphone data).
 Using ZeroMQ as connections to different modules adds flexibility to our system, as we can design applications such as the drawing application we have used in our system to interface only with the localization module and disregard how the microphone data is collected.

### 2015-07-02 reports/paper/conclusion.tex
thesis update

@@
 In this thesis, we described and built an inexpensive, portable yet reasonably accurate sound localization system with two microphone arrays.
-We have analyzed different array architectures and showed that the localization accuracy varies with the array geometry and the location of the sound source (in terms of both the distance and the angle).
-In particular, regions close to the line connecting two microphones are generally more difficult to localize than regions close the line bisecting two microphones.
+We have analyzed different array architectures and showed that the localization accuracy varies with the array geometry and the location of the sound source (both in terms of the distance and the angle).
+In particular, regions close to the line connecting two microphones are generally more difficult to localize than regions close to the line bisecting two microphones.
 This problem can be alleviated by employing multiple microphones that are placed at large distance from each other, but such arrangement affects the portability of the system.
 As a trade off between accuracy and portability, we demonstrated that a two array architecture achieves better accuracy compared to a single array system of similar size.
-By using cross-correlation output as the likelihood for different arrival time differences, each array generates a likelihood map for each possible location in the area.
+By using cross-correlation output as the likelihood estimate for different arrival time differences, each array generates a likelihood map for each possible location in the area.
 We demonstrated that by merging likelihood maps from the two arrays, we were able to achieve good localization accuracy (less than $3$ cm average error) in a local one meter by one meter region.

### 2015-07-02 reports/paper/introduction.tex
thesis update

@@
 To test the accuracy of our localization system, we perform acoustic localization experiments in an one meter by one meter area.
-The experiment area size is adequate for our purposes because human arm movement will generally be constrained in this area.
+The experiment area is adequate for our purposes because human arm movement will generally be constrained in this area.
 Our experiment show that our system localizes both static and moving acoustic source with good accuracy in this area.
@@
 
-This system can also be used to in Augmented Reality (AR) applications.
+This system can also be used in Augmented Reality (AR) applications.
 For example, the user can wear a music tag on one finger, and then the system would be able to track the user's finger movements.
@@
 \section{Contribution}
-The main contributions of the thesis are:
+The main contributions of this thesis are:
 \begin{enumerate}
-\item We provided a detailed simulation analyzing the impact of the arriving time estimation accuracy on localization error. 
-\item We evaluated different array architectures, and showed that a two array architecture achieves good localization accuracy while maintaining good portability. 
-\item We implemented and built the proposed sound localization system and compared different state of the art methods for localization and tracking. 
+\item We provided a detailed simulation analyzing the impact of arriving time difference estimation on localization error. 
+\item We evaluated different array architectures, and showed that a two array architecture achieves good localization accuracy while maintaining portability. 
+\item We implemented and built the proposed sound localization system and compared different state of the art methods for point localization and movement tracking. 
 \end{enumerate}
-As a fun demonstration of the technology, we also demonstrated that our system can be used as a tool for painting in space.
+As a fun showcase of our technology, we also demonstrated that our system can be used as a tool for painting in space.
 
 \section{Outline}
-We first discussed in Chapter II prior relevant research and approaches in sound localization.
-We also discussed various ways of combining past localization estimates for movement tracking.
+We first discussed in Chapter II the relevant research and approaches developed for sound localization.
+We also discussed various ways of combining recent localization estimates for movement tracking.
 In Chapter III, we evaluated different array architectures and their impact on localization accuracy.
 We demonstrated that a two array system outperforms a single array system of similar physical dimension.
-In Chapter IV, we first presented the chosen architecture along with hardware details, followed by experiment details and results.
+In Chapter IV, we first presented the chosen architecture along with its hardware details.
+Then we explained the experiment details and discussed the corresponding results.

### 2015-07-10 reports/paper/introduction.tex
thesis update

@@
 A set of $32$ satellites equipped with synchronized atomic clock orbit around the earth and broadcast their time and position information at fixed intervals~\cite{intro:gps2}.
-A GPS receiver listens to at least 4 satellites and uses the broadcasted timing information to infer the location.
+A GPS receiver listens to at least 4 satellites and uses the broadcasted timing information to infer the source location.
 Commercial grade GPS has an average error of a few meters, depending on the size and quality of the receiver~\cite{intro:gps}.
@@
 These systems employ multiple ultrasound receivers and use the arrival time difference between the sound source and receivers to infer the source location.
-Although obtaining good accuracy, ultrasound systems require the use of expensive transducers.
+Although these ultrasound based systems can obtain good indoor localization accuracy, they require the use of expensive transducers.
 
@@
 The purpose of our system is to give applications to designs that allow remote human computer interaction.
+We will demonstrate in the experiment section that our system can be used as a tool to paint in space.
 To test the accuracy of our localization system, we perform acoustic localization experiments in an one meter by one meter area.

### 2015-07-10 reports/paper/prior.tex
thesis update

@@
 
-For a point sound source in a direct field, the signal intensity decays proportional to the square of the distance between the source and the microphone.
+For a point sound source in a direct field, the signal intensity decays in proportion to the square of the distance between the source and the microphone.
 Let $I_i$ denote the received signal intensity at microphone $i$:
@@
 Any obstacle object between the sound source and any microphone would produce a significant distortion in the measured intensity ratio.
-We find this approach to be too restrictive for our purposes, since in an interactive system we do not control whether or not the user places any obstacle between the sound source and the microphones.
+Additionally, any background noise would also potentially distort the measured intensity ratio.
+We find the ILD approach to be too restrictive for our purposes, since in an interactive system we do not control whether or not the user places any obstacle between the sound source and the microphones.
 
@@
 \item[GCC] $W(\omega) = 1$. No prefiltering is done. This is unfiltered normal cross correlation.
-\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divided by its magnitude. Only phase information contributes to delay estimation.
+\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$.
+Each frequency is divided by its magnitude.
+Only phase information contributes to delay estimation\cite{tdoa:gcc1,tdoa:gcc2,tdoa:gcc3}.
 \item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$.
